### 2024-09-02 manuscript.tex
initial commit

\documentclass[article]{ajs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{etoolbox}

%% almost as usual
\author{Matthias Templ\,\orcidlink{0000-0002-8638-5276}\\ University of Applied Sciences and Arts \\ Northwestern Switzerland \And 
        Second Author\,\orcidlink{0000-0000-0000-0000}\\ Affiliation}
%% \AND can be used instead of \And and starts a new line
\title{Template and Guidelines for Submitting a Manuscript to the Austrian Journal of Statistics}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Matthias Templ, Second Author} %% comma-separated
\Plaintitle{A Capitalized Title: Something about a Topic} %% without formatting
\Shorttitle{A Capitalized Title} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
This guidelines are tested for new versions of texlive and current version of overleaf's LaTeX version. 
%If you run into problems using ealier versions of texlive, please report to \email{matthias.templ@gmail.com}.

  This template and guide of the Austrian Journal of Statistics (AJS) shows the required style of manuscripts submitted to the Austrian Journal of Statistics. We strongly recommend
to follow the given instructions exactly already when writing the paper. When submitting papers to AJS, it is essential to follow these guidelines. The present document can be used directly as a template. The file with the name \texttt{guidelinesAJS.tex} can be obtained from the following website \href{https://github.com/matthias-da/ajs-public}{https://github.com/matthias-da/ajs-public}, or directly as overleaf template in their academic journal section.

Manuscripts submitted without consideration of the AJS style guide will most likely not be forwarded for review, but rejected with the possibility of resubmission.
}
\Keywords{keywords, comma-separated, non-capitalized, \proglang{R}}
\Plainkeywords{keywords, comma-separated, non-capitalized,, R} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}
%% \setcounter{page}{1}
\Pages{1--xx}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Matthias Templ\\
  Institue of Data Analysis and Process Design\\
  University of Applied Sciences and Arts Northwestern Switzerland\\
  CH-4600 Olten, Switzerland\\
  E-mail: \email{matthias.templ@gmail.com}\\
  URL: \url{https://www.fhnw.ch/de/personen/matthias-templ}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%% include your article here, just as usual

Before we look at style in more detail, the most common mistakes and errors are described:

\section{Comments on frequently made mistakes and concerns}

To avoid a reject in first place:

\begin{description}
\item[C1.word] Do not submit articles written with prohibited software like Microsoft Word or similar WYSWYG software. We like to send your manuscript with the AJS style to reviewers. 
\item[C1.conclusion] Write a conclusion and/or discussion discussing your results.  Discuss how your method improved science and give value to the community, and discuss differences to other alternative methods. 
\item[C1.data.documentation] Give sufficient documentation on data you are using.
\item[C1.application] In case you have numerical examples, use convincing applications and interesting data. Do not use \texttt{toy} data with 50 observations and no explanatory variables / covariables, for example. 
\end{description}

To do justice to the style from the very beginning:

\begin{description}
\item[C2.title] Title in Capitalized Style. See the title of the guidelines for illustration.
\item[C2.subsections]  Subsection in lower case style. See the subsection of the guidelines for illustration
\item[C2.references]  Captialized Title for references. E.g. \cite{fleet20} should appear in the reference list as ``Fleet Management in Free-Floating Bike Sharing Systems Using
  Predictive Modelling and Explorative Tools.'' and NOT as ``Fleet management in free-floating bike sharing systems using
  predictive modelling and explorative tools.''.
\item[C2.doi]  Include DOI numbers for each citation (if possible)
\item[C2.graphics]  Use vector-based graphics such as \texttt{pdf}. Do not use pixel-based graphics like \texttt{png}, \texttt{jpeg}, and do not convert from \texttt{png} or \texttt{jpeg} to \texttt{pdf}
\item{C2.tables} Produce tables in \LaTeX{}. (no screenshots)
\end{description}

\section{Introduction}

The style file of the Austrian Journal of Statsitics is based on the style file of the Journal of Statistical Software (JSS) including minor modifications. We thank Achim Zeileis for his great work on the JSS style file.

The Austrian Journal of Statistics is freely available over the net. Free and open-source tools used for production of your manuscript are therefore also highly supported. Code to reproduce the figures and tables in manuscripts as well as methods implemented in software can be published as supplementary material. This increases usually the number of 
downloads of your article.

After review, final sources can also be sent as \texttt{knitr} \citep{knitr} .Rnw file.

%Graphics should be inserted as PostScript files using the
%package {\tt psfig} or {\tt graphicx}. Bibliographic references
%should be given using {\tt bibtex} in the {\em author-date\/}
%style.

\section{Writing your manuscript}

In order to obtain a uniform layout and to ease the production as
much as possible it is highly recommended to use \LaTeXe\ for the
text processing.

Technically, you have three possibilities to choose the software for writing your manuscript:

- Select from the overleaf template for academic journals the template for the Austrian Journal of Statistics.
- Download the template from \href{https://github.com/matthias-da/ajs-public}{https://github.com/matthias-da/ajs-public}
- use the R package \texttt{rtcles}, function \texttt{ajs\_article}.
- use the overleaf template for the Austrian Journal of Statistics, found online at \href{https://www.overleaf.com/latex/templates/tagged/academic-journal}{https://www.overleaf.com/latex/templates/tagged/academic-journal}

Each article must have a summary in English which
inform the reader quickly on the contents and the substantial
results. The length surely depends on the contents. 
Mathematical formulas and references should be
avoided in the summary.

Headings and footings as well as pagination will be added during
the production. Footnotes should be avoided.

The article should be structured by enumerated sections and
subsections. The title should be in title case style, sections and subsections in section case style. Graphics and tables must contain captions, above
tables and below graphics (figures). Examples are (see Figure~\ref{fig:Fig3.4} and Table~\ref{Tab3.4}).
Graphics \textbf{should be produced in \textbf{pdf} format} and included 
 at suitable places of the manuscript.

\begin{figure}[hbt]
\begin{center}
\includegraphics{ajs-download20180210}
\caption{\label{fig:Fig3.4}Download statistics of articles per month at the Austrian Journal of Statistics.}  
\end{center}
\end{figure}

\begin{table}[hbt]
\caption{\label{Tab3.4}80 measured values of access time.}
\vspace*{-5mm}
\small
\hspace{1.5cm}
\begin{center}
\begin{tabular}{|llllllll|}
\toprule
70.0 & 68.6 & 67.9 & 66.3 & 71.0 & 64.2 & 69.6 & 71.0\\
... & ... & ... & ... & ... & ... & ... & ...\\
%69.7 & 69.0 & 73.4 & 69.0 & 70.1 & 69.8 & 69.0 & 73.0\\
%70.0 & 69.1 & 69.5-& 67.9 & 72.8 & 72.1 & 69.5+& 70.1\\
%70.2 & 69.1 & 68.9 & 68.3 & 74.9 & 68.4 & 69.1 & 66.6\\
%71.1 & 69.2 & 71.2 & 68.9 & 70.9 & 70.6 & 69.9 & 69.9\\
%69.4 & 69.5+& 68.5+& 70.9 & 71.6 & 68.9 & 72.0 & 70.3\\
%68.6 & 68.5-& 67.8 & 72.2 & 68.7 & 70.6 & 66.9 & 69.3\\
%71.4 & 68.7 & 74.2 & 68.8 & 71.4 & 71.8 & 67.5-& 70.4\\
%71.4 & 67.4 & 69.5-& 72.4 & 70.4 & 69.3 & 68.2 & 67.0\\
71.7 & 70.5+& 72.5-& 68.2 & 67.6 & 68.6 & 70.5-& 65.8\\
\bottomrule
\end{tabular}
\end{center}
\end{table}

Mathematical formulas should be enumerated as far as they are
referenced \textbf{only}. E.g.~Equation~(\ref{form3.1}) shows the density 
\begin{equation}\label{form3.1}
f(x) =
\frac1{\sqrt{2\pi}\sigma}e^{-\frac{(x-\mu)^2}{2\sigma^2}}\,.
\end{equation}

Please consider

\begin{itemize}
\item  capitalized title case for titles and for all references and 
\item sentence, non-capitalized case for sections and subsections.
\end{itemize}

Cited literature should be summarized at the end of the article in
form of a non-enumerated section. \cite{knitr} is a reference
to an article. If the citation is in parentheses, it
should look like \citep{knitr}. The latter verison is used for indirect citations while the first one to directly cite an article.
References should be in title case.

Correspondence addresses of author(s) should be added at the end
of the manuscript.

\section{Submission and refereeing}

\subsection{Submission}

Consideration of the previously stated rules are of great help for
the editing process. Whenever technically possible, manuscripts
should be submitted through the editorial system at \href{www.ajs.or.at}{www.ajs.or.at}. The authors must be registered at this website.

\subsection{Refereing}

All contributions will be anonymously refereed. Cited literature which is
hardly available should accompany the submitted manuscript. It
should also be considered to place used and analyzed data at
disposal for the referees (if there are no legal or technical
arguments against).

Editor and referees must trust that the contribution has not been
submitted for publication at the same time at another place. It is
fair that the submitting author notifies if an earlier version has
already been submitted somewhere before.

\subsection{Copyrights and fees}

The Austrian Journal of Statistics is free of
charge. Generally, we use the Creative Commons Attribution License (CC-BY).

For copyright issues, publication ethics and malpractice statements, please have a look at 
\href{www.ajs.or.at}{www.ajs.or.at}

\section{Checklist}

We list the most common mistakes when submitting manuscripts to the journal and formulate them in form of questions.

\begin{itemize}
\item Did you upload a pdf of your manuscript for review? (no Word, no .tex, ...)
\item Did you use the style of AJS?  
\item Are your graphics produced in pdf format? 
\item Is your title captialized?
\item Are the titles of your cited papers in your references capitalized?
\item Are your section and subsection headers non-capitalized form?
\item Are your tables produced with \LaTeX ? (no screenshots)
\item Have you cited the software you used, e.g. R itself and R packages used? To not cite software is plagiarism and violates intellectual properties of the authors of the software.
\item Did you proposed at least three independent reviewers?
\end{itemize}

\section{Frequently asked questions}

\noindent 
\begin{enumerate}
\item \textit{My citation shows all authors and not the typically shown et al.}. \\ Answer: this is ok. We like to show all authors in the text.
\item \textit{When I submit a paper, how long does it take to be published} \\ Answer: The average time from submission to publication is approx. 132 days. But it heavily depends if you can make satisfying revisions in time. 
\item \textit{In which databases is AJS indexed.} \\
Answer: we are indexed in many of them, most notable is Scopus (from Elsevier). More information can be found at the website of AJS. 
\item \textit{Can I upload my paper on \href{https://arxiv.org/}{https://arxiv.org/} and distribute before the paper is online at AJS? } \\
Answer: Of course, you can do it. However, please clearly point out in the header of the paper that this paper is under review in the Austrian Journal of Statistics and will likely be published there. Note that the readers should check if the paper is already be published at AJS for correct citation, i.e. finally the journal should be cited. Update and upload  \href{https://arxiv.org/}{https://arxiv.org/} or similar sites with the final version in AJS style after the paper is published  in AJS.

\end{enumerate}

\section{Known issues}

While in the documentclass arcticle a section header with special characters works without problems, e.g. \verb|\section{A \& B}|, this results in an error in the documentclass ajs. The solution is \verb|\section[A and B]{A \& B}|.

%\bibliographystyle{plainat}
\bibliography{template-guidelinesAJS}

\end{document}

### 2024-09-02 manuscript.tex
basic changes

@@
 %% almost as usual
-\author{Matthias Templ\,\orcidlink{0000-0002-8638-5276}\\ University of Applied Sciences and Arts \\ Northwestern Switzerland \And 
-        Second Author\,\orcidlink{0000-0000-0000-0000}\\ Affiliation}
+\author{Matthias Medl\,\orcidlink{0000-0002-3354-4579}\\ Institute of Statistics \\ BOKU University \\ Vienna \And 
+        Dianne Cook\,\orcidlink{0000-0002-3813-7155}\\ Monash Data Futures Institute \\ Monash University \\ Melbourne \And
+        Ursula Laa\,\orcidlink{0000-0002-0249-6439}\\ Institute of Statistics \\ BOKU University \\ Vienna}
 %% \AND can be used instead of \And and starts a new line
-\title{Template and Guidelines for Submitting a Manuscript to the Austrian Journal of Statistics}
+\title{pytourr: Interactive Visualization Interface for the tourr Package}
 
 %% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Matthias Templ, Second Author} %% comma-separated
-\Plaintitle{A Capitalized Title: Something about a Topic} %% without formatting
-\Shorttitle{A Capitalized Title} %% a short title (if necessary)
+\Plainauthor{Matthias Medl, Dianne Cook, Ursula Laa} %% comma-separated
+\Plaintitle{pytourr: Interactive Visualization Interface for the tourr Package} %% without formatting
+\Shorttitle{pytourr} %% a short title (if necessary)
 
@@
 }
-\Keywords{keywords, comma-separated, non-capitalized, \proglang{R}}
+\Keywords{interactive visualization, tourr, exploratory data analysis, \proglang{R}, \proglang{python}}
 \Plainkeywords{keywords, comma-separated, non-capitalized,, R} %% without formatting

### 2024-09-05 manuscript.tex
major initial restructuring

@@
 \Abstract{
-This guidelines are tested for new versions of texlive and current version of overleaf's LaTeX version. 
-%If you run into problems using ealier versions of texlive, please report to \email{matthias.templ@gmail.com}.
-
-  This template and guide of the Austrian Journal of Statistics (AJS) shows the required style of manuscripts submitted to the Austrian Journal of Statistics. We strongly recommend
-to follow the given instructions exactly already when writing the paper.
-When submitting papers to AJS, it is essential to follow these guidelines.
-The present document can be used directly as a template.
-The file with the name \texttt{guidelinesAJS.tex} can be obtained from the following website \href{https://github.com/matthias-da/ajs-public}{https://github.com/matthias-da/ajs-public}, or directly as overleaf template in their academic journal section.
-
-Manuscripts submitted without consideration of the AJS style guide will most likely not be forwarded for review, but rejected with the possibility of resubmission.
+Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum
+Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum
+Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum
+Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum
+Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum
+Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum
+Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum
+Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum
 }
-\Keywords{interactive visualization, tourr, exploratory data analysis, \proglang{R}, \proglang{python}}
+\Keywords{interactive graphics, tourr, exploratory data analysis, \proglang{R}, \proglang{python}}
 \Plainkeywords{keywords, comma-separated, non-capitalized,, R} %% without formatting
@@
 \Address{
-  Matthias Templ\\
-  Institue of Data Analysis and Process Design\\
-  University of Applied Sciences and Arts Northwestern Switzerland\\
-  CH-4600 Olten, Switzerland\\
-  E-mail: \email{matthias.templ@gmail.com}\\
-  URL: \url{https://www.fhnw.ch/de/personen/matthias-templ}
+  Matthias Medl\\
+  Institue of Statistics\\
+  BOKU University Vienna\\
+  E-mail: \email{matthias.medl@boku.ac.at}\\
 }
@@
 
-%% include your article here, just as usual
-
-Before we look at style in more detail, the most common mistakes and errors are described:
-
-\section{Comments on frequently made mistakes and concerns}
-
-To avoid a reject in first place:
-
-\begin{description}
-\item[C1.word] Do not submit articles written with prohibited software like Microsoft Word or similar WYSWYG software. We like to send your manuscript with the AJS style to reviewers. 
-\item[C1.conclusion] Write a conclusion and/or discussion discussing your results.
-Discuss how your method improved science and give value to the community, and discuss differences to other alternative methods.
-\item[C1.data.documentation] Give sufficient documentation on data you are using.
-\item[C1.application] In case you have numerical examples, use convincing applications and interesting data.
-Do not use \texttt{toy} data with 50 observations and no explanatory variables / covariables, for example.
-\end{description}
-
-To do justice to the style from the very beginning:
-
-\begin{description}
-\item[C2.title] Title in Capitalized Style. See the title of the guidelines for illustration.
-\item[C2.subsections]  Subsection in lower case style. See the subsection of the guidelines for illustration
-\item[C2.references]  Captialized Title for references. E.g. \cite{fleet20} should appear in the reference list as ``Fleet Management in Free-Floating Bike Sharing Systems Using
-  Predictive Modelling and Explorative Tools.'' and NOT as ``Fleet management in free-floating bike sharing systems using
-  predictive modelling and explorative tools.''.
-\item[C2.doi]  Include DOI numbers for each citation (if possible)
-\item[C2.graphics]  Use vector-based graphics such as \texttt{pdf}.
-Do not use pixel-based graphics like \texttt{png}, \texttt{jpeg}, and do not convert from \texttt{png} or \texttt{jpeg} to \texttt{pdf}
-\item{C2.tables} Produce tables in \LaTeX{}. (no screenshots)
-\end{description}
-
 \section{Introduction}
 
-The style file of the Austrian Journal of Statsitics is based on the style file of the Journal of Statistical Software (JSS) including minor modifications.
-We thank Achim Zeileis for his great work on the JSS style file.
+\section{Similar work}
 
-The Austrian Journal of Statistics is freely available over the net.
-Free and open-source tools used for production of your manuscript are therefore also highly supported.
-Code to reproduce the figures and tables in manuscripts as well as methods implemented in software can be published as supplementary material.
-This increases usually the number of
-downloads of your article.
+Previous attempts have been made to bring interactivtiy to the tourr package.
+The galahr package \citep{galahr} was an attempt to so by means of the R packages shiny \citep{shiny} and plotly \citep{plotly}, but ultimately suffered from performance issues when the size of the investigated dataset became too large (n>10 000).
+To overcome these issues, the detourr package \citep{detourr}, which precomputes multiple projections in R and subsequently visualizes them in an interactive display using Javascript, has been developed.
+While detourr showed better performance, the interactivtiy is still limited as only precomputed projections can be displayed and there is no option to show multiple linked displays simultaneously.
+The mmtour package \citep{laa2023new} is a \proglang{Mathamatica} implementation allowing for interaction with manual and slice tours.
 
-After review, final sources can also be sent as \texttt{knitr} \citep{knitr} .Rnw file.
+\section{The pytourr package}
 
-%Graphics should be inserted as PostScript files using the
-%package {\tt psfig} or {\tt graphicx}. Bibliographic references
-%should be given using {\tt bibtex} in the {\em author-date\/}
-%style.
+\subsection{Overview}
 
-\section{Writing your manuscript}
+The primary objective of this work is to develop an R package that integrates high-performance interactivity through Python packages, such as \texttt{TKinter} \citep{lundh1999introduction}, \texttt{CustomTKinter} \citep{schimansky24}, and \texttt{matplotlib} \citep{Hunter:2007}, with the tourr package \citep{tourr}.
+This integration is achieved via \texttt{reticulate} \citep{reticulate}, a framework that facilitates seamless interoperability between Python and R.
 
-In order to obtain a uniform layout and to ease the production as
-much as possible it is highly recommended to use \LaTeXe\ for the
-text processing.
+At the core of the \texttt{pytourr} package is a graphical user interface (GUI) designed to display multiple linked visualizations concurrently.
+Users can interact with one of the plots, for example, by subselecting specific data points, and the other linked plots will dynamically update to reflect these modifications.
+This interconnected functionality allows for efficient and immediate analysis of high-dimensional datasets without necessitating extensive coding, thereby streamlining the exploratory data analysis process.
 
-Technically, you have three possibilities to choose the software for writing your manuscript:
+In addition to this core functionality, \texttt{pytourr} offers a variety of interactive plot types, providing users with the flexibility to visualize their data from multiple perspectives.
+The ability to navigate through various projections of the displayed tours directly within the GUI enables users to explore different aspects of the dataset with ease.
+Furthermore, users can initiate new tours directly from the interface.
+The GUI also supports interactive feature selection, allowing users to specify which subset of features should be visualized in the plots.
+Once users have identified interesting views or settings, \texttt{pytourr} allows them to save the displayed projections, subsets, and plots.
+This functionality ensures that analysis states can be preserved for further examination or reporting, making the package particularly useful for iterative analysis where findings may need to be revisited or shared with collaborators.
 
-- Select from the overleaf template for academic journals the template for the Austrian Journal of Statistics.
-- Download the template from \href{https://github.com/matthias-da/ajs-public}{https://github.com/matthias-da/ajs-public}
-- use the R package \texttt{rtcles}, function \texttt{ajs\_article}.
-- use the overleaf template for the Austrian Journal of Statistics, found online at \href{https://www.overleaf.com/latex/templates/tagged/academic-journal}{https://www.overleaf.com/latex/templates/tagged/academic-journal}
+With its high level of interactivity, performance, and ease of use, \texttt{pytourr} streamlines the exploration of complex datasets, offering a powerful tool for researchers working with high-dimensional data.
 
-Each article must have a summary in English which
-inform the reader quickly on the contents and the substantial
-results. The length surely depends on the contents. 
-Mathematical formulas and references should be
-avoided in the summary.
+\subsection{The graphical user interface}
 
-Headings and footings as well as pagination will be added during
-the production. Footnotes should be avoided.
+The \texttt{pytourr} GUI can be launched using the function \texttt{interactive\textunderscore tour()}.
+At a minimum, the user needs to provide both the dataset and the instructions for constructing the desired plots.
+The dataset must be supplied as a \texttt{data.table}, while the plotting instructions should be passed as a list containing the named elements \texttt{type} and \texttt{obj}.
+The \texttt{type} element specifies the type of display to generate, such as "scatter" for a scatterplot or "2d\textunderscore tour" for a 2-dimensional tour.
+The \texttt{obj} element further defines the properties of the chosen display.
+For example, to create a 2-dimensional tour, the user must provide a \texttt{tour\textunderscore history} object, which can be generated using the \texttt{tourr} package.
+For a scatterplot, the user needs to provide a vector of strings specifying the names of the features to be displayed.
+The user can optionally specify the feature names, the arrangement of the plots, predefined subsets of the data (e.g., cluster solutions), custom names for these subsets, the number of available subsets, and the size of each plot.
 
-The article should be structured by enumerated sections and
-subsections. The title should be in title case style, sections and subsections in section case style. Graphics and tables must contain captions, above
-tables and below graphics (figures). Examples are (see Figure~\ref{fig:Fig3.4} and Table~\ref{Tab3.4}).
-Graphics \textbf{should be produced in \textbf{pdf} format} and included 
- at suitable places of the manuscript.
+The GUI is divided into two main sections: a sidebar on the left, which contains a comprehensive set of interactive controls to customize and interact with the displays, and the display area on the right.
+An overview of the GUI can be seen in figure x.
 
-\begin{figure}[hbt]
-\begin{center}
-\includegraphics{ajs-download20180210}
-\caption{\label{fig:Fig3.4}Download statistics of articles per month at the Austrian Journal of Statistics.}  
-\end{center}
-\end{figure}
+At the top of the sidebar, users can select and deselect features via checkboxes, determining which features are displayed in the plots.
+Below this, each subset of the data has a corresponding checkbox, which is used to select the active subset.
+When a subset is selected, any data points marked in the plots will be moved to this active subset and be colored accordingly.
+The colored boxes next to the subset names indicate how the datapoints within each subset are colored.
+Users can adjust the transparency of each subset by clicking these colored boxes.
+This can be useful when one wants to highlight one or multiple subsets, for instance when wanting to compare to subsets in particular.
+The subset selection can be reset to its original state at any time by clicking the "Reset original selection" button.
 
-\begin{table}[hbt]
-\caption{\label{Tab3.4}80 measured values of access time.}
-\vspace*{-5mm}
-\small
-\hspace{1.5cm}
-\begin{center}
-\begin{tabular}{|llllllll|}
-\toprule
-70.0 & 68.6 & 67.9 & 66.3 & 71.0 & 64.2 & 69.6 & 71.0\\
-... & ... & ... & ... & ... & ... & ... & ...\\
-%69.7 & 69.0 & 73.4 & 69.0 & 70.1 & 69.8 & 69.0 & 73.0\\
-%70.0 & 69.1 & 69.5-& 67.9 & 72.8 & 72.1 & 69.5+& 70.1\\
-%70.2 & 69.1 & 68.9 & 68.3 & 74.9 & 68.4 & 69.1 & 66.6\\
-%71.1 & 69.2 & 71.2 & 68.9 & 70.9 & 70.6 & 69.9 & 69.9\\
-%69.4 & 69.5+& 68.5+& 70.9 & 71.6 & 68.9 & 72.0 & 70.3\\
-%68.6 & 68.5-& 67.8 & 72.2 & 68.7 & 70.6 & 66.9 & 69.3\\
-%71.4 & 68.7 & 74.2 & 68.8 & 71.4 & 71.8 & 67.5-& 70.4\\
-%71.4 & 67.4 & 69.5-& 72.4 & 70.4 & 69.3 & 68.2 & 67.0\\
-71.7 & 70.5+& 72.5-& 68.2 & 67.6 & 68.6 & 70.5-& 65.8\\
-\bottomrule
-\end{tabular}
-\end{center}
-\end{table}
+At the top of the sidebar, users can select and deselect features via checkboxes, controlling which features are displayed in the plots.
+Below the feature selection, each subset has a checkbox that designates the active subset.
+When datapoints within the plots are manually selected, they will be moved to the active subset and colored accordingly.
+The colored boxes next to the subset names indicate the colors that the datapoints within each subset are colored.
+By clicking these colored boxes, users can adjust the transparency of the datapoints within the corrisponding subet, which is particularly useful for highlighting and comparing specific subsets.
+The initial subset selection (in case one was provided) can be restored at any time by clicking the "Reset original selection" button.
 
-Mathematical formulas should be enumerated as far as they are
-referenced \textbf{only}. E.g.~Equation~(\ref{form3.1}) shows the density 
-\begin{equation}\label{form3.1}
-f(x) =
-\frac1{\sqrt{2\pi}\sigma}e^{-\frac{(x-\mu)^2}{2\sigma^2}}\,.
-\end{equation}
+One can move through the projections of a tour by pressing the arrow keys or by specifying the index of the projection the be displayed and pressing the "Update frames" button.
+Users can animate the tours by toggling the "Animate" checkbox and specifying a time interval, so that the displays automatically move to the next projections after the specified time interval.
 
-Please consider
+When clicking the "Save projections and subsets" button, a file browser is first launched, allowing users to specify the destination for the output files.
+These files include a PNG image of the current plots, CSV files containing the projection matrices for all displayed tours, and a mapping of the data points to their respective subsets.
+Additionally, users can initiate new tours of various directly from the sidebar.
 
-\begin{itemize}
-\item  capitalized title case for titles and for all references and 
-\item sentence, non-capitalized case for sections and subsections.
-\end{itemize}
+\subsection{R/Python interface}
 
-Cited literature should be summarized at the end of the article in
-form of a non-enumerated section. \cite{knitr} is a reference
-to an article. If the citation is in parentheses, it
-should look like \citep{knitr}. The latter verison is used for indirect citations while the first one to directly cite an article.
-References should be in title case.
+The majority of \texttt{pytourr} was written in Python, while the R side of the package handles setting up the Python environment within the interactive interface is being run, launching the interactive tour, and generating new tours when initiated through the interface.
+To incorporate the functionality of the \texttt{tourr} package without translating large portions of its code from R to Python, the \texttt{reticulate} package was used.
+This approach allows \texttt{pytourr} to automatically benefit from updates to \texttt{tourr}.
 
-Correspondence addresses of author(s) should be added at the end
-of the manuscript.
+However, to reduce inefficiencies caused by cross-language communication and to simplify debugging, \texttt{tourr} functions were only accessed when necessary.
+Core functionalities, such as performing data transformations using projections, were implemented directly in Python.
+These transformations are based on well-established mathematical principles and are straightforward to replicate, ensuring they remain stable and efficient.
 
-\section{Submission and refereeing}
+\subsection{Structure of the Python code}
 
-\subsection{Submission}
+The central component of the Python code is the \texttt{customTKinter} class \texttt{InteractiveTourInterface}.
+This class centrally stores attributes related to all plots, such as the dataset, subselections, feature selections, and other shared information.
+Plot-specific data is organized in dictionaries (the Python equivalent of named lists in R), including the display type, construction instructions, tour projections (only in case of tour displays), color schemes for the displayed data, and, where applicable, the selector and manual projection manipulation classes.
 
-Consideration of the previously stated rules are of great help for
-the editing process. Whenever technically possible, manuscripts
-should be submitted through the editorial system at \href{www.ajs.or.at}{www.ajs.or.at}. The authors must be registered at this website.
+The selector classes handle the behavior when users manually select data points to move them to the active subset.
+After a selection is made, the selector class updates the centrally stored subselection attribute and ensures all other displays reflect these changes.
+The manual projection manipulation classes construct the arrows representing the projection axes in the displays and manage the manual adjustment of projections.
+Users can right-click and drag the arrowheads to modify the projections, after which the class orthonormalizes the projection axes and updates both the projection and the transformed data accordingly.
 
-\subsection{Refereing}
+\subsection{Bit blit}
 
-All contributions will be anonymously refereed. Cited literature which is
-hardly available should accompany the submitted manuscript. It
-should also be considered to place used and analyzed data at
-disposal for the referees (if there are no legal or technical
-arguments against).
+The implementation of bit blitting was crucial to ensuring fast plot updates and providing a smooth user experience.
+With bit blit, the static elements of the display, such as the outer frames of the plots, are stored as a background image.
+When a plot is manipulated, only the affected plot is updated, and within that, only the interactive elements, such as the data points and projection axes in a 2-dimensional tour, are rendered on top of the background image.
 
-Editor and referees must trust that the contribution has not been
-submitted for publication at the same time at another place. It is
-fair that the submitting author notifies if an earlier version has
-already been submitted somewhere before.
+In practice, this means that the background, without the interactive elements, must be captured either during initialization or after major updates.
+The entire plot is first rendered without the interactive elements, the background is then saved, and finally, the plot is redrawn with the interactive elements blended in.
+Since this process is relatively slow, full updates are only triggered during initialization or after significant changes, such as modifications to the set of active features.
 
-\subsection{Copyrights and fees}
+\section{Datasets}
 
-The Austrian Journal of Statistics is free of
-charge. Generally, we use the Creative Commons Attribution License (CC-BY).
+Two datasets were analysed as part of this study. The Austrian Vacation Activities \cite{dolnicar2003winter} dataset and the the Australian Vacation Activities \cite{cliff2009formative} dataset.
 
-For copyright issues, publication ethics and malpractice statements, please have a look at 
-\href{www.ajs.or.at}{www.ajs.or.at}
+\subsection{Austrian Vacation Activities dataset}
 
-\section{Checklist}
+The Austrian Vacation Activities dataset comprises responses from 2,961 adult tourists who spent their holiday in Austria during the 1997/98 season.
+Participants were asked to evaluate the importance of 27 different activities during their vacation.
+The survey categorized responses based on four levels of importance: "totally important," "mostly important," "a bit important," and "not important." For analysis, the responses were binarized: a value of 1 was assigned if the activity was rated as "totally important," and a value of 0 if any of the other categories were selected.
+The survey was conducted by the Europäisches Tourismus Institut GmbH at the University of Trier and focused exclusively on tourists who did not stay in the country's capital cities.
 
-We list the most common mistakes when submitting manuscripts to the journal and formulate them in form of questions.
+\subsection{Australian Vacation Activities dataset}
 
-\begin{itemize}
-\item Did you upload a pdf of your manuscript for review? (no Word, no .tex, ...)
-\item Did you use the style of AJS?  
-\item Are your graphics produced in pdf format? 
-\item Is your title captialized?
-\item Are the titles of your cited papers in your references capitalized?
-\item Are your section and subsection headers non-capitalized form?
-\item Are your tables produced with \LaTeX ? (no screenshots)
-\item Have you cited the software you used, e.g. R itself and R packages used? To not cite software is plagiarism and violates intellectual properties of the authors of the software.
-\item Did you proposed at least three independent reviewers?
-\end{itemize}
+The second dataset, the Australian Vacation Activities dataset, includes responses from 1,003 adult Australians who were surveyed through a permission-based internet panel.
+The survery was conducted in 2007.
+Participants were asked whether they engaged in 44 specific vacation activities during their most recent vacation within Australia.
+Similar to the Austrian dataset, responses were binarized: a value of 1 indicates that the participant took part in the activity, while a value of 0 signifies they did not.
 
-\section{Frequently asked questions}
+\section{Examples}
 
-\noindent 
-\begin{enumerate}
-\item \textit{My citation shows all authors and not the typically shown et al.}. \\ Answer: this is ok. We like to show all authors in the text.
-\item \textit{When I submit a paper, how long does it take to be published} \\ Answer: The average time from submission to publication is approx.
-132 days.
-But it heavily depends if you can make satisfying revisions in time.
-\item \textit{In which databases is AJS indexed.} \\
-Answer: we are indexed in many of them, most notable is Scopus (from Elsevier). More information can be found at the website of AJS. 
-\item \textit{Can I upload my paper on \href{https://arxiv.org/}{https://arxiv.org/} and distribute before the paper is online at AJS? } \\
-Answer: Of course, you can do it.
-However, please clearly point out in the header of the paper that this paper is under review in the Austrian Journal of Statistics and will likely be published there.
-Note that the readers should check if the paper is already be published at AJS for correct citation, i.e.
-finally the journal should be cited.
-Update and upload  \href{https://arxiv.org/}{https://arxiv.org/} or similar sites with the final version in AJS style after the paper is published  in AJS.
+\section{Discussion}
 
-\end{enumerate}
-
-\section{Known issues}
-
-While in the documentclass arcticle a section header with special characters works without problems, e.g.
-\verb|\section{A \& B}|, this results in an error in the documentclass ajs.
-The solution is \verb|\section[A and B]{A \& B}|.
+\section{Conclusion}
 
 %\bibliographystyle{plainat}
-\bibliography{template-guidelinesAJS}
+\bibliography{pytourr_references.bib}
 

### 2024-09-06 manuscript.tex
small changes

@@
 \author{Matthias Medl\,\orcidlink{0000-0002-3354-4579}\\ Institute of Statistics \\ BOKU University \\ Vienna \And 
-        Dianne Cook\,\orcidlink{0000-0002-3813-7155}\\ Monash Data Futures Institute \\ Monash University \\ Melbourne \And
+        Dianne Cook\,\orcidlink{0000-0002-3813-7155}\\ Econometrics and Business Statistics \\ Monash University \\ Melbourne \And
         Ursula Laa\,\orcidlink{0000-0002-0249-6439}\\ Institute of Statistics \\ BOKU University \\ Vienna}
 %% \AND can be used instead of \And and starts a new line
-\title{pytourr: Interactive Visualization Interface for the tourr Package}
+\title{Interactively Visualizing Clustering in Multivariate Data Using pytourr}
 
@@
 \Plainauthor{Matthias Medl, Dianne Cook, Ursula Laa} %% comma-separated
-\Plaintitle{pytourr: Interactive Visualization Interface for the tourr Package} %% without formatting
+\Plaintitle{Interactively Visualizing Clustering in Multivariate Data Using pytourr} %% without formatting
 \Shorttitle{pytourr} %% a short title (if necessary)
@@
 
-\section{Similar work}
+\begin{itemize} \itemsep 0in
+\item Background to clustering analysis
+\item Visualising clusters, including tours
+\item Difference between clustering and partitioning
+\item How do you know whether the results are useful
+\end{itemize}
+
+\section{Related Work}
 
@@
 
-\section{Datasets}
+\section{Applications}
 
-Two datasets were analysed as part of this study. The Austrian Vacation Activities \cite{dolnicar2003winter} dataset and the the Australian Vacation Activities \cite{cliff2009formative} dataset.
+The Austrian Vacation Activities \cite{dolnicar2003winter} and the Australian Vacation Activities \cite{cliff2009formative} datasets are used to illustrate the methods.
 

### 2024-09-09 manuscript.tex
first draft of winterActiv data analysis

@@
 \usepackage{etoolbox}
+\usepackage{amsmath}
+\usepackage{graphicx}
+\graphicspath{{images/}}
 
@@
 
+To gain further insight into the dataset a k-means clustering as described in the book "Market Segmentation Analysis" has been performed \citep{leisch2018market}.
+Therefore, the function \texttt{stepcclust} of the R package \texttt{flexclust
+}\citep{flexclust} with k=6 and nrep=20 was used.
+
+\subsubsection{Feature selection}
+
+The original dataset contains \(p\)=27 features.
+Some of these features are more informative than others.
+We only want to keep the most informative ones.
+Additionally, interacting with the pytourr GUI becomes cumbersome when handling more than \textasciitilde
+ 15 features, making it necessary to reduce the dimensionality of the dataset.
+An effective and intuitive way to perform feature selection is by using the heatmap display within pytourr.
+The heatmap supports three different display metrics for feature selection:
+
+\[
+\text{Total Fraction:} \quad \frac{\displaystyle\sum_{i = 1}^{k} \sum_{j = 1}^{p} c_{ij}}{n}
+\]
+The \textbf{Total Fraction} shows the counts \(c_{ij}\) of the \(i\)-th cluster and \(j\)-th feature, normalized by the total number of observations \(n\).
+
+\[
+\text{Intra Cluster Fraction:} \quad \sum_{i = 1}^{k} \sum_{j = 1}^{p} \frac{c_{ij}}{c_{i}}
+\]
+The \textbf{Intra Cluster Fraction} displays the contribution of each feature within a specific cluster, providing insight into how much a feature contributes to the cluster compared to the overall cluster size \(c_i\).
+
+\[
+\text{Intra Feature Fraction:} \quad \sum_{j = 1}^{p} \sum_{i = 1}^{k} \frac{c_{ij}}{c_{j}}
+\]
+The \textbf{Intra Feature Fraction} highlights how a feature contributes across different clusters relative to the overall importance of that feature \(c_j\).
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{winteractiv_heatmap.png}
+    \caption{A: heatmap displaying the intra cluster fraction. B: heatmap displaying the intra feature fraction.}
+    \label{fig:winteractiv_heatmap}
+\end{figure}
+
+In Figure \ref{fig:winteractiv_heatmap}A, we can observe the general interests of tourists within each cluster.
+For instance, tourists in clusters 1, 3, 5, and 6 predominantly engaged in alpine skiing, while those in clusters 2 and 4 did not.
+Additionally, we see that activities such as ski touring and horseback riding were generally unpopular across all clusters.
+
+In Figure \ref{fig:winteractiv_heatmap}B, we can determine whether tourists in a particular cluster had a strong preference for certain activities.
+For example, nearly all tourists who visited museums are grouped in cluster 2, and those who used health facilities are primarily attributed to cluster 4.
+We can also identify activities that were similarly popular across all clusters, such as relaxing.
+
+By using the heatmap with these metrics, we can perform feature selection by removing unpopular activities and those that were consistently similar across all clusters.
+After performing the feature selection by unchecking the corresponding checkboxes in the GUI using this strategy, the following 12 activities remained: alpine skiing, going to a spa, using health facilities, hiking, going for walks, excursions, going out in the evening, going to discos/bars, shopping, sightseeing, museums, and pool/sauna.
+
+We can now repeat the k-means clustering with \texttt{stepcclust} on the reduced dataset.
+Silhouette plots of both cluster solutions can be seen in Figure \ref{fig:silhouette_comparison}.
+By comparing both silhouette plots, we can see that the cluster solution with the reduced dataset results in a clustering of higher quality.
+Thus, we will continue with the analysis on the reduced dataset with the corresponsig cluster solution.
+We can also see in Figure \ref{fig:silhouette_comparison}B that cluster 3 is of comparatively low quality.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{silhouette_comparison.png}
+    \caption{A: silhouette plot of the k-means solution of the full dataset. B: silhouette plot of the k-means solution with the reduced dataset.}
+    \label{fig:silhouette_comparison}
+\end{figure}
+
+We can further explore the similarities between the clusters by initializing an \texttt{interactive\_tour()} with a 2D tour based on the linear discriminant analysis (LDA) projection pursuit index.
+By navigating through the tour, we can observe various projections, and when a projection that separates the clusters is found, we highlight each cluster sequentially.
+The different highlighted clusters can be seen in Figure \ref{fig:silhouette_comparison}.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{winter_activ_cluster_highlights.png}
+    \caption{Cluster highlights: \textbf{A} - Cluster 1 (blue), \textbf{B} - Cluster 2 (orange), \textbf{C} - Cluster 3 (green), \textbf{D} - Cluster 4 (red), \textbf{E} - Cluster 5 (violet), \textbf{F} - Cluster 6 (brown).}
+    \label{fig:winter_activ_cluster_highlights}
+\end{figure}
+
+This process allows us to visually assess the separation and similarities between the clusters, providing insight into the structure of the dataset.
+By highlighting each cluster individually, we can evaluate their distinctiveness in different projections.
+The most influential features shown in Figure \ref{fig:winter_activ_cluster_highlights} are pool/sauna, alpine skiing, museums, and going to the spa.
+The projection roughly separates clusters 1, 2, and 3 from each other and the other three clusters, which appear to be quite similar.
+
+By manually manipulating the projection axes or initiating a local tour, we can gain further insight into the similarities between the different clusters.
+This interactive exploration allows for a more nuanced understanding of the relationships between clusters and the influence of key features on the separation of the data.
+
+\subsubsection{Manual cluster improvement}
+
+We have seen in Figure \ref{fig:silhouette_comparison} that clusters 1 and 3 contain some data points that do not fit well into their respective clusters.
+Manual selection of data points can be employed to further improve the quality of the cluster solution.
+To do this, we first initialize an \texttt{interactive\_tour()} with the following components:
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{winter_cl7_init.png}
+    \caption{Top left: 2D tour with the linear discriminant analysis projection pursuit index.
+Top right: heatmap with the intra-cluster fraction.
+Bottom left: 1D tour with the linear discriminant analysis projection pursuit index.
+Bottom right: mosaic plot.}
+    \label{fig:winter_cl7_init}
+\end{figure}
+
+\begin{itemize}
+    \item A 2D tour with the linear discriminant analysis projection pursuit index,
+    \item A heatmap displaying the intra-cluster fraction,
+    \item A 1D tour with the linear discriminant analysis projection pursuit index, and
+    \item A mosaic plot.
+\end{itemize}
+
+This setup produces the display shown in Figure \ref{fig:winter_cl7_init}.
+We can observe that clusters 1 and 3 overlap significantly in the features "sightseeing," "museums," and "going out in the evenings." By highlighting clusters 1 and 3, and adjusting the projection axes, we can confirm that there is indeed overlap between these clusters, as shown in Figure \ref{fig:winter_cl7_pre}.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{winter_cl7_pre.png}
+    \caption{Top left: 2D tour with the linear discriminant analysis projection pursuit index.
+Top right: heatmap with the intra-cluster fraction.
+Bottom left: 1D tour with the linear discriminant analysis projection pursuit index.
+Bottom right: mosaic plot.}
+    \label{fig:winter_cl7_pre}
+\end{figure}
+
+To improve the quality of the k-means solution, we can select the checkmark for subset 7 and manually select the region of overlap.
+The selected data points will then be shifted to a new cluster: cluster 7.
+The interface with the newly formed cluster can be seen in Figure \ref{fig:winter_cl7_post}.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{winter_cl7_post.png}
+    \caption{Top left: 2D tour with the linear discriminant analysis projection pursuit index.
+Top right: heatmap with the intra-cluster fraction.
+Bottom left: 1D tour with the linear discriminant analysis projection pursuit index.
+Bottom right: mosaic plot.}
+    \label{fig:winter_cl7_post}
+\end{figure}
+
+To assess the quality of the manually enhanced cluster solution, we can compare the silhouette plots of this new solution with those from a k-means solution where \(k = 7\).
+These silhouette plots are shown in Figure \ref{fig:silhouette_comparison_k7_man}.
+It is evident that the manually enhanced cluster solution better describes the data compared to the k-means solution with \(k = 7\).
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{silhouette_comparison_k7_man.png}
+    \caption{A: Silhouette plot of the k-means solution with \(k=7\). B: Silhouette plot of the manually enhanced k-means solution with \(k=6\).}
+    \label{fig:silhouette_comparison_k7_man}
+\end{figure}
+
+Additionally, manually enhancing a cluster solution has the advantage of maintaining the base structure of the clustering, whereas clusters in a new k-means solution can be entirely different.
+Moreover, tourists in cluster 7 now appear primarily interested in cultural activities, such as visiting museums, and not in activities like alpine skiing, going to the spa, using health facilities, or going to discos/bars.
+
+!Note to this section!
+comparing the average shilouette scores indictes that the original k=6 solutiuon is best, even though in my opinion the plots look better with the manual solution.
+The scores are:
+
+k=6: 0.1460762
+
+k=7: 0.1336373
+
+k=6 + manual: 0.140716
+
+Let me know what you think
+
 \subsection{Australian Vacation Activities dataset}
@@
 Similar to the Austrian dataset, responses were binarized: a value of 1 indicates that the participant took part in the activity, while a value of 0 signifies they did not.
-
-\section{Examples}
 

### 2024-09-17 manuscript.tex
added reproducability section and reworked manipulation of austrian cluster solution

@@
 
-When clicking the "Save projections and subsets" button, a file browser is first launched, allowing users to specify the destination for the output files.
-These files include a PNG image of the current plots, CSV files containing the projection matrices for all displayed tours, and a mapping of the data points to their respective subsets.
+The "Save projections and subsets" button can be used to save the current state of the analysis.
+These states can be recovered by using pressing the "Load projections and subsets" button.
 Additionally, users can initiate new tours of various directly from the sidebar.
@@
 Since this process is relatively slow, full updates are only triggered during initialization or after significant changes, such as modifications to the set of active features.
+
+\subsection{Reproducibility}
+
+Ensuring the reproducibility of data analysis is a fundamental principle in scientific research.
+It allows others to verify the validity of the findings and is key to the integrity of the scientific process.
+Reproducibility not only builds trust in the research outcomes but also enables the scientific community to build upon existing work.
+When analyses can be replicated, it can be validated whether the conclusions drawn from the data are robust and not dependent on the specific conditions or idiosyncrasies of the original analyst.
+Moreover, reproducible research can serve as a foundational building block for subsequent studies, fostering incremental advancements in knowledge.
+
+One challenge in the context of interactive data analysis is that not all steps of the analysis are precisely documented in the form of code, especially when using graphical user interfaces (GUIs) where user-driven interactions might not leave a traceable history.
+This lack of documentation can hinder the ability of others to reproduce the analysis or to understand how specific results were obtained.
+To mitigate this challenge, it is essential to implement mechanisms that allow users to easily save and share intermediate snapshots of their analyses.
+
+One measure to combat this is to make saving intermediate snapshots of the analysis easy and accessable.
+Specifically, the "Save projections and subsets" button enables users to take snapshots of their analysis, including visual representations, selected data, and parameter settings.
+Upon pressing this button, a file browser is triggered, allowing users to specify the destination for saving these snapshots.
+Each save operation generates multiple files:
+
+\begin{itemize}
+    \item A \texttt{.png} file containing the currently displayed graphics,
+    \item \texttt{.csv} files that capture the feature and subset selection as well as projections of the tours displayed at the time of the snapshot,
+    \item two \texttt{.pkl} files that contain state variables of the GUI, allowing for complete recovery of the snapshot.
+\end{itemize}
+
+These files provide dual utility.
+First, they allow users to fully recover the state of the analysis within the GUI.
+This can be achieved either by using the "Load projections and subsets" button, or by launching a new GUI instance with the \texttt{load\_interactive\_tour()} function.
+The latter approach, using \texttt{load\_interactive\_tour()}, has the added flexibility of only requiring the original dataset and the directory containing the saved files.
+This function also allows users to modify display settings, such as adjusting the size of the interactive plots or changing the arrangement of the display grid.
+In contrast, when loading the saved state directly from within the GUI, it is crucial that the active session was initiated with the same dataset and plot objects that were present at the time of saving.
+This ensures that the analysis environment is accurately replicated.
+
+Second, the saved \texttt{.csv} files provide a way to inspect and further analyze the data outside of the original interface.
+This opens up opportunities for deeper analysis and extensions of the work.
+
+This level of interactivity and documentation is crucial for reproducibility, as it ensures that even exploratory, interactive data analysis can be retraced and validated by others.
+Ultimately, these features facilitate a reproducible workflow that balances the flexibility of interactive exploration with the rigor of reproducible research.
 
@@
 
-\subsubsection{Manual cluster improvement}
+\subsubsection{Manual cluster selection}
 
-We have seen in Figure \ref{fig:silhouette_comparison} that clusters 1 and 3 contain some data points that do not fit well into their respective clusters.
-Manual selection of data points can be employed to further improve the quality of the cluster solution.
-To do this, we first initialize an \texttt{interactive\_tour()} with the following components:
+There are several reasons why we might want to manually modify a clustering solution.
+One is to capture observations that do not fit well within their assigned clusters.
+Another reason is to explore specific features in more detail.
+The advantage of manual cluster selection is that it preserves most of the current clustering structure, allowing us to adjust specific parts of the solution without starting from scratch.
+This approach is particularly useful when we already have a cluster solution that reveals interesting patterns in the data.
+
+In Figure \ref{fig:silhouette_comparison}, we observed that clusters 1 and 3 contained data points that did not fit well into their respective clusters.
+To further investigate this, we can initialize an \texttt{interactive\_tour()} with the following components:
 
@@
 \begin{itemize}
-    \item A 2D tour with the linear discriminant analysis projection pursuit index,
-    \item A heatmap displaying the intra-cluster fraction,
+    \item A 2D tour using the linear discriminant analysis projection pursuit index,
+    \item A heatmap showing the intra-cluster fraction,
     \item A 1D tour with the linear discriminant analysis projection pursuit index, and
@@
 This setup produces the display shown in Figure \ref{fig:winter_cl7_init}.
-We can observe that clusters 1 and 3 overlap significantly in the features "sightseeing," "museums," and "going out in the evenings." By highlighting clusters 1 and 3, and adjusting the projection axes, we can confirm that there is indeed overlap between these clusters, as shown in Figure \ref{fig:winter_cl7_pre}.
+It can also be seen that both clusters 1 and 3 contain tourists that didn't go alpine skiing and the main difference between them is that tourists in cluster 3 enjoyed going to the spa and health facilities as well as going to the pool, while the ones in cluster 1 didn't.
+Other than that, the clusters were mostly similar.
+Now we might be interested in the subset of tourists that enjoy going to museums.
+Therefore, we can adjust the projection axes so that these axes are elongated and point into different directions.
+We can see that there is indeed overlap between clusters 1 and 3, as shown in Figure \ref{fig:winter_cl7_pre}.
+As a next step we can reassign the overlapping section to a new cluster - cluster 7.
+By selecting the checkbox for subset 7 and manually selecting the region of overlap, we can form a new cluster, which is visualized in Figure \ref{fig:winter_cl7_post}.
 
@@
 
-To improve the quality of the k-means solution, we can select the checkmark for subset 7 and manually select the region of overlap.
-The selected data points will then be shifted to a new cluster: cluster 7.
-The interface with the newly formed cluster can be seen in Figure \ref{fig:winter_cl7_post}.
-
 \begin{figure}[h!]
@@
 
-To assess the quality of the manually enhanced cluster solution, we can compare the silhouette plots of this new solution with those from a k-means solution where \(k = 7\).
-These silhouette plots are shown in Figure \ref{fig:silhouette_comparison_k7_man}.
-It is evident that the manually enhanced cluster solution better describes the data compared to the k-means solution with \(k = 7\).
+To evaluate whether the manually enhanced cluster solution reduces the number of poorly fitting observations, we can compare the silhouette plots of this new solution with those from a k-means solution where \(k = 7\).
+Figure \ref{fig:silhouette_comparison_k7_man} shows that the manually enhanced cluster solution better describes the data compared to the k-means solution with \(k = 7\).
 
@@
 
-Additionally, manually enhancing a cluster solution has the advantage of maintaining the base structure of the clustering, whereas clusters in a new k-means solution can be entirely different.
-Moreover, tourists in cluster 7 now appear primarily interested in cultural activities, such as visiting museums, and not in activities like alpine skiing, going to the spa, using health facilities, or going to discos/bars.
+We can observe slight behavioral differences between tourists in clusters 1 and 7.
+Tourists in cluster 7 all enjoyed both museums and sightseeing, whereas most tourists in cluster 1 engaged in sightseeing but showed no interest in museums.
+Instead, participants in cluster 1 exhibited a greater preference for hiking.
+Despite this, tourists in both clusters generally shared similar interests.
+This insight could be valuable for enhancing museum marketing strategies.
+While clusters 1 and 7 have overlapping interests, it appears that current marketing efforts may not effectively reach tourists in cluster 1.
+By increasing targeted marketing at hiking trails, popular excursion destinations, and shopping centers, it may be possible to attract more interest in museums from tourists in cluster 1.
 
@@
 
-k=6 + manual: 0.140716
+k=6 + manual: 0.1408945
 

### 2024-09-18 manuscript.tex
added overview figure and reworkt description of rework. changed the metrics section

@@
 
-The GUI is divided into two main sections: a sidebar on the left, which contains a comprehensive set of interactive controls to customize and interact with the displays, and the display area on the right.
-An overview of the GUI can be seen in figure x.
+The GUI is divided into two main sections: a sidebar on the left, which contains a comprehensive set of interactive controls, and the display area on the right, where the selected plots are shown (see Figure \ref{fig:GUI_overview}).
 
-At the top of the sidebar, users can select and deselect features via checkboxes, determining which features are displayed in the plots.
-Below this, each subset of the data has a corresponding checkbox, which is used to select the active subset.
-When a subset is selected, any data points marked in the plots will be moved to this active subset and be colored accordingly.
-The colored boxes next to the subset names indicate how the datapoints within each subset are colored.
-Users can adjust the transparency of each subset by clicking these colored boxes.
-This can be useful when one wants to highlight one or multiple subsets, for instance when wanting to compare to subsets in particular.
-The subset selection can be reset to its original state at any time by clicking the "Reset original selection" button.
+At the top of the sidebar, users can select and deselect features via checkboxes (Figure \ref{fig:GUI_overview}C), controlling which features are displayed in the plots.
+Below that, each subset has a checkbox to designate the active subset (Figure \ref{fig:GUI_overview}D).
+When data points are manually selected in the plots, they will be assigned to the active subset and colored accordingly.
+The colored boxes next to the subset names indicate the color of the data points.
+Clicking these boxes adjusts the transparency of the points, which is useful for highlighting and comparing subsets.
+Subsets can also be renamed using the textboxes.
+The "Reset original selection" button allows users to revert the subset selections to the initial state.
 
-At the top of the sidebar, users can select and deselect features via checkboxes, controlling which features are displayed in the plots.
-Below the feature selection, each subset has a checkbox that designates the active subset.
-When datapoints within the plots are manually selected, they will be moved to the active subset and colored accordingly.
-The colored boxes next to the subset names indicate the colors that the datapoints within each subset are colored.
-By clicking these colored boxes, users can adjust the transparency of the datapoints within the corrisponding subet, which is particularly useful for highlighting and comparing specific subsets.
-The initial subset selection (in case one was provided) can be restored at any time by clicking the "Reset original selection" button.
+The frame selection interface (Figure \ref{fig:GUI_overview}E) displayes the frames of the currently shown tours and enables users to jump between frames of the tour objects.
+Below this, users can save and load projections and subsets using the respective buttons (Figure \ref{fig:GUI_overview}F).
+New tours can be started using the current settings via the interface at the bottom (Figure \ref{fig:GUI_overview}G), and users can select different metrics for some plot types e.g.
+heatmaps (Figure \ref{fig:GUI_overview}H).
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{GUI_overview.png}
+    \caption{Overview of the GUI, consisting of the sidebar (left) and display area (right).
+A: Sidebar.
+B: Display area.
+C: Feature selection checkboxes.
+D: Subset selection with color indicators.
+E: Frame selection interface.
+F: Save and load buttons.
+G: Interface for starting new tours.
+H: Metric selection interface.}
+    \label{fig:GUI_overview}
+\end{figure}
 
@@
 An effective and intuitive way to perform feature selection is by using the heatmap display within pytourr.
-The heatmap supports three different display metrics for feature selection:
+The heatmap supports three different display metrics \(m\) for feature selection:
 
 \[
-\text{Total Fraction:} \quad \frac{\displaystyle\sum_{i = 1}^{k} \sum_{j = 1}^{p} c_{ij}}{n}
+\text{Total Fraction: } m_{ij} = \frac{c_{ij}}{n}
 \]
@@
 \[
-\text{Intra Cluster Fraction:} \quad \sum_{i = 1}^{k} \sum_{j = 1}^{p} \frac{c_{ij}}{c_{i}}
+\text{Intra Cluster Fraction: }  m_{ij} = \frac{c_{ij}}{n_{i}} 
 \]
-The \textbf{Intra Cluster Fraction} displays the contribution of each feature within a specific cluster, providing insight into how much a feature contributes to the cluster compared to the overall cluster size \(c_i\).
+The \textbf{Intra Cluster Fraction} shows each feature's contribution to a specific cluster, relative to the cluster's size \(n_i\).
 
 \[
-\text{Intra Feature Fraction:} \quad \sum_{j = 1}^{p} \sum_{i = 1}^{k} \frac{c_{ij}}{c_{j}}
+\text{Intra Feature Fraction: } m_{ij} = \frac{c_{ij}}{c_{j}} 
 \]
-The \textbf{Intra Feature Fraction} highlights how a feature contributes across different clusters relative to the overall importance of that feature \(c_j\).
+The \textbf{Intra Feature Fraction} highlights how a feature contributes across different clusters relative to the overall count of that feature \(c_j\).
 

### 2024-09-19 manuscript.tex
introduction drafted

@@
 %% \AND can be used instead of \And and starts a new line
-\title{Interactively Visualizing Clustering in Multivariate Data Using pytourr}
+\title{Interactively Visualizing Multivariate Market Segmentation Using pytourr}
 
@@
 \Plainauthor{Matthias Medl, Dianne Cook, Ursula Laa} %% comma-separated
-\Plaintitle{Interactively Visualizing Clustering in Multivariate Data Using pytourr} %% without formatting
+\Plaintitle{Interactively Visualizing Multivariate Market Segmentation Using pytourr} %% without formatting
 \Shorttitle{pytourr} %% a short title (if necessary)
@@
 
-\begin{itemize} \itemsep 0in
-\item Background to clustering analysis
-\item Visualising clusters, including tours
-\item Difference between clustering and partitioning
-\item How do you know whether the results are useful
-\end{itemize}
+Clustering algorithms are often used to find a smaller number of observations (the cluster means) that adequately summarize a much larger number of observations.
+For market segmentation, clustering can allow partitioning of observations into a small number of groups, by incorporating associations between the variables.
+Market segmentation supports targeted approaches to different groups of customers based on common traits.
+It provides a data-driven solution to partitioning customer data.
+\cite{leisch2018market} provides an extensive overview of using clustering for market segmentation.
+
+A difference between clustering analysis and partitioning is typically the nature of the data.
+With cluster analysis, we usually envision data that contains separated clusters, and a successful clustering result is one that divides the data based on these gaps.
+With partitioning, it is usual that there are no gaps in the data, but it is still useful to partition the data.
+Figure \ref{kmeans-partition} illustrates how the $k$-means algorithm would partition a 2D data set into four groups depending on the correlation between the two variables.
+When the correlation is high, the partitioning will be along the combination of variables that produces the highest variance.
+With lower correlation, it will segment the bottom and top, and divide the middle into two parts in the opposite direction.
+When there is no association, the partitioning is radial like a windmill.
+
+\begin{figure*}[h]
+\centerline{\includegraphics[width=1\textwidth]{images/intro1.png}}
+\caption{Examples of how $k$-means partitions 2D data with different association structure. If the correlation is high (c), the partitioning happens along the primary direction of the association.}
+\label{kmeans-partition}
+\end{figure*}
+
+One can see that the data is perfectly divided into four parts.
+However, if one were to plot the two variables individually, this would not be obvious.
+Figure \ref{kmeans-histogram} shows histograms of the two variables, V1, V2, with the colour matching the four partitions.
+From the histograms, we can see some differences in the partitions for the four different association structures, but they are all overlapping.
+The distinct border between the partitions can only be seen from the scatterplots of both variables.
+
+\begin{figure*}[ht]
+\centerline{\includegraphics[width=1\textwidth]{images/intro2.png}}
+\caption{The four $k$-means partitions plotted as histograms of the individual variables.
+While differences between groups can be seen, the clear separation cannot.
+This is important for understanding why high-dimensional visualisation methods are useful for summarising partitioning results.}
+\label{kmeans-histogram}
+\end{figure*}
+
+When there are more than two variables, histograms of the individual variables are commonly used to display the partitioning results.
+This means that the analyst likely cannot understand how the partitioning divides the data.
+All that they can observe is roughly how the individual variables relate to the partition, which is useful but inadequate.
+Here is where using tour methods to view high dimensions can be helpful.
+A tour (XXX general tour and recent tour) can be used to show scatterplots of combinations of variables and thus provide views like that in Figure \ref{kmeans-partition} where distinct differences between partitions can be observed.
+High dimensions are still tricky, and a combination of animations of the linear combinations, and interactive control (XXX REF radial/manual tour) over the combinations is important.
+A scatterplot of a combination of variables can be considered to be a projection of the data, and thus like a shadow of a 3D object, some aspects of the data (object) can be obscured.
+Using slices of the projected data (XXX REF) can be a useful addition to projections.
+This paper illustrates how to do this to better understand partitioning results for multivariate data.
+
+%\begin{itemize} \itemsep 0in
+%\item Background to clustering analysis
+%\item Visualising clusters, including tours
+%\item Difference between clustering and partitioning
+%\item How do you know whether the results are useful
+%\end{itemize}
+
+This paper is organised as follows XXX The methods are illustrated using Austrian and Australian tourism data provided in \cite{leisch2018market}.
 

### 2024-09-19 scripts/intro.R

@@ -1,0 +1,129 @@
+# Example in introduction
+library(tidyverse)
+library(mvtnorm)
+library(patchwork)
+library(colorspace)
+
+theme_set(theme_bw(base_size = 14) +
+            theme(
+              aspect.ratio = 1,
+              plot.background = element_rect(fill = 'transparent', colour = NA),
+              plot.title.position = "plot",
+              plot.title = element_text(size = 24),
+              panel.background = element_rect(fill = 'transparent', colour = NA),
+              legend.background = element_rect(fill = 'transparent', colour = NA),
+              legend.key = element_rect(fill = 'transparent', colour = NA)
+            )
+)
+
+f_std <- function(x) {(x-min(x))/(max(x)-min(x))}
+set.seed(914)
+blob1 <- rmvnorm(n=155, mean=c(0,0), 
+                 sigma=matrix(c(1, 0, 0, 1), 
+                              ncol=2, byrow=TRUE)) |> 
+  as_tibble() |>
+  mutate_all(f_std)
+blob2 <- rmvnorm(n=155, mean=c(0,0), 
+                 sigma=matrix(c(1, 0.6, 0.6, 1), 
+                              ncol=2, byrow=TRUE)) |> 
+  as_tibble() |>
+  mutate_all(f_std)
+blob3 <- rmvnorm(n=155, mean=c(0,0), 
+                 sigma=matrix(c(1, 0.9, 0.9, 1), 
+                              ncol=2, byrow=TRUE)) |> 
+  as_tibble() |>
+  mutate_all(f_std)
+b1 <- ggplot(blob1, aes(V1, V2)) + 
+  geom_point() +
+  theme(axis.text = element_blank(),
+        axis.title = element_blank())
+b2 <- ggplot(blob2, aes(V1, V2)) + 
+  geom_point() +
+  theme(axis.text = element_blank(),
+        axis.title = element_blank())
+b3 <- ggplot(blob3, aes(V1, V2)) + 
+  geom_point() +
+  theme(axis.text = element_blank(),
+        axis.title = element_blank())
+b1 + b2 + b3 + plot_layout(ncol=3)
+
+
+set.seed(855)
+b1_km <- kmeans(blob1, 4)
+b2_km <- kmeans(blob2, 4)
+b3_km <- kmeans(blob3, 4)
+blob1_cl <- blob1 |>
[…]

### 2024-09-20 manuscript.tex
Some restructuring of the sections

@@
 
-\section{Related Work}
+\section{Interactive interface for partitioning}
 
-Previous attempts have been made to bring interactivtiy to the tourr package.
-The galahr package \citep{galahr} was an attempt to so by means of the R packages shiny \citep{shiny} and plotly \citep{plotly}, but ultimately suffered from performance issues when the size of the investigated dataset became too large (n>10 000).
-To overcome these issues, the detourr package \citep{detourr}, which precomputes multiple projections in R and subsequently visualizes them in an interactive display using Javascript, has been developed.
-While detourr showed better performance, the interactivtiy is still limited as only precomputed projections can be displayed and there is no option to show multiple linked displays simultaneously.
-The mmtour package \citep{laa2023new} is a \proglang{Mathamatica} implementation allowing for interaction with manual and slice tours.
+The aim of this work is to build an interface that allows for an interactive exploration of partitioning solutions of multivariate data. The interface should enable
 
-\section{The pytourr package}
+\begin{itemize}
+\item visualizing a clustering solution with different tour methods, including manual changes to the viewing direction (manual tour)
+\item brushing of (groups of) points for manual refinement of the cluster solution (spin-and-brush capabilities)
+\item additional linked displays for comprehensive exploration of the solution
+\item ability to update the displays based on user selections
+\end{itemize}
 
-\subsection{Overview}
+While tour animations are best obtained within R using the tourr package~\citep{tourr}, it does not enable the interactivity required for example for a manual tour~\citep{laa2023new}.
+Interactive graphics are available when using Javascript, as implemented in the detourr package \citep{detourr}.
+This allows to replay a recorded tour path with interactive graphics, and can also be linked with additional displays, but lacks capabilities for manual tours.
 
-The primary objective of this work is to develop an R package that integrates high-performance interactivity through Python packages, such as \texttt{TKinter} \citep{lundh1999introduction}, \texttt{CustomTKinter} \citep{schimansky24}, and \texttt{matplotlib} \citep{Hunter:2007}, with the tourr package \citep{tourr}.
-This integration is achieved via \texttt{reticulate} \citep{reticulate}, a framework that facilitates seamless interoperability between Python and R.
+To integrate user interactions with the capabilities of the tourr package an active communitcation with the interface is required.
+For example, we may wish to explore the local neighbourhood of a projection selected by the user with a local tour animation provided by tourr, or we may want to optimize a guided tour path using groups identified via brushing.
+Our solution is using Python for high-performance interactivity, through the packages \texttt{TKinter} \citep{lundh1999introduction}, \texttt{CustomTKinter} \citep{schimansky24}, and \texttt{matplotlib} \citep{Hunter:2007}, with integration to the tourr package \citep{tourr} via \texttt{reticulate} \citep{reticulate}, a framework that facilitates seamless interoperability between Python and R.
+\textcolor{red}{also rpy2?}
 
-At the core of the \texttt{pytourr} package is a graphical user interface (GUI) designed to display multiple linked visualizations concurrently.
+%At the core of the \texttt{pytourr} package is a graphical user interface (GUI) designed to display multiple linked visualizations concurrently.
 Users can interact with one of the plots, for example, by subselecting specific data points, and the other linked plots will dynamically update to reflect these modifications.
@@
 
-In addition to this core functionality, \texttt{pytourr} offers a variety of interactive plot types, providing users with the flexibility to visualize their data from multiple perspectives.
-The ability to navigate through various projections of the displayed tours directly within the GUI enables users to explore different aspects of the dataset with ease.
+The interface was implemented in \texttt{pytourr} and offers a variety of linked interactive plot types, providing users with the flexibility to visualize their data from multiple perspectives.
+The ability to navigate through various projections of the displayed tours directly within the GUI enables users to explore different aspects of the dataset.
 Furthermore, users can initiate new tours directly from the interface.
-The GUI also supports interactive feature selection, allowing users to specify which subset of features should be visualized in the plots.
+The GUI also supports interactive variable selection, allowing users to specify which subset of variables should be visualized in the plots.
 Once users have identified interesting views or settings, \texttt{pytourr} allows them to save the displayed projections, subsets, and plots.
@@
 Since this process is relatively slow, full updates are only triggered during initialization or after significant changes, such as modifications to the set of active features.
+
+\section{Workflow with the pytourr package}
+
+Before launching the pytourr interface the user should have performed a partitioning of their choice, and provide the initial clustering solution used by the interface.
+We launch the interface to explore and potentially refine this solution.
+
+\subsection{Variable selection}
+
+A first step may look into what variables are important for the cluster solution.
+The interface provides different capabilites that support variable selection.
+First, a 2D tour view can be used to understand the sensitivity of the grouping to individual varaibles, in particular when using the manual tour to remove one variable at a time.
+This is in particular useful when the starting view was obtained through a projection pursuit guided tour REF that optimized for the separation between the labeled groups REF.
+
+In our applications we are considering market segmentation analysis on binary variables.
+For this setting we have developed a heatmap display that we have found to be particularly useful in variable selection.
+The idea is to understand differences in the binary ratings between clusters, where we use different normalizations to better explore the data.
+For \(c_{ij}\) the count (number of true values in the binary variable) in the \(i\)-th cluster and \(j\)-th variable, we define
+
+\[
+\text{Overall count fraction}\ f_{ij} = \frac{c_{ij}}{n}
+\]
+with normalization to the total number of observations $n$ for a general overview,
+
+\[
+\text{Cluster count fraction}\  f_{ij} = \frac{c_{ij}}{n_{i}}
+\]
+showing the counts relative to the cluster's size \(n_i\), to see the fraction of observations in cluster $i$ for which variable $j$ was true, and
+
+\[
+\text{Variable count fraction}\ f_{ij} = \frac{c_{ij}}{c_{j}}
+\]
+to see how the true values for variable $j$ are distributed accross the different clusters relative, with normalization to the overall count of that feature \(c_j = \sum_i c_{ij}\).
+
+\textcolor{red}{different symbol for each? indicat on the index value somehow?}
+
+\subsection{Subset selection}
+
+The spin-and-brush approach suggests to cluster data manually when using a tour: we run a tour animation, stop when we see a group of points that are different from the rest of the distribution, brush them, and then continue.
+Different projections will enable the separation of different groups, and for well-separated clusters we will be able to recover a full cluster solutions in this manner.
+
+A similar approach can be used to refine a partitioning solution.
+This is in particular useful to integrate prior knowledge or business interests in a given cluster solution.
+In the interface we can keep the provided clustering, but separate out new subsets via manual selection, for example after we found a group of particular interest via a manual tour.
 

### 2024-09-21 manuscript.tex
minor changes to section 2. Addition of analysis of australian dataset. Rework of some figures and saves

@@
 
-At the top of the sidebar, users can select and deselect features via checkboxes (Figure \ref{fig:GUI_overview}C), controlling which features are displayed in the plots.
-Below that, each subset has a checkbox to designate the active subset (Figure \ref{fig:GUI_overview}D).
+At the top of the sidebar, users can select and deselect features using checkboxes (Figure \ref{fig:GUI_overview}C), controlling which features are displayed in the plots.
+Below this, each subset has its own checkbox to designate the active subset (Figure \ref{fig:GUI_overview}D).
 When data points are manually selected in the plots, they will be assigned to the active subset and colored accordingly.
-The colored boxes next to the subset names indicate the color of the data points.
-Clicking these boxes adjusts the transparency of the points, which is useful for highlighting and comparing subsets.
-Subsets can also be renamed using the textboxes.
-The "Reset original selection" button allows users to revert the subset selections to the initial state.
+For scatterplots, data points can be selected by encircling them directly on the plot while holding down the left mouse button.
+For barplots, clicking on a specific bar selects the data represented by that bar.
+The colored boxes next to the subset names indicate the assigned colors for the data points.
+Clicking on these boxes adjusts the transparency of the points, which is helpful for highlighting and comparing subsets.
+Subsets can also be renamed using the provided text boxes.
+The "Reset original selection" button allows users to revert the subset selections to their initial state.
 
@@
 After a selection is made, the selector class updates the centrally stored subselection attribute and ensures all other displays reflect these changes.
-The manual projection manipulation classes construct the arrows representing the projection axes in the displays and manage the manual adjustment of projections.
+The manual projection manipulation classes construct the arrows representing the projection axes in the displays and manage the manual adjustment of projections, and thus enable manual tours.
 Users can right-click and drag the arrowheads to modify the projections, after which the class orthonormalizes the projection axes and updates both the projection and the transformed data accordingly.
@@
 Additionally, interacting with the pytourr GUI becomes cumbersome when handling more than \textasciitilde
- 15 features, making it necessary to reduce the dimensionality of the dataset.
-An effective and intuitive way to perform feature selection is by using the heatmap display within pytourr.
-The heatmap supports three different display metrics \(m\) for feature selection:
-
-\[
-\text{Total Fraction: } m_{ij} = \frac{c_{ij}}{n}
-\]
-The \textbf{Total Fraction} shows the counts \(c_{ij}\) of the \(i\)-th cluster and \(j\)-th feature, normalized by the total number of observations \(n\).
-
-\[
-\text{Intra Cluster Fraction: }  m_{ij} = \frac{c_{ij}}{n_{i}} 
-\]
-The \textbf{Intra Cluster Fraction} shows each feature's contribution to a specific cluster, relative to the cluster's size \(n_i\).
-
-\[
-\text{Intra Feature Fraction: } m_{ij} = \frac{c_{ij}}{c_{j}} 
-\]
-The \textbf{Intra Feature Fraction} highlights how a feature contributes across different clusters relative to the overall count of that feature \(c_j\).
+ 15 features, making it necessary to reduce the dimensionality of the dataset. An effective and intuitive way to perform feature selection is by using the heatmap display within pytourr.
 
@@
 We can also see in Figure \ref{fig:silhouette_comparison}B that cluster 3 is of comparatively low quality.
+It is important to note that the silhouette scores were generally quite low, reflecting the lack of clearly separable clusters in the data.
+Consequently, there is no objectively correct or optimal solution.
+The primary aim of this analysis, therefore, is to gain insights into the data rather than to identify the optimal clustering configuration.
 
@@
 
-To evaluate whether the manually enhanced cluster solution reduces the number of poorly fitting observations, we can compare the silhouette plots of this new solution with those from a k-means solution where \(k = 7\).
-Figure \ref{fig:silhouette_comparison_k7_man} shows that the manually enhanced cluster solution better describes the data compared to the k-means solution with \(k = 7\).
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=1\textwidth]{silhouette_comparison_k7_man.png}
-    \caption{A: Silhouette plot of the k-means solution with \(k=7\). B: Silhouette plot of the manually enhanced k-means solution with \(k=6\).}
-    \label{fig:silhouette_comparison_k7_man}
-\end{figure}
-
 We can observe slight behavioral differences between tourists in clusters 1 and 7.
@@
 
-!Note to this section!
-comparing the average shilouette scores indictes that the original k=6 solutiuon is best, even though in my opinion the plots look better with the manual solution.
-The scores are:
-
-k=6: 0.1460762
-
-k=7: 0.1336373
-
-k=6 + manual: 0.1408945
-
-Let me know what you think
-
 \subsection{Australian Vacation Activities dataset}
@@
 Similar to the Austrian dataset, responses were binarized: a value of 1 indicates that the participant took part in the activity, while a value of 0 signifies they did not.
+Suvreys where participants claimed they partook in more than 40 activities or no activity at all were removed as they are considered faulty.
+
+\subsubsection{Feature selection}
+
+At first, hierarchical clustering using Ward's method \citep{ward2} and the Jaccard index was applied to the features.
+The resulting dendrogram is shown in Figure \ref{fig:aus_feature _clustering}.
+Based on this clustering, \( k = 15 \) clusters were identified, and generally, only one representative feature from each cluster was selected for further analysis.
+Clusters containing unpopular activities, such as "Adventure", which only had 42 participants, were discarded.
+The cluster containing popular features like "Beach," "Swimming," "ScenicWalks," "Markets," "Sightseeing," "Friends," "Pubs," "BBQ," "Shopping," "Eating," "EatingHigh," "Movies," and "Relaxing" was treated differently.
+Multiple features from this cluster were retained to preserve as much information as possible.
+After feature selection, the observations were reclustered using k-means with \( k = 15 \).
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{aus_feature_clustering.png}
+    \caption{Dendogram of the features of the Australian Vacation Activities dataset using Ward's method with the Jaccard index.}
+    \label{fig:aus_feature_clustering}
+\end{figure}
+
+\subsubsection{Manual cluster selection}
+
+In the heatmap displaying the intra-cluster fraction shown in Figure \ref{fig:aus_heatmapg}, we observe that subsets 1, 2, and 6 tend to prefer travelling alone.
+As a tourist agency, we might be interested in targeting solo travellers more effectively.
+However, subsets 2 and 6 also include individuals who enjoy spending time with their friends.
+To further explore the dataset, we can launch an \texttt{interactive\_tour()} with the same configuration as in the previous example.
+To achieve better separation between the subsets, we can skip to the last frames of a 2 dimensional tour optimized for the LDA index.
+
+Since we are particularly interested in tourists who did not spend time with friends, we can extend the projection axis "Friends" outward to separate the data based on that feature.
+Datapoints in the opposite direction of the "Friends" axes are the solo travellers we are interested in, which can be seen in Figure \ref{fig:aus_preselection} on the left side of the top left plot.
+Subsequently, we can pull all other projection axes in one direction to separate data points based on their overall activity level.
+Tourists that fall in the direction of the axes generally engaged in more activities compared to those in the opposite direction of the projection axes.
+This separation is evident, as observations in subset 6 are located opposite the axes, and the heatmap shows that they did not engage in many activities.
+Similarly we can also see that subset 2, which contains quite active tourists, is shifted towards the direction of the axes.
+
+Using this logic, we can segment the data points on the left into three segments: active tourists (more upward), medium active tourists (center), and largely inactive tourists (bottom).
+The result of this segmentation can be seen in Figure \ref{fig:aus_selection}.
+Analyzing the heatmap, we can identify several interesting patterns.
+
+By comparing subset 2, which contains active tourists who spent time with their friends, with subset 7, the active tourists who traveled alone, we notice that subset 7 showed less interest in visiting the casino, theatre, and chartering a boat.
+
+We also observe distinctive differences in the interests of the subsets that traveled alone.
+Subset 7 was interested in relaxing, shopping, sightseeing, wildlife, going to the beach, visiting pubs, and exploring farms.
+Given that subset 7 showed a notable interest in going to pubs, we can assume they want to meet new people.
+This insight can be utilized in a marketing campaign that bundles the activities subset 7 is interested in, creating a package tailored for solo travelers.
+This way, they can meet other solo travelers while engaging in activities they enjoy.
+
+Although subset 8 was generally less active, almost everyone still engaged in sightseeing.
+For them, the focus was on sightseeing, relaxing, shopping, and going to the beach, with much less interest in other activities.
+It can be assumed that they value their time alone.
+These tourists could potentially be targeted more effectively by offering sightseeing options with minimal interaction, such as using a phone app to provide information about interesting locations, rather than relying on a tour guide.
+
+Finally, since subset 9 is notably inactive, one might infer that they prefer spending much of their time in their accommodation.
+Consequently, these individuals might be most interested in accommodations that offer well-equipped, comfortable living spaces with amenities that cater to relaxation and leisure.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{aus_heatmap.png}
+    \caption{Heatmap of the intra cluster fraction of the k-means solution with \( k = 6 \) and the reduced feature subset.}
+    \label{fig:aus_heatmapg}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{aus_preselection.png}
+    \caption{Separation of the data based on the feature "Friends" and the general activity.
+Top left: 2D tour.
+Top right: 1D tour.
+Bottom left: Mosaic plot.
+Bottom right: Heatmap with the intra cluster fraction.}
+    \label{fig:aus_preselection}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1\textwidth]{aus_selection.png}
+    \caption{Separation of the data based on the feature "Friends" and the general activity after reselection of subsets. 
+    Tourists in subsets 7 (pink), 8 (grey) and 9 (gold) all responded thay didn't engage in activities with their friends.
+Subset 7 was the most active, subset 8 medium active and subset 9 largely inactive.
+Top left: 2D tour.
+Top right: 1D tour.
+Bottom left: Mosaic plot.
+Bottom right: Heatmap with the intra cluster fraction.}
+    \label{fig:aus_selection}
+\end{figure}
 

### 2024-09-23 manuscript.tex
added results and discussion. Modified the feature clustering figure

@@
 
-\subsubsection{Manual cluster selection}
+\subsubsection{Redefing cluster assignments - learning more about museum goers}
 
@@
 
-\subsubsection{Manual cluster selection}
+\subsubsection{Segmentation of solo travellers}
 
@@
 
+One might question the necessity of manual exploratory data analysis, considering it is inherently subjective and relies heavily on intuition.
+However, in situations like those presented here, where there is no clearly defined or optimal clustering solution or feature selection, manual exploration becomes indispensable.
+While it is possible to optimize clustering metrics to improve separation between clusters, this alone may not yield conclusions that are useful for practical applications.
+The lack of clear boundaries and the overlapping nature of clusters in the datasets underscore the limitations of purely automated methods in capturing the complexity and nuance of real-world data.
+
+In such cases, manual exploration allows analysts to interweave expert knowledge, intuition, and specific objectives with the initial clustering solution, which serves as the backbone of the analysis.
+This approach is particularly valuable when no single solution can be deemed “correct.” For instance, in the Austrian dataset, leveraging the interactive GUI for feature selection enabled us to isolate the most informative activities and gain deeper insights into the preferences of tourists who might be interested in visiting museums.
+This understanding facilitated targeted recommendations for increasing museum attendance by focusing on tourists frequenting hiking trails, excursion spots, and shopping centers.
+Such insights are challenging to extract through automated optimization alone.
+
+Similarly, in the Australian dataset, manual exploration allowed us to refine the understanding of solo travelers by dividing them into three distinct subsegments: highly active, moderately active, and largely inactive tourists.
+This nuanced segmentation, derived from a blend of automated clustering and manual adjustments, offers a deeper understanding of the varied needs and behaviors within this group, enabling more effective marketing strategies.
+
+While some plots supported by the pytourr package are specifically designed for analyzing binary survey data, it is important to emphasize that its capabilities extend far beyond this data type.
+The versatility of tours has been demonstrated across various applications in the past, showcasing their effectiveness in exploring complex, high-dimensional datasets.
+The interactive GUI enables seamless exploration of both one- and two-dimensional tours, regardless of the data being analyzed, providing a powerful tool for uncovering patterns and insights across diverse domains.
+
 \section{Conclusion}
+
+Ultimately, manual data exploration serves as a crucial complement to automated methods, providing the flexibility to incorporate context, expert judgment, and specific analytical goals.
+This approach enables analysts to refine initial results and adapt them to the complexities of real-world scenarios, leading to more nuanced interpretations and actionable insights.
+The pytourr package offers a user-friendly and powerful tool for conducting such analyses, bridging the gap between automated clustering and interactive exploration.
+
+By integrating interactive visualization capabilities, the pytourr package empowers users to dynamically engage with their data, making it possible to uncover subtle patterns and relationships that might otherwise remain hidden.
+This flexibility is particularly valuable in complex datasets where traditional methods may fall short.
+Additionally, the package's broad applicability across various data types and analytical contexts highlights its potential as a versatile tool for a wide range of research and practical applications.
+
+Looking forward, expanding the interactive features and supporting additional visualization and data types could further enhance the package’s utility, making it an even more powerful tool for exploratory data analysis.
+Future developments could also include integrating advanced machine learning techniques, such as enabling users to generate new cluster solutions directly within the GUI.
+Another promising direction would be the implementation of sliced tours, which would enhance the package's capability to analyze concave and nonlinear structures.
+
+In summary, pytourr represents a significant advancement in the toolkit of data analysts, offering a novel way to balance automated analysis with human intuition and domain expertise, thereby facilitating a deeper and more comprehensive understanding of complex datasets.
 

### 2024-09-25 manuscript.tex
modified conclusion

@@
 
-Ultimately, manual data exploration serves as a crucial complement to automated methods, providing the flexibility to incorporate context, expert judgment, and specific analytical goals.
+Ultimately, manual data exploration serves as a useful complement to automated methods, providing the flexibility to incorporate context, expert judgment, and specific analytical goals.
 This approach enables analysts to refine initial results and adapt them to the complexities of real-world scenarios, leading to more nuanced interpretations and actionable insights.
-The pytourr package offers a user-friendly and powerful tool for conducting such analyses, bridging the gap between automated clustering and interactive exploration.
+The pytourr package offers an organised and responsive interactive tool for conducting such analyses, bridging the gap between automated clustering and exploration.
 
 By integrating interactive visualization capabilities, the pytourr package empowers users to dynamically engage with their data, making it possible to uncover subtle patterns and relationships that might otherwise remain hidden.
-This flexibility is particularly valuable in complex datasets where traditional methods may fall short.
-Additionally, the package's broad applicability across various data types and analytical contexts highlights its potential as a versatile tool for a wide range of research and practical applications.
+This is especially valuable in tackling complex datasets with the mindset that an automated solution needs to be validated.
+The package has broad applicability across various data types and analytical contexts where clustering is used.
+The flexibility of setting up the GUI elements from the command line allow it to be tailored for different applications.
 
-Looking forward, expanding the interactive features and supporting additional visualization and data types could further enhance the package’s utility, making it an even more powerful tool for exploratory data analysis.
-Future developments could also include integrating advanced machine learning techniques, such as enabling users to generate new cluster solutions directly within the GUI.
-Another promising direction would be the implementation of sliced tours, which would enhance the package's capability to analyze concave and nonlinear structures.
+Future developments in the software might include expanding the range of interactive features and providing additional visualization methods.
+Because it is built on R and python, it would also be possible to integrate more closely with the machine learning algorithms, such as generating new cluster solutions directly from the GUI.
+For more complex shapes, containing concavities, or non-linear boundaries, implementation of sliced tours, would be recommended.
 

### 2024-09-25 manuscript.tex
caption for fig 4 and refs for intro

@@
 Here is where using tour methods to view high dimensions can be helpful.
-A tour (XXX general tour and recent tour) can be used to show scatterplots of combinations of variables and thus provide views like that in Figure \ref{kmeans-partition} where distinct differences between partitions can be observed.
-High dimensions are still tricky, and a combination of animations of the linear combinations, and interactive control (XXX REF radial/manual tour) over the combinations is important.
+A tour (\cite{Asimov1985-xr}, and see \cite{lee2022}) can be used to show scatterplots of combinations of variables and thus provide views like that in Figure \ref{kmeans-partition} where distinct differences between partitions can be observed.
+High dimensions are still tricky, and a combination of animations of the linear combinations, and interactive control \citep{cook_manual_1997, laa2023new} over the combinations is important.
 A scatterplot of a combination of variables can be considered to be a projection of the data, and thus like a shadow of a 3D object, some aspects of the data (object) can be obscured.
-Using slices of the projected data (XXX REF) can be a useful addition to projections.
+Using slices of the projected data \citep{Laa2020} can be a useful addition to projections.
 This paper illustrates how to do this to better understand partitioning results for multivariate data.
@@
 
-This paper is organised as follows XXX The methods are illustrated using Austrian and Australian tourism data provided in \cite{leisch2018market}.
+This paper is organised as follows.
+Section \ref{interface} describes the software interface.
+The user workflow is explained in \ref{workflow}.
+The methods are illustrated in Section \ref{applications} using Austrian and Australian tourism data provided in \cite{leisch2018market}.
+Sections \ref{discussion} and \ref{conclusion} discuss the limitations and potential future developments.
 
-\section{Interactive interface for partitioning}
+\section{Interactive interface for partitioning}~\label{interface}
 
@@
 While tour animations are best obtained within R using the tourr package~\citep{tourr}, it does not enable the interactivity required for example for a manual tour~\citep{laa2023new}.
-Interactive graphics are available when using Javascript, as implemented in the detourr package \citep{detourr}.
+Interactive graphics are available when using Javascript, as implemented in the detourr package \citep{RJ-2023-052}.
 This allows to replay a recorded tour path with interactive graphics, and can also be linked with additional displays, but lacks capabilities for manual tours.
@@
 The interface was implemented in \texttt{pytourr} and offers a variety of linked interactive plot types, providing users with the flexibility to visualize their data from multiple perspectives.
-The ability to navigate through various projections of the displayed tours directly within the GUI enables users to explore different aspects of the dataset.
+The ability to navigate through various projections of the displayed tours directly within a graphical user interface (GUI) enables users to explore different aspects of the dataset.
 Furthermore, users can initiate new tours directly from the interface.
@@
 
-\subsection{The graphical user interface}
+\subsection{Overview of the graphical user interface (GUI)}
 
@@
     \includegraphics[width=1\textwidth]{GUI_overview.png}
-    \caption{Overview of the GUI, consisting of the sidebar (left) and display area (right).
-A: Sidebar.
-B: Display area.
-C: Feature selection checkboxes.
-D: Subset selection with color indicators.
-E: Frame selection interface.
-F: Save and load buttons.
-G: Interface for starting new tours.
-H: Metric selection interface.}
+    \caption{Overview of the GUI.
+A: Sidebar controls; B: Display area;  C: Feature selection checkboxes; D: Subset selection with color indicators; E: Frame selection interface; F: Save and load buttons; G: Interface for starting new tours; H: Metric selection interface.}
     \label{fig:GUI_overview}
@@
 
-\section{Workflow with the pytourr package}
+\section{Workflow with the pytourr package}~\label{workflow}
 
@@
 
-\section{Applications}
+\section{Applications}~\label{applications}
 
@@
     \includegraphics[width=1\textwidth]{winteractiv_heatmap.png}
-    \caption{A: heatmap displaying the intra cluster fraction. B: heatmap displaying the intra feature fraction.}
+    \caption{Traditional overview of clusters.
+Color represents  (A) the intra-cluster fraction, and (B) the intra-feature fraction, with lighter indicating higher values.
+From A, we can see that some activities dominate most clusters, e.g.
+relaxing and going for walks.
+Some activities are unique to clusters, e.g.
+skiing is a primary activity for clusters 1, 3, 5, 6.
+From B, we can see which activities dominate each cluster, e.g.
+cluster 4 tourists like going to a spa and using health activities.}
     \label{fig:winteractiv_heatmap}
@@
 
-In Figure \ref{fig:winteractiv_heatmap}A, we can observe the general interests of tourists within each cluster.
+\textcolor{red}{It might be better to standardise these intra fractions.
+For A, one should divide by cluster count, and then compute activity fraction.
+For B, one should divide by activity count and then compute cluster fraction.} In Figure \ref{fig:winteractiv_heatmap}A, we can observe the general interests of tourists within each cluster.
 For instance, tourists in clusters 1, 3, 5, and 6 predominantly engaged in alpine skiing, while those in clusters 2 and 4 did not.
@@
 
-\section{Discussion}
+\section{Discussion}~\label{discussion}
 
@@
 
-\section{Conclusion}
+\section{Conclusion}~\label{conclusion}
 
@@
 Because it is built on R and python, it would also be possible to integrate more closely with the machine learning algorithms, such as generating new cluster solutions directly from the GUI.
-For more complex shapes, containing concavities, or non-linear boundaries, implementation of sliced tours, would be recommended.
+For more complex shapes, containing concavities, or non-linear boundaries, implementation of sliced tours~\citep{Laa2020}, would be recommended.
 

### 2024-09-26 manuscript.tex
References added

@@
 First, a 2D tour view can be used to understand the sensitivity of the grouping to individual varaibles, in particular when using the manual tour to remove one variable at a time.
-This is in particular useful when the starting view was obtained through a projection pursuit guided tour REF that optimized for the separation between the labeled groups REF.
+This is in particular useful when the starting view was obtained through a projection pursuit guided tour~\cite{CBCH94} that optimized for the separation between the labeled groups~\cite{lckl2005}.
 

### 2024-09-26 manuscript.tex
changed figure descriptions and added color indicators to the main text

--- 2024-09-25 manuscript.tex
@@
     \includegraphics[width=1\textwidth]{silhouette_comparison.png}
-    \caption{A: silhouette plot of the k-means solution of the full dataset. B: silhouette plot of the k-means solution with the reduced dataset.}
+    \caption{Comparison of silhouette plots of two k-means cluster solutions of the Austrian vacation activities dataset with k=6.
+(A) shows the silhouette plot of the k-means solution of the full dataset and (B) the silhouette plot of the k-means solution of the dataset after manual feature selection.
+We can see that the cluster solution with the reduced dataset achieved better silhouette scores and that clusters 1 and 3 contain observations with negative silhouette scores.}
     \label{fig:silhouette_comparison}
@@
     \includegraphics[width=1\textwidth]{winter_activ_cluster_highlights.png}
-    \caption{Cluster highlights: \textbf{A} - Cluster 1 (blue), \textbf{B} - Cluster 2 (orange), \textbf{C} - Cluster 3 (green), \textbf{D} - Cluster 4 (red), \textbf{E} - Cluster 5 (violet), \textbf{F} - Cluster 6 (brown).}
+    \caption{Display of a projection of the Austrian vacation activities dataset with the six clusters of the k-means cluster solution highlighted in different colors.
+The colors indicate which clusters the highlighted observations belong to with cluster 1 being blue (A), cluster 2 being orange (B), cluster 3 being green (C), cluster 4 being red (D), cluster 5 being violett (E) and cluster 6 being brown (F).
+We can see that the projection separates some clusters well, however also that there is considerable overlap of clusters 4, 5 and 6}
     \label{fig:winter_activ_cluster_highlights}
@@
 The most influential features shown in Figure \ref{fig:winter_activ_cluster_highlights} are pool/sauna, alpine skiing, museums, and going to the spa.
-The projection roughly separates clusters 1, 2, and 3 from each other and the other three clusters, which appear to be quite similar.
+The projection roughly separates clusters 1 (blue), 2 (orange), and 3 (green) from each other and the other three clusters (red, violett and brown), which appear to be quite similar.
 
@@
     \includegraphics[width=1\textwidth]{winter_cl7_init.png}
-    \caption{Top left: 2D tour with the linear discriminant analysis projection pursuit index.
+    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset.
+Top left: 2D tour with the linear discriminant analysis projection pursuit index.
 Top right: heatmap with the intra-cluster fraction.
 Bottom left: 1D tour with the linear discriminant analysis projection pursuit index.
-Bottom right: mosaic plot.}
+Bottom right: mosaic plot.
+Touirists in both clusters 1 and 3 didn't participate in skiing a lot, but tourists in cluster 3 were much more interested in going to the pool, spa and health facilities compared to cluster 1.}
     \label{fig:winter_cl7_init}
@@
 We can see that there is indeed overlap between clusters 1 and 3, as shown in Figure \ref{fig:winter_cl7_pre}.
-As a next step we can reassign the overlapping section to a new cluster - cluster 7.
+As a next step we can reassign the overlapping section to a new cluster - cluster 7 (pink).
 By selecting the checkbox for subset 7 and manually selecting the region of overlap, we can form a new cluster, which is visualized in Figure \ref{fig:winter_cl7_post}.
@@
     \includegraphics[width=1\textwidth]{winter_cl7_pre.png}
-    \caption{Top left: 2D tour with the linear discriminant analysis projection pursuit index.
+    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset, with manually adjusted projections.
+Top left: 2D tour with the linear discriminant analysis projection pursuit index.
 Top right: heatmap with the intra-cluster fraction.
 Bottom left: 1D tour with the linear discriminant analysis projection pursuit index.
-Bottom right: mosaic plot.}
+Bottom right: mosaic plot.
+Changing the projeciton axes of "going to a spa", "museums", "sight seeing" and "using health facilities" reveals the preferences and overlap of clusters 1 and 3.}
     \label{fig:winter_cl7_pre}
@@
     \includegraphics[width=1\textwidth]{winter_cl7_post.png}
-    \caption{Top left: 2D tour with the linear discriminant analysis projection pursuit index.
+    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset, after new subsets have been selected.
+Top left: 2D tour with the linear discriminant analysis projection pursuit index.
 Top right: heatmap with the intra-cluster fraction.
 Bottom left: 1D tour with the linear discriminant analysis projection pursuit index.
-Bottom right: mosaic plot.}
+Bottom right: mosaic plot.
+We can see that now almost all museum goers are in the new manually selected subset 7 and what their preferences are compared to the other clusters.}
     \label{fig:winter_cl7_post}
@@
 
-At first, hierarchical clustering using Ward's method \citep{ward2} and the Jaccard index was applied to the features.
-The resulting dendrogram is shown in Figure \ref{fig:aus_feature _clustering}.
+At first, hierarchical clustering using Ward's method \citep{murtagh2014ward} and the Jaccard index was applied to the features.
+The resulting dendrogram is shown in Figure \ref{fig:aus_feature_clustering}.
 Based on this clustering, \( k = 15 \) clusters were identified, and generally, only one representative feature from each cluster was selected for further analysis.
@@
     \includegraphics[width=1\textwidth]{aus_feature_clustering.png}
-    \caption{Dendogram of the features of the Australian Vacation Activities dataset using Ward's method with the Jaccard index.}
+    \caption{Dendogram of the features of the Australian Vacation Activities dataset using Ward's method with the Jaccard index.
+Features which were clustered together are marked by the colored boxes.
+We can see, which activties had similar patterns in tourst interest.}
     \label{fig:aus_feature_clustering}
@@
 Tourists that fall in the direction of the axes generally engaged in more activities compared to those in the opposite direction of the projection axes.
-This separation is evident, as observations in subset 6 are located opposite the axes, and the heatmap shows that they did not engage in many activities.
-Similarly we can also see that subset 2, which contains quite active tourists, is shifted towards the direction of the axes.
+This separation is evident, as observations in subset 6 (brown) are located opposite the axes, and the heatmap shows that they did not engage in many activities.
+Similarly we can also see that subset 2 (orange), which contains quite active tourists, is shifted towards the direction of the axes.
 
-Using this logic, we can segment the data points on the left into three segments: active tourists (more upward), medium active tourists (center), and largely inactive tourists (bottom).
+Using this logic, we can segment the data points on the left into three segments: active tourists (more upward), moderately active tourists (center), and largely inactive tourists (bottom).
 The result of this segmentation can be seen in Figure \ref{fig:aus_selection}.
@@
 
-By comparing subset 2, which contains active tourists who spent time with their friends, with subset 7, the active tourists who traveled alone, we notice that subset 7 showed less interest in visiting the casino, theatre, and chartering a boat.
+By comparing subset 3 (green), which contains active tourists who spent time with their friends, with subset 7 (pink), the active tourists who traveled alone, we notice that subset 7 showed less interest in visiting the casino, theatre, and chartering a boat.
 
@@
 
-Although subset 8 was generally less active, almost everyone still engaged in sightseeing.
+Although subset 8 (grey) was generally less active, almost everyone still engaged in sightseeing.
 For them, the focus was on sightseeing, relaxing, shopping, and going to the beach, with much less interest in other activities.
@@
 
-Finally, since subset 9 is notably inactive, one might infer that they prefer spending much of their time in their accommodation.
+Finally, since subset 9 (gold) is notably inactive, one might infer that they prefer spending much of their time in their accommodation.
 Consequently, these individuals might be most interested in accommodations that offer well-equipped, comfortable living spaces with amenities that cater to relaxation and leisure.
@@
     \includegraphics[width=1\textwidth]{aus_heatmap.png}
-    \caption{Heatmap of the intra cluster fraction of the k-means solution with \( k = 6 \) and the reduced feature subset.}
+    \caption{Heatmap of the intra cluster fraction (indicated by color) of the k-means solution with \( k = 6 \) and the reduced feature subset.
+We can observe that tourists in subsets 1, 2 and 6 all prefer to travel without their friends.}
     \label{fig:aus_heatmapg}
@@
     \includegraphics[width=1\textwidth]{aus_preselection.png}
-    \caption{Separation of the data based on the feature "Friends" and the general activity.
+    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Australian vacation activities dataset, with the projection axes of the feature "Friends" pointing into one direction and all the other ones into another one.
 Top left: 2D tour.
@@
 Bottom left: Mosaic plot.
-Bottom right: Heatmap with the intra cluster fraction.}
+Bottom right: Heatmap with the intra cluster fraction.
+We can see how the feature "Friends" and the general activity level (represented by the other projection axes pointing into one direction) separates the data.}
     \label{fig:aus_preselection}
@@
     \includegraphics[width=1\textwidth]{aus_selection.png}
-    \caption{Separation of the data based on the feature "Friends" and the general activity after reselection of subsets. 
-    Tourists in subsets 7 (pink), 8 (grey) and 9 (gold) all responded thay didn't engage in activities with their friends.
-Subset 7 was the most active, subset 8 medium active and subset 9 largely inactive.
+    \caption{Interactive tour GUI as seen in \ref{fig:aus_preselection}, but after subselection of subsets 7 (pink), 8 (grey) and 9 (gold).
 Top left: 2D tour.
@@
 Bottom left: Mosaic plot.
-Bottom right: Heatmap with the intra cluster fraction.}
+Bottom right: Heatmap with the intra cluster fraction.
+We can observe the preferences of three new subsegments, which can be interpreted as very active (pink), moderately active (grey) and inactive (gold) tourists tavelling without their friends.}
     \label{fig:aus_selection}

### 2024-09-26 manuscript.tex
gave renaming the package and title a shot

@@
 %% \AND can be used instead of \And and starts a new line
-\title{Interactively Visualizing Multivariate Market Segmentation Using pytourr}
+\title{LIONFISH: an expLoratory Interactive tOol for dyNamic visualization and identiFicatIon multidimenSional mecHanisms}
 
@@
 \Plainauthor{Matthias Medl, Dianne Cook, Ursula Laa} %% comma-separated
-\Plaintitle{Interactively Visualizing Multivariate Market Segmentation Using pytourr} %% without formatting
-\Shorttitle{pytourr} %% a short title (if necessary)
+\Plaintitle{LIONFISH} %% without formatting
+\Shorttitle{LIONFISH} %% a short title (if necessary)
 

### 2024-09-27 manuscript.tex
fixing typos and spell check

@@
   Matthias Medl\\
-  Institue of Statistics\\
+  Institute of Statistics\\
   BOKU University Vienna\\
@@
 
-To integrate user interactions with the capabilities of the tourr package an active communitcation with the interface is required.
+To integrate user interactions with the capabilities of the tourr package an active communication with the interface is required.
 For example, we may wish to explore the local neighbourhood of a projection selected by the user with a local tour animation provided by tourr, or we may want to optimize a guided tour path using groups identified via brushing.
@@
 The dataset must be supplied as a \texttt{data.table}, while the plotting instructions should be passed as a list containing the named elements \texttt{type} and \texttt{obj}.
-The \texttt{type} element specifies the type of display to generate, such as "scatter" for a scatterplot or "2d\textunderscore tour" for a 2-dimensional tour.
+The \texttt{type} element specifies the type of display to generate, such as \texttt{scatter} for a scatterplot or \texttt{2d\textunderscore tour} for a 2-dimensional tour.
 The \texttt{obj} element further defines the properties of the chosen display.
@@
 Subsets can also be renamed using the provided text boxes.
-The "Reset original selection" button allows users to revert the subset selections to their initial state.
+The \texttt{Reset original selection} button allows users to revert the subset selections to their initial state.
 
-The frame selection interface (Figure \ref{fig:GUI_overview}E) displayes the frames of the currently shown tours and enables users to jump between frames of the tour objects.
+The frame selection interface (Figure \ref{fig:GUI_overview}E) displays the frames of the currently shown tours and enables users to jump between frames of the tour objects.
 Below this, users can save and load projections and subsets using the respective buttons (Figure \ref{fig:GUI_overview}F).
@@
 
-One can move through the projections of a tour by pressing the arrow keys or by specifying the index of the projection the be displayed and pressing the "Update frames" button.
-Users can animate the tours by toggling the "Animate" checkbox and specifying a time interval, so that the displays automatically move to the next projections after the specified time interval.
+One can move through the projections of a tour by pressing the arrow keys or by specifying the index of the projection the be displayed and pressing the \texttt{Update frames} button.
+Users can animate the tours by toggling the \texttt{Animate} checkbox and specifying a time interval, so that the displays automatically move to the next projections after the specified time interval.
 
-The "Save projections and subsets" button can be used to save the current state of the analysis.
-These states can be recovered by using pressing the "Load projections and subsets" button.
+The \texttt{Save projections and subsets} button can be used to save the current state of the analysis.
+These states can be recovered by using pressing the \texttt{Load projections and subsets} button.
 Additionally, users can initiate new tours of various directly from the sidebar.
@@
 The central component of the Python code is the \texttt{customTKinter} class \texttt{InteractiveTourInterface}.
-This class centrally stores attributes related to all plots, such as the dataset, subselections, feature selections, and other shared information.
+This class centrally stores attributes related to all plots, such as the dataset, sub-selections, feature selections, and other shared information.
 Plot-specific data is organized in dictionaries (the Python equivalent of named lists in R), including the display type, construction instructions, tour projections (only in case of tour displays), color schemes for the displayed data, and, where applicable, the selector and manual projection manipulation classes.
@@
 The selector classes handle the behavior when users manually select data points to move them to the active subset.
-After a selection is made, the selector class updates the centrally stored subselection attribute and ensures all other displays reflect these changes.
+After a selection is made, the selector class updates the centrally stored sub-selection attribute and ensures all other displays reflect these changes.
 The manual projection manipulation classes construct the arrows representing the projection axes in the displays and manage the manual adjustment of projections, and thus enable manual tours.
@@
 A first step may look into what variables are important for the cluster solution.
-The interface provides different capabilites that support variable selection.
-First, a 2D tour view can be used to understand the sensitivity of the grouping to individual varaibles, in particular when using the manual tour to remove one variable at a time.
+The interface provides different capabilities that support variable selection.
+First, a 2D tour view can be used to understand the sensitivity of the grouping to individual variables, in particular when using the manual tour to remove one variable at a time.
 This is in particular useful when the starting view was obtained through a projection pursuit guided tour~\cite{CBCH94} that optimized for the separation between the labeled groups~\cite{lckl2005}.
@@
 \]
-to see how the true values for variable $j$ are distributed accross the different clusters relative, with normalization to the overall count of that feature \(c_j = \sum_i c_{ij}\).
+to see how the true values for variable $j$ are distributed across the different clusters relative, with normalization to the overall count of that feature \(c_j = \sum_i c_{ij}\).
 
@@
 
-One measure to combat this is to make saving intermediate snapshots of the analysis easy and accessable.
-Specifically, the "Save projections and subsets" button enables users to take snapshots of their analysis, including visual representations, selected data, and parameter settings.
+One measure to combat this is to make saving intermediate snapshots of the analysis easy and accessible.
+Specifically, the \texttt{Save projections and subsets} button enables users to take snapshots of their analysis, including visual representations, selected data, and parameter settings.
 Upon pressing this button, a file browser is triggered, allowing users to specify the destination for saving these snapshots.
@@
 First, they allow users to fully recover the state of the analysis within the GUI.
-This can be achieved either by using the "Load projections and subsets" button, or by launching a new GUI instance with the \texttt{load\_interactive\_tour()} function.
+This can be achieved either by using the \texttt{Load projections and subsets} button, or by launching a new GUI instance with the \texttt{load\_interactive\_tour()} function.
 The latter approach, using \texttt{load\_interactive\_tour()}, has the added flexibility of only requiring the original dataset and the directory containing the saved files.
@@
 Participants were asked to evaluate the importance of 27 different activities during their vacation.
-The survey categorized responses based on four levels of importance: "totally important," "mostly important," "a bit important," and "not important." For analysis, the responses were binarized: a value of 1 was assigned if the activity was rated as "totally important," and a value of 0 if any of the other categories were selected.
+The survey categorized responses based on four levels of importance: ``totally important", ``mostly important", ``a bit important", and ``not important".
+For analysis, the responses were binarized: a value of 1 was assigned if the activity was rated as ``totally important", and a value of 0 if any of the other categories were selected.
 The survey was conducted by the Europäisches Tourismus Institut GmbH at the University of Trier and focused exclusively on tourists who did not stay in the country's capital cities.
 
-To gain further insight into the dataset a k-means clustering as described in the book "Market Segmentation Analysis" has been performed \citep{leisch2018market}.
-Therefore, the function \texttt{stepcclust} of the R package \texttt{flexclust
-}\citep{flexclust} with k=6 and nrep=20 was used.
+To gain further insight into the dataset a k-means clustering as described in \cite{leisch2018market} has been performed.
+Therefore, the function \texttt{stepcclust} of the R package \texttt{flexclust}\citep{flexclust} with $k=6$ and $nrep=20$ was used.
 
@@
 By comparing both silhouette plots, we can see that the cluster solution with the reduced dataset results in a clustering of higher quality.
-Thus, we will continue with the analysis on the reduced dataset with the corresponsig cluster solution.
+Thus, we will continue with the analysis on the reduced dataset with the corresponding cluster solution.
 We can also see in Figure \ref{fig:silhouette_comparison}B that cluster 3 is of comparatively low quality.
@@
     \caption{Display of a projection of the Austrian vacation activities dataset with the six clusters of the k-means cluster solution highlighted in different colors.
-The colors indicate which clusters the highlighted observations belong to with cluster 1 being blue (A), cluster 2 being orange (B), cluster 3 being green (C), cluster 4 being red (D), cluster 5 being violett (E) and cluster 6 being brown (F).
+The colors indicate which clusters the highlighted observations belong to with cluster 1 being blue (A), cluster 2 being orange (B), cluster 3 being green (C), cluster 4 being red (D), cluster 5 being violet (E) and cluster 6 being brown (F).
 We can see that the projection separates some clusters well, however also that there is considerable overlap of clusters 4, 5 and 6}
@@
 The most influential features shown in Figure \ref{fig:winter_activ_cluster_highlights} are pool/sauna, alpine skiing, museums, and going to the spa.
-The projection roughly separates clusters 1 (blue), 2 (orange), and 3 (green) from each other and the other three clusters (red, violett and brown), which appear to be quite similar.
+The projection roughly separates clusters 1 (blue), 2 (orange), and 3 (green) from each other and the other three clusters (red, violet and brown), which appear to be quite similar.
 
@@
 
-\subsubsection{Redefing cluster assignments - learning more about museum goers}
+\subsubsection{Redefining cluster assignments - learning more about museum goers}
 
@@
 Bottom right: mosaic plot.
-Touirists in both clusters 1 and 3 didn't participate in skiing a lot, but tourists in cluster 3 were much more interested in going to the pool, spa and health facilities compared to cluster 1.}
+Tourists in both clusters 1 and 3 didn't participate in skiing a lot, but tourists in cluster 3 were much more interested in going to the pool, spa and health facilities compared to cluster 1.}
     \label{fig:winter_cl7_init}
@@
 Bottom right: mosaic plot.
-Changing the projeciton axes of "going to a spa", "museums", "sight seeing" and "using health facilities" reveals the preferences and overlap of clusters 1 and 3.}
+Changing the projection axes of ``going to a spa", ``museums", ``sightseeing" and ``using health facilities" reveals the preferences and overlap of clusters 1 and 3.}
     \label{fig:winter_cl7_pre}
@@
 The second dataset, the Australian Vacation Activities dataset, includes responses from 1,003 adult Australians who were surveyed through a permission-based internet panel.
-The survery was conducted in 2007.
+The survey was conducted in 2007.
 Participants were asked whether they engaged in 44 specific vacation activities during their most recent vacation within Australia.
 Similar to the Austrian dataset, responses were binarized: a value of 1 indicates that the participant took part in the activity, while a value of 0 signifies they did not.
-Suvreys where participants claimed they partook in more than 40 activities or no activity at all were removed as they are considered faulty.
+Surveys where participants claimed they partook in more than 40 activities or no activity at all were removed as they are considered faulty.
 
@@
 Based on this clustering, \( k = 15 \) clusters were identified, and generally, only one representative feature from each cluster was selected for further analysis.
-Clusters containing unpopular activities, such as "Adventure", which only had 42 participants, were discarded.
-The cluster containing popular features like "Beach," "Swimming," "ScenicWalks," "Markets," "Sightseeing," "Friends," "Pubs," "BBQ," "Shopping," "Eating," "EatingHigh," "Movies," and "Relaxing" was treated differently.
+Clusters containing unpopular activities, such as ``Adventure", which only had 42 participants, were discarded.
+The cluster containing popular features like ``Beach", ``Swimming", ``ScenicWalks", ``Markets", ``Sightseeing", ``Friends", ``Pubs", ``BBQ", ``Shopping", ``Eating", ``EatingHigh", ``Movies," and ``Relaxing" was treated differently.
 Multiple features from this cluster were retained to preserve as much information as possible.
@@
 Features which were clustered together are marked by the colored boxes.
-We can see, which activties had similar patterns in tourst interest.}
+We can see, which activities had similar patterns in tourist interest.}
     \label{fig:aus_feature_clustering}
@@
 
-Since we are particularly interested in tourists who did not spend time with friends, we can extend the projection axis "Friends" outward to separate the data based on that feature.
-Datapoints in the opposite direction of the "Friends" axes are the solo travellers we are interested in, which can be seen in Figure \ref{fig:aus_preselection} on the left side of the top left plot.
+Since we are particularly interested in tourists who did not spend time with friends, we can extend the projection axis ``Friends" outward to separate the data based on that feature.
+Datapoints in the opposite direction of the ``Friends" axes are the solo travellers we are interested in, which can be seen in Figure \ref{fig:aus_preselection} on the left side of the top left plot.
 Subsequently, we can pull all other projection axes in one direction to separate data points based on their overall activity level.
@@
 Given that subset 7 showed a notable interest in going to pubs, we can assume they want to meet new people.
-This insight can be utilized in a marketing campaign that bundles the activities subset 7 is interested in, creating a package tailored for solo travelers.
-This way, they can meet other solo travelers while engaging in activities they enjoy.
+This insight can be utilized in a marketing campaign that bundles the activities subset 7 is interested in, creating a package tailored for solo travellers.
+This way, they can meet other solo travellers while engaging in activities they enjoy.
 
@@
     \includegraphics[width=1\textwidth]{aus_preselection.png}
-    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Australian vacation activities dataset, with the projection axes of the feature "Friends" pointing into one direction and all the other ones into another one.
+    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Australian vacation activities dataset, with the projection axes of the feature ``Friends" pointing into one direction and all the other ones into another one.
 Top left: 2D tour.
@@
 Bottom right: Heatmap with the intra cluster fraction.
-We can see how the feature "Friends" and the general activity level (represented by the other projection axes pointing into one direction) separates the data.}
+We can see how the feature ``Friends" and the general activity level (represented by the other projection axes pointing into one direction) separates the data.}
     \label{fig:aus_preselection}
@@
     \includegraphics[width=1\textwidth]{aus_selection.png}
-    \caption{Interactive tour GUI as seen in \ref{fig:aus_preselection}, but after subselection of subsets 7 (pink), 8 (grey) and 9 (gold).
+    \caption{Interactive tour GUI as seen in \ref{fig:aus_preselection}, but after sub-selection of subsets 7 (pink), 8 (grey) and 9 (gold).
 Top left: 2D tour.
@@
 Bottom right: Heatmap with the intra cluster fraction.
-We can observe the preferences of three new subsegments, which can be interpreted as very active (pink), moderately active (grey) and inactive (gold) tourists tavelling without their friends.}
+We can observe the preferences of three new subsegments, which can be interpreted as very active (pink), moderately active (grey) and inactive (gold) tourists travelling without their friends.}
     \label{fig:aus_selection}
@@
 
-Similarly, in the Australian dataset, manual exploration allowed us to refine the understanding of solo travelers by dividing them into three distinct subsegments: highly active, moderately active, and largely inactive tourists.
+Similarly, in the Australian dataset, manual exploration allowed us to refine the understanding of solo travellers by dividing them into three distinct subsegments: highly active, moderately active, and largely inactive tourists.
 This nuanced segmentation, derived from a blend of automated clustering and manual adjustments, offers a deeper understanding of the varied needs and behaviors within this group, enabling more effective marketing strategies.

### 2024-09-27 manuscript.tex
refs to citep in places

@@
 First, a 2D tour view can be used to understand the sensitivity of the grouping to individual variables, in particular when using the manual tour to remove one variable at a time.
-This is in particular useful when the starting view was obtained through a projection pursuit guided tour~\cite{CBCH94} that optimized for the separation between the labeled groups~\cite{lckl2005}.
+This is in particular useful when the starting view was obtained through a projection pursuit guided tour~\citep{CBCH94} that optimized for the separation between the labeled groups~\citep{lckl2005}.
 

### 2024-09-27 manuscript.tex
quotes to proper double quotes

@@
 Participants were asked to evaluate the importance of 27 different activities during their vacation.
-The survey categorized responses based on four levels of importance: ``totally important", ``mostly important", ``a bit important", and ``not important".
-For analysis, the responses were binarized: a value of 1 was assigned if the activity was rated as ``totally important", and a value of 0 if any of the other categories were selected.
+The survey categorized responses based on four levels of importance: ``totally important'', ``mostly important'', ``a bit important'', and ``not important".
+For analysis, the responses were binarized: a value of 1 was assigned if the activity was rated as ``totally important'', and a value of 0 if any of the other categories were selected.
 The survey was conducted by the Europäisches Tourismus Institut GmbH at the University of Trier and focused exclusively on tourists who did not stay in the country's capital cities.
@@
 Bottom right: mosaic plot.
-Changing the projection axes of ``going to a spa", ``museums", ``sightseeing" and ``using health facilities" reveals the preferences and overlap of clusters 1 and 3.}
+Changing the projection axes of ``going to a spa'', ``museums'', ``sightseeing" and ``using health facilities" reveals the preferences and overlap of clusters 1 and 3.}
     \label{fig:winter_cl7_pre}
@@
 Based on this clustering, \( k = 15 \) clusters were identified, and generally, only one representative feature from each cluster was selected for further analysis.
-Clusters containing unpopular activities, such as ``Adventure", which only had 42 participants, were discarded.
-The cluster containing popular features like ``Beach", ``Swimming", ``ScenicWalks", ``Markets", ``Sightseeing", ``Friends", ``Pubs", ``BBQ", ``Shopping", ``Eating", ``EatingHigh", ``Movies," and ``Relaxing" was treated differently.
+Clusters containing unpopular activities, such as ``Adventure'', which only had 42 participants, were discarded.
+The cluster containing popular features like ``Beach'', ``Swimming'', ``ScenicWalks'', ``Markets'', ``Sightseeing'', ``Friends'', ``Pubs'', ``BBQ'', ``Shopping'', ``Eating'', ``EatingHigh'', ``Movies'', and ``Relaxing'' was treated differently.
 Multiple features from this cluster were retained to preserve as much information as possible.

### 2024-09-27 manuscript.tex
more quotes to proper double quotes

@@
 Participants were asked to evaluate the importance of 27 different activities during their vacation.
-The survey categorized responses based on four levels of importance: ``totally important'', ``mostly important'', ``a bit important'', and ``not important".
+The survey categorized responses based on four levels of importance: ``totally important'', ``mostly important'', ``a bit important'', and ``not important''.
 For analysis, the responses were binarized: a value of 1 was assigned if the activity was rated as ``totally important'', and a value of 0 if any of the other categories were selected.
@@
 Bottom right: mosaic plot.
-Changing the projection axes of ``going to a spa'', ``museums'', ``sightseeing" and ``using health facilities" reveals the preferences and overlap of clusters 1 and 3.}
+Changing the projection axes of ``going to a spa'', ``museums'', ``sightseeing'' and ``using health facilities'' reveals the preferences and overlap of clusters 1 and 3.}
     \label{fig:winter_cl7_pre}
@@
 
-Since we are particularly interested in tourists who did not spend time with friends, we can extend the projection axis ``Friends" outward to separate the data based on that feature.
-Datapoints in the opposite direction of the ``Friends" axes are the solo travellers we are interested in, which can be seen in Figure \ref{fig:aus_preselection} on the left side of the top left plot.
+Since we are particularly interested in tourists who did not spend time with friends, we can extend the projection axis ``Friends'' outward to separate the data based on that feature.
+Datapoints in the opposite direction of the ``Friends'' axes are the solo travellers we are interested in, which can be seen in Figure \ref{fig:aus_preselection} on the left side of the top left plot.
 Subsequently, we can pull all other projection axes in one direction to separate data points based on their overall activity level.
@@
     \includegraphics[width=1\textwidth]{aus_preselection.png}
-    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Australian vacation activities dataset, with the projection axes of the feature ``Friends" pointing into one direction and all the other ones into another one.
+    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Australian vacation activities dataset, with the projection axes of the feature ``Friends'' pointing into one direction and all the other ones into another one.
 Top left: 2D tour.
@@
 Bottom right: Heatmap with the intra cluster fraction.
-We can see how the feature ``Friends" and the general activity level (represented by the other projection axes pointing into one direction) separates the data.}
+We can see how the feature ``Friends'' and the general activity level (represented by the other projection axes pointing into one direction) separates the data.}
     \label{fig:aus_preselection}

### 2024-09-27 manuscript.tex
small typo

@@
 In such cases, manual exploration allows analysts to interweave expert knowledge, intuition, and specific objectives with the initial clustering solution, which serves as the backbone of the analysis.
-This approach is particularly valuable when no single solution can be deemed “correct.” For instance, in the Austrian dataset, leveraging the interactive GUI for feature selection enabled us to isolate the most informative activities and gain deeper insights into the preferences of tourists who might be interested in visiting museums.
+This approach is particularly valuable when no single solution can be deemed “correct''.
+For instance, in the Austrian dataset, leveraging the interactive GUI for feature selection enabled us to isolate the most informative activities and gain deeper insights into the preferences of tourists who might be interested in visiting museums.
 This understanding facilitated targeted recommendations for increasing museum attendance by focusing on tourists frequenting hiking trails, excursion spots, and shopping centers.

### 2024-09-27 manuscript.tex
three gt projections of the austrian data

@@
 The survey was conducted by the Europäisches Tourismus Institut GmbH at the University of Trier and focused exclusively on tourists who did not stay in the country's capital cities.
+Figure \ref{winter-gt} shows three projections from a grand tour of this data: when mostly two variables contribute to a projection the data will look apparently clustered, due to the binary values, but when more variables contribute to the projection it looks more continuous.
+
+\begin{figure*}[h]
+\centerline{\includegraphics[width=0.3\textwidth]{images/winter_proj1.pdf}\includegraphics[width=0.3\textwidth]{images/winter_proj2.pdf}\includegraphics[width=0.3\textwidth]{images/winter_proj3.pdf}}
+\caption{Three 2D projections from a grand tour on the Austrian Vacation Activities data.
+Because the 27 variables are binary, you'll see some apparent clusters (left) but with this many binary variables the data mostly looks continuous.}
+\label{winter-gt}
+\end{figure*}
 

### 2024-09-27 scripts/austrian_tourism.R

@@ -1,0 +1,33 @@
+# Application: Austrian tourism
+# Load libraries
+library(tourr)
+library(pytourr)
+
+# Check random projections
+data(winterActiv)
+winterActiv_std <- apply(winterActiv, 2, scale)
+set.seed(630)
+t1 <- save_history(winterActiv, max=20)
+t1i <- interpolate(t1)
+animate_xy(winterActiv_std, planned_tour(t1), half_range=10)
+render_gif(winterActiv_std, 
+           planned_tour(t1), 
+           display_xy(axes="off", half_range=10), 
+           gif_file = "images/winterActiv.gif",
+           width = 400, height = 400)
+
+# Save projections to illustrate the binary nature
+proj <- matrix(t1i[,,3], nrow=27, ncol=2)
+proj <- matrix(t1i[,,60], nrow=27, ncol=2)
+proj <- matrix(t1i[,,30], nrow=27, ncol=2)
+p <- render_proj(winterActiv_std, proj)
+ggplot() +
+  geom_point(data=p$data_prj, aes(x=P1, y=P2)) +
+  #xlim(-1,1) + ylim(-1, 1) +
+  theme_bw() +
+  theme(aspect.ratio=1,
+        axis.text=element_blank(),
+        axis.title=element_blank(),
+        axis.ticks=element_blank(),
+        panel.grid=element_blank())
+

### 2024-09-27 manuscript.tex
changed pytourr to lionfish

--- 2024-09-27 manuscript.tex
@@
 %% \AND can be used instead of \And and starts a new line
-\title{LIONFISH: an expLoratory Interactive tOol for dyNamic visualization and identiFicatIon multidimenSional mecHanisms}
+\title{Interactively Visualizing Clustering in Multivariate Data Using lionfish}
 
@@
 \Plainauthor{Matthias Medl, Dianne Cook, Ursula Laa} %% comma-separated
-\Plaintitle{LIONFISH} %% without formatting
-\Shorttitle{LIONFISH} %% a short title (if necessary)
+\Plaintitle{lionfish} %% without formatting
+\Shorttitle{lionfish} %% a short title (if necessary)
 
@@
 
-%At the core of the \texttt{pytourr} package is a graphical user interface (GUI) designed to display multiple linked visualizations concurrently.
+%At the core of the \texttt{lionfish} package is a graphical user interface (GUI) designed to display multiple linked visualizations concurrently.
 Users can interact with one of the plots, for example, by subselecting specific data points, and the other linked plots will dynamically update to reflect these modifications.
@@
 
-The interface was implemented in \texttt{pytourr} and offers a variety of linked interactive plot types, providing users with the flexibility to visualize their data from multiple perspectives.
+The interface was implemented in \texttt{lionfish} and offers a variety of linked interactive plot types, providing users with the flexibility to visualize their data from multiple perspectives.
 The ability to navigate through various projections of the displayed tours directly within a graphical user interface (GUI) enables users to explore different aspects of the dataset.
@@
 The GUI also supports interactive variable selection, allowing users to specify which subset of variables should be visualized in the plots.
-Once users have identified interesting views or settings, \texttt{pytourr} allows them to save the displayed projections, subsets, and plots.
+Once users have identified interesting views or settings, \texttt{lionfish} allows them to save the displayed projections, subsets, and plots.
 This functionality ensures that analysis states can be preserved for further examination or reporting, making the package particularly useful for iterative analysis where findings may need to be revisited or shared with collaborators.
 
-With its high level of interactivity, performance, and ease of use, \texttt{pytourr} streamlines the exploration of complex datasets, offering a powerful tool for researchers working with high-dimensional data.
+With its high level of interactivity, performance, and ease of use, \texttt{lionfish} streamlines the exploration of complex datasets, offering a powerful tool for researchers working with high-dimensional data.
 
@@
 
-The \texttt{pytourr} GUI can be launched using the function \texttt{interactive\textunderscore tour()}.
+The \texttt{lionfish} GUI can be launched using the function \texttt{interactive\textunderscore tour()}.
 At a minimum, the user needs to provide both the dataset and the instructions for constructing the desired plots.
@@
 
-The majority of \texttt{pytourr} was written in Python, while the R side of the package handles setting up the Python environment within the interactive interface is being run, launching the interactive tour, and generating new tours when initiated through the interface.
+The majority of \texttt{lionfish} was written in Python, while the R side of the package handles setting up the Python environment within the interactive interface is being run, launching the interactive tour, and generating new tours when initiated through the interface.
 To incorporate the functionality of the \texttt{tourr} package without translating large portions of its code from R to Python, the \texttt{reticulate} package was used.
-This approach allows \texttt{pytourr} to automatically benefit from updates to \texttt{tourr}.
+This approach allows \texttt{lionfish} to automatically benefit from updates to \texttt{tourr}.
 
@@
 
-\section{Workflow with the pytourr package}~\label{workflow}
+\section{Workflow with the lionfish package}~\label{workflow}
 
-Before launching the pytourr interface the user should have performed a partitioning of their choice, and provide the initial clustering solution used by the interface.
+Before launching the lionfish interface the user should have performed a partitioning of their choice, and provide the initial clustering solution used by the interface.
 We launch the interface to explore and potentially refine this solution.
@@
 We only want to keep the most informative ones.
-Additionally, interacting with the pytourr GUI becomes cumbersome when handling more than \textasciitilde
- 15 features, making it necessary to reduce the dimensionality of the dataset. An effective and intuitive way to perform feature selection is by using the heatmap display within pytourr.
+Additionally, interacting with the lionfish GUI becomes cumbersome when handling more than \textasciitilde
+ 15 features, making it necessary to reduce the dimensionality of the dataset. An effective and intuitive way to perform feature selection is by using the heatmap display within lionfish.
 
@@
 
-While some plots supported by the pytourr package are specifically designed for analyzing binary survey data, it is important to emphasize that its capabilities extend far beyond this data type.
+While some plots supported by the lionfish package are specifically designed for analyzing binary survey data, it is important to emphasize that its capabilities extend far beyond this data type.
 The versatility of tours has been demonstrated across various applications in the past, showcasing their effectiveness in exploring complex, high-dimensional datasets.
@@
 This approach enables analysts to refine initial results and adapt them to the complexities of real-world scenarios, leading to more nuanced interpretations and actionable insights.
-The pytourr package offers an organised and responsive interactive tool for conducting such analyses, bridging the gap between automated clustering and exploration.
+The lionfish package offers an organised and responsive interactive tool for conducting such analyses, bridging the gap between automated clustering and exploration.
 
-By integrating interactive visualization capabilities, the pytourr package empowers users to dynamically engage with their data, making it possible to uncover subtle patterns and relationships that might otherwise remain hidden.
+By integrating interactive visualization capabilities, the lionfish package empowers users to dynamically engage with their data, making it possible to uncover subtle patterns and relationships that might otherwise remain hidden.
 This is especially valuable in tackling complex datasets with the mindset that an automated solution needs to be validated.
@@
 
-In summary, pytourr represents a significant advancement in the toolkit of data analysts, offering a novel way to balance automated analysis with human intuition and domain expertise, thereby facilitating a deeper and more comprehensive understanding of complex datasets.
+In summary, lionfish represents a significant advancement in the toolkit of data analysts, offering a novel way to balance automated analysis with human intuition and domain expertise, thereby facilitating a deeper and more comprehensive understanding of complex datasets.
 
 %\bibliographystyle{plainat}
-\bibliography{pytourr_references.bib}
+\bibliography{lionfish_references.bib}
 

### 2024-09-27 manuscript.tex
changed typesetting of references to lowercase

@@
 
-Before launching the lionfish interface the user should have performed a partitioning of their choice, and provide the initial clustering solution used by the interface.
+Before launching the \texttt{lionfish} interface the user should have performed a partitioning of their choice, and provide the initial clustering solution used by the interface.
 We launch the interface to explore and potentially refine this solution.
@@
 We only want to keep the most informative ones.
-Additionally, interacting with the lionfish GUI becomes cumbersome when handling more than \textasciitilde
- 15 features, making it necessary to reduce the dimensionality of the dataset. An effective and intuitive way to perform feature selection is by using the heatmap display within lionfish.
+Additionally, interacting with the \texttt{lionfish} GUI becomes cumbersome when handling more than \textasciitilde
+ 15 features, making it necessary to reduce the dimensionality of the dataset. An effective and intuitive way to perform feature selection is by using the heatmap display within \texttt{lionfish}.
 
@@
 
-While some plots supported by the lionfish package are specifically designed for analyzing binary survey data, it is important to emphasize that its capabilities extend far beyond this data type.
+While some plots supported by the \texttt{lionfish} package are specifically designed for analyzing binary survey data, it is important to emphasize that its capabilities extend far beyond this data type.
 The versatility of tours has been demonstrated across various applications in the past, showcasing their effectiveness in exploring complex, high-dimensional datasets.
@@
 This approach enables analysts to refine initial results and adapt them to the complexities of real-world scenarios, leading to more nuanced interpretations and actionable insights.
-The lionfish package offers an organised and responsive interactive tool for conducting such analyses, bridging the gap between automated clustering and exploration.
+The \texttt{lionfish} package offers an organised and responsive interactive tool for conducting such analyses, bridging the gap between automated clustering and exploration.
 
-By integrating interactive visualization capabilities, the lionfish package empowers users to dynamically engage with their data, making it possible to uncover subtle patterns and relationships that might otherwise remain hidden.
+By integrating interactive visualization capabilities, the \texttt{lionfish} package empowers users to dynamically engage with their data, making it possible to uncover subtle patterns and relationships that might otherwise remain hidden.
 This is especially valuable in tackling complex datasets with the mindset that an automated solution needs to be validated.
@@
 
-In summary, lionfish represents a significant advancement in the toolkit of data analysts, offering a novel way to balance automated analysis with human intuition and domain expertise, thereby facilitating a deeper and more comprehensive understanding of complex datasets.
+In summary, \texttt{lionfish} represents a significant advancement in the toolkit of data analysts, offering a novel way to balance automated analysis with human intuition and domain expertise, thereby facilitating a deeper and more comprehensive understanding of complex datasets.
 

### 2024-09-27 manuscript.tex
changed variable to feature and adjusted the metric descriptions

@@
 With partitioning, it is usual that there are no gaps in the data, but it is still useful to partition the data.
-Figure \ref{kmeans-partition} illustrates how the $k$-means algorithm would partition a 2D data set into four groups depending on the correlation between the two variables.
-When the correlation is high, the partitioning will be along the combination of variables that produces the highest variance.
+Figure \ref{kmeans-partition} illustrates how the $k$-means algorithm would partition a 2D data set into four groups depending on the correlation between the two features.
+When the correlation is high, the partitioning will be along the combination of features that produces the highest variance.
 With lower correlation, it will segment the bottom and top, and divide the middle into two parts in the opposite direction.
@@
 One can see that the data is perfectly divided into four parts.
-However, if one were to plot the two variables individually, this would not be obvious.
-Figure \ref{kmeans-histogram} shows histograms of the two variables, V1, V2, with the colour matching the four partitions.
+However, if one were to plot the two features individually, this would not be obvious.
+Figure \ref{kmeans-histogram} shows histograms of the two features, V1, V2, with the colour matching the four partitions.
 From the histograms, we can see some differences in the partitions for the four different association structures, but they are all overlapping.
-The distinct border between the partitions can only be seen from the scatterplots of both variables.
+The distinct border between the partitions can only be seen from the scatterplots of both features.
 
@@
 \centerline{\includegraphics[width=1\textwidth]{images/intro2.png}}
-\caption{The four $k$-means partitions plotted as histograms of the individual variables.
+\caption{The four $k$-means partitions plotted as histograms of the individual features.
 While differences between groups can be seen, the clear separation cannot.
@@
 
-When there are more than two variables, histograms of the individual variables are commonly used to display the partitioning results.
+When there are more than two features, histograms of the individual features are commonly used to display the partitioning results.
 This means that the analyst likely cannot understand how the partitioning divides the data.
-All that they can observe is roughly how the individual variables relate to the partition, which is useful but inadequate.
+All that they can observe is roughly how the individual features relate to the partition, which is useful but inadequate.
 Here is where using tour methods to view high dimensions can be helpful.
-A tour (\cite{Asimov1985-xr}, and see \cite{lee2022}) can be used to show scatterplots of combinations of variables and thus provide views like that in Figure \ref{kmeans-partition} where distinct differences between partitions can be observed.
+A tour (\cite{Asimov1985-xr}, and see \cite{lee2022}) can be used to show scatterplots of combinations of features and thus provide views like that in Figure \ref{kmeans-partition} where distinct differences between partitions can be observed.
 High dimensions are still tricky, and a combination of animations of the linear combinations, and interactive control \citep{cook_manual_1997, laa2023new} over the combinations is important.
-A scatterplot of a combination of variables can be considered to be a projection of the data, and thus like a shadow of a 3D object, some aspects of the data (object) can be obscured.
+A scatterplot of a combination of features can be considered to be a projection of the data, and thus like a shadow of a 3D object, some aspects of the data (object) can be obscured.
 Using slices of the projected data \citep{Laa2020} can be a useful addition to projections.
@@
 Our solution is using Python for high-performance interactivity, through the packages \texttt{TKinter} \citep{lundh1999introduction}, \texttt{CustomTKinter} \citep{schimansky24}, and \texttt{matplotlib} \citep{Hunter:2007}, with integration to the tourr package \citep{tourr} via \texttt{reticulate} \citep{reticulate}, a framework that facilitates seamless interoperability between Python and R.
-\textcolor{red}{also rpy2?}
 
@@
 Furthermore, users can initiate new tours directly from the interface.
-The GUI also supports interactive variable selection, allowing users to specify which subset of variables should be visualized in the plots.
+The GUI also supports interactive feature selection, allowing users to specify which subset of features should be visualized in the plots.
 Once users have identified interesting views or settings, \texttt{lionfish} allows them to save the displayed projections, subsets, and plots.
@@
 
-\subsection{Variable selection}
+\subsection{Feature selection based on fractional analysis}
 
-A first step may look into what variables are important for the cluster solution.
-The interface provides different capabilities that support variable selection.
-First, a 2D tour view can be used to understand the sensitivity of the grouping to individual variables, in particular when using the manual tour to remove one variable at a time.
+A first step may look into what features are important for the cluster solution.
+The interface provides different capabilities that support feature selection.
+First, a 2D tour view can be used to understand the sensitivity of the grouping to individual features, in particular when using the manual tour to remove one feature at a time.
 This is in particular useful when the starting view was obtained through a projection pursuit guided tour~\citep{CBCH94} that optimized for the separation between the labeled groups~\citep{lckl2005}.
 
-In our applications we are considering market segmentation analysis on binary variables.
-For this setting we have developed a heatmap display that we have found to be particularly useful in variable selection.
+In our applications we are considering market segmentation analysis on binary features.
+For this setting we have developed a heatmap display that we have found to be particularly useful in feature selection.
 The idea is to understand differences in the binary ratings between clusters, where we use different normalizations to better explore the data.
-For \(c_{ij}\) the count (number of true values in the binary variable) in the \(i\)-th cluster and \(j\)-th variable, we define
+For \(c_{ij}\) the count (number of true values in the binary feature) in the \(i\)-th cluster and \(j\)-th feature, we define
 
 \[
-\text{Overall count fraction}\ f_{ij} = \frac{c_{ij}}{n}
+\text{Total fraction}\ f_{ij}^{tot} = \frac{c_{ij}}{n}
 \]
@@
 \[
-\text{Cluster count fraction}\  f_{ij} = \frac{c_{ij}}{n_{i}}
+\text{Intra cluster fraction}\  f_{ij}^{clust} = \frac{c_{ij}}{n_{i}}
 \]
-showing the counts relative to the cluster's size \(n_i\), to see the fraction of observations in cluster $i$ for which variable $j$ was true, and
+showing the counts relative to the cluster's size \(n_i\), to see the fraction of observations in cluster $i$ for which feature $j$ was true, and
 
 \[
-\text{Variable count fraction}\ f_{ij} = \frac{c_{ij}}{c_{j}}
+\text{Intra feature fraction}\ f_{ij}^{feat} = \frac{c_{ij}}{c_{j}}
 \]
-to see how the true values for variable $j$ are distributed across the different clusters relative, with normalization to the overall count of that feature \(c_j = \sum_i c_{ij}\).
+to see how the true values for feature $j$ are distributed across the different clusters relative, with normalization to the overall count of that feature \(c_j = \sum_i c_{ij}\).
 
@@
     \item \texttt{.csv} files that capture the feature and subset selection as well as projections of the tours displayed at the time of the snapshot,
-    \item two \texttt{.pkl} files that contain state variables of the GUI, allowing for complete recovery of the snapshot.
+    \item two \texttt{.pkl} files that contain state features of the GUI, allowing for complete recovery of the snapshot.
 \end{itemize}
@@
 The survey was conducted by the Europäisches Tourismus Institut GmbH at the University of Trier and focused exclusively on tourists who did not stay in the country's capital cities.
-Figure \ref{winter-gt} shows three projections from a grand tour of this data: when mostly two variables contribute to a projection the data will look apparently clustered, due to the binary values, but when more variables contribute to the projection it looks more continuous.
+Figure \ref{winter-gt} shows three projections from a grand tour of this data: when mostly two features contribute to a projection the data will look apparently clustered, due to the binary values, but when more features contribute to the projection it looks more continuous.
 
@@
 \caption{Three 2D projections from a grand tour on the Austrian Vacation Activities data.
-Because the 27 variables are binary, you'll see some apparent clusters (left) but with this many binary variables the data mostly looks continuous.}
+Because the 27 features are binary, you'll see some apparent clusters (left) but with this many binary features the data mostly looks continuous.}
 \label{winter-gt}

### 2024-09-29 scripts/feature_selection.R
script for explaining the heatmaps

@@ -1,0 +1,24 @@
+# Feature selection based on fractions
+library(tidyverse)
+library(tourr)
+set.seed(446)
+n <- 120
+d <- tibble(x1 = sample(c(0,1), n, replace=TRUE),
+            x2 = sample(c(0,1), n, replace=TRUE),
+            x3 = sample(c(0,1), n, replace=TRUE),
+            x4 = sample(c(0,1), n, replace=TRUE))
+d_km <- kmeans(d, 4)
+
+d <- d |>
+  mutate(cl = factor(d_km$cluster))
+# animate_xy(d[,1:4], col=d$cl)
+d_tab <- d_km$centers |>
+  as_tibble() |>
+  mutate(cl = paste0("cl", 1:4)) |>
+  pivot_longer(x1:x4, names_to = "feature", values_to = "mean")
+
+ggplot(d_tab, aes(x=cl, y=feature, fill=mean)) +
+  geom_tile() +
+  scale_fill_viridis_c(option="magma")
+
+

### 2024-09-30 scripts/feature_selection.R
completed feature selection example

@@ -21,4 +21,27 @@ ggplot(d_tab, aes(x=cl, y=feature, fill=mean)) +
   geom_tile() +
   scale_fill_viridis_c(option="magma")
 
+summed_counts <- d %>%
+  group_by(cl) %>%
+  summarise(across(x1:x4, sum))
 
+# Calculate the row sums for each cluster
+col_sums <- colSums(select(summed_counts, x1:x4))
+
+# Divide each value in the columns x1 to x4 by its respective row sum
+summed_counts_normalized <- summed_counts %>%
+  mutate(across(x1:x4, ~ .x / col_sums))
+
+summed_counts_long <- summed_counts_normalized %>%
+  pivot_longer(cols = x1:x4, names_to = "feature", values_to = "mean")
+
+# Plot heatmap
+ggplot(summed_counts_long, aes(x=cl, y=feature, fill=mean)) +
+  geom_tile() +
+  scale_fill_viridis_c(option="magma")
+
+# In figure 1 (intra cluster fraction) we can see that all observations in
+# cluster 4 had a x4 value of 1.
+# Figure 2 (intra feature fraction) reveals that cluster 1 had more positive x4
+# values. In fact, cluster 1 has the largest accumulation of x4. We cannot see
+# this in figure 1, but I consider this important information.

### 2024-10-01 manuscript.tex
script for explaining the heatmaps
rewrite of normalization section

@@
 %% \AND can be used instead of \And and starts a new line
-\title{Interactively Visualizing Clustering in Multivariate Data Using lionfish}
+\title{Interactively Visualizing Multivariate Market Segmentation Using the R Package Lionfish}
 
@@
 
-\begin{figure*}[h]
+\begin{figure*}[ht]
 \centerline{\includegraphics[width=1\textwidth]{images/intro1.png}}
@@
 \end{figure*}
-
-One can see that the data is perfectly divided into four parts.
-However, if one were to plot the two features individually, this would not be obvious.
-Figure \ref{kmeans-histogram} shows histograms of the two features, V1, V2, with the colour matching the four partitions.
-From the histograms, we can see some differences in the partitions for the four different association structures, but they are all overlapping.
-The distinct border between the partitions can only be seen from the scatterplots of both features.
 
@@
 \end{figure*}
+
+One can see that the data is perfectly divided into four parts.
+However, if one were to plot the two features individually, this would not be obvious.
+Figure \ref{kmeans-histogram} shows histograms of the two features, V1, V2, with the colour matching the four partitions.
+From the histograms, we can see some differences in the partitions for the four different association structures, but they are all overlapping.
+The distinct border between the partitions can only be seen from the scatterplots of both features.
 
@@
 
-The aim of this work is to build an interface that allows for an interactive exploration of partitioning solutions of multivariate data. The interface should enable
+The aim of this work is to build an interface that allows for an interactive exploration of partitions generated by clustering of multivariate market data.
+In some combinations of the features, one should be able to the separations as can be seen in the 2D example in Figure \ref{kmeans-partition}.
+The objectives for the interface are to enable:
 
-\begin{itemize}
-\item visualizing a clustering solution with different tour methods, including manual changes to the viewing direction (manual tour)
-\item brushing of (groups of) points for manual refinement of the cluster solution (spin-and-brush capabilities)
-\item additional linked displays for comprehensive exploration of the solution
-\item ability to update the displays based on user selections
-\end{itemize}
+\begin{enumerate}
+\item Visualizing the partitions between clusters in combinations of features using tour methods, with both a grand tour and manual tour.
+\item Interactively brushing groups of points to refine the cluster solution, like done with spin-and-brush tools.
+\item Linking multiple displays to focus on particular clusters, and simplify the problem to better understand the clustering solution. 
+\item Update displays based on user selections such as feature selection or cluster selection, or re-scaling.
+\end{enumerate}
 
@@
 
-\subsection{Feature selection based on fractional analysis}
+\subsection{Feature and cluster relationships}
 
-A first step may look into what features are important for the cluster solution.
-The interface provides different capabilities that support feature selection.
-First, a 2D tour view can be used to understand the sensitivity of the grouping to individual features, in particular when using the manual tour to remove one feature at a time.
-This is in particular useful when the starting view was obtained through a projection pursuit guided tour~\citep{CBCH94} that optimized for the separation between the labeled groups~\citep{lckl2005}.
+A first step is to assess which features are important for the cluster solution. The interface provides different capabilities that support feature selection: tour view and summary heatmaps.
 
-In our applications we are considering market segmentation analysis on binary features.
-For this setting we have developed a heatmap display that we have found to be particularly useful in feature selection.
+The 2D tour view can be used to understand the sensitivity of the grouping to individual features.
+When using the manual tour to remove one feature at a time, the relationship between this variable and the partition between clusters can be examined.
+The ideal starting view is that obtained through a projection pursuit guided tour~\citep{CBCH94} that optimized for the separation between the labeled groups~\citep{lckl2005}.
+
+The summary heatmaps provide an overview of the cluster compositions relative to the features. Consider the matrix,
+
+\[
+C = \left[ \begin{array}{cccc} c_{11} & c_{12} & \dots & c_{1p} \\
+                                              c_{21} & c_{22} & \dots & c_{2p} \\
+                                              \vdots &  \vdots & \ddots &  \vdots \\
+                                              c_{k1} & c_{k2} & \dots & c_{kp}
+                                              \end{array} \right]
+\]
+
+where $c_{ij}, i=1, ..., p$ (number of features); $j=1, ..., k$ (number of clusters) are a summary of each feature in each cluster.
+Because our examples use binary features,  the $c_{ij}$ is the number of 1's of feature $j$ in cluster $i$.
+%For this setting we have developed a heatmap display that we have found to be particularly useful in feature selection.
 The idea is to understand differences in the binary ratings between clusters, where we use different normalizations to better explore the data.
 For \(c_{ij}\) the count (number of true values in the binary feature) in the \(i\)-th cluster and \(j\)-th feature, we define
+There are several ways that these values can be normalized to examine different aspects: $f_{ij}^{o} = \frac{c_{ij}}{n}$, $f_{ij}^{c} = \frac{c_{ij}}{n_{i}}$ and $f_{ij}^{f} = \frac{c_{ij}}{n_{j}}$.
+The first is the overall fraction, where counts are normalised by the overall number of observations $n$.
+It is generally not particularly useful.
+The second, $f^c$ is normalized relative to the size of each cluster, and is useful for comparing the features which contribute most to each cluster.
+The last, $f_f$ is normalized relative to each feature, so it is useful to examine how features are distributed across clusters \textcolor{red}{Not sure this last explanation is good - at least I don't understand what I've written}
 
-\[
-\text{Total fraction}\ f_{ij}^{tot} = \frac{c_{ij}}{n}
-\]
-with normalization to the total number of observations $n$ for a general overview,
+%\[
+%\text{Overall fraction}\ f_{ij}^{o} = \frac{c_{ij}}{n}
+%\]
+%with normalization to the total number of observations $n$ for a general overview,
 
-\[
-\text{Intra cluster fraction}\  f_{ij}^{clust} = \frac{c_{ij}}{n_{i}}
-\]
-showing the counts relative to the cluster's size \(n_i\), to see the fraction of observations in cluster $i$ for which feature $j$ was true, and
+%\[
+%\text{Intra cluster fraction}\  f_{ij}^{clust} = \frac{c_{ij}}{n_{i}}
+%\]
+%showing the counts relative to the cluster's size \(n_i\), to see the fraction of observations in cluster $i$ for which feature $j$ was true, and
 
-\[
-\text{Intra feature fraction}\ f_{ij}^{feat} = \frac{c_{ij}}{c_{j}}
-\]
-to see how the true values for feature $j$ are distributed across the different clusters relative, with normalization to the overall count of that feature \(c_j = \sum_i c_{ij}\).
-
-\textcolor{red}{different symbol for each? indicat on the index value somehow?}
+%\[
+%\text{Intra feature fraction}\ f_{ij}^{feat} = \frac{c_{ij}}{c_{j}}
+%\]
+%to see how the true values for feature $j$ are distributed across the different clusters relative, with normalization to the overall count of that feature \(c_j = \sum_i c_{ij}\).
 

### 2024-10-01 scripts/australian_tourism_clustering.R
added scripts for recreating all figures

@@ -1,0 +1,19 @@
+library(tourr)
+library(lionfish)
+library(flexclust)
+library(stats)
+library(dplyr)
+
+data("ausActiv")
+ausActiv <- ausActiv[rowSums(ausActiv) > 0 & rowSums(ausActiv) <= 40, ]
+set.seed(1234)
+
+# Figure 11
+dist_matrix_f <- dist(t(ausActiv), method="binary")
+ward_cluster_f <- hclust(dist_matrix_f, "ward.D2")
+plot(ward_cluster_f)
+
+clusters_f <- cutree(ward_cluster_f, k = 15)
+plot(ward_cluster_f, main = "Dendrogram of Australian Vacation Activities",
+     sub = "", xlab = "", ylab = "Height")
+rect_info <- rect.hclust(ward_cluster_f, k = 15, border = 2:16)

### 2024-10-01 scripts/austrian_tourism_clustering.R

@@ -1,0 +1,27 @@
+library(tourr)
+library(flexclust)
+library(lionfish)
+
+# perform initial k-means clustering
+set.seed(1234)
+data(winterActiv)
+clusters_full = stepcclust(winterActiv, k=6, nrep=20,  save.data=TRUE)
+
+# Figure 5
+init_env()
+obj1 <- list(type = "heatmap", obj = c("Intra Cluster Fraction"))
+interactive_tour(data=winterActiv,
+                 plot_objects = list(obj1),
+                 feature_names= colnames(winterActiv),
+                 preselection = clusters_full@cluster,
+                 n_subsets = 6)
+
+# Figure 6 A
+plot(Silhouette(clusters_full))
+
+# Figure 6 B
+winterActiv_features <- read.csv("saves/aut_saves/init/feature_selection.csv",
+                                 header = FALSE)
+winterActiv_feat_subset <- winterActiv[, colnames(winterActiv) %in% winterActiv_features$V1]
+clusters_feat_subset = stepcclust(winterActiv_feat_subset, k=6, nrep=20,  save.data=TRUE)
+plot(Silhouette(clusters_feat_subset))

### 2024-10-01 scripts/load_snapshots.R

@@ -1,0 +1,38 @@
+# install lionfish
+install.packages("remotes")
+remotes::install_github("mmedl94/lionfish")
+library(lionfish)
+library(dplyr)
+
+# set working directory
+setwd("..")
+
+# load Austrian Vacation Activities dataset
+data("winterActiv")
+winterActiv_features <- read.csv("saves/aut_saves/init/feature_selection.csv",
+                                 header = FALSE)
+winterActiv <- winterActiv[, colnames(winterActiv) %in% winterActiv_features$V1]
+
+
+# load Australian Vacation Activities dataset
+data("ausActiv")
+ausActiv <- ausActiv[rowSums(ausActiv) > 0 & rowSums(ausActiv) <= 40, ]
+ausActiv_features <- read.csv("saves/aus_saves/init/feature_selection.csv",
+                                 header = FALSE)
+ausActiv <- ausActiv[, colnames(ausActiv) %in% ausActiv_features$V1]
+ausActiv <- ausActiv[, ausActiv_features$V1]
+
+# Figure 7 + figure 8
+load_interactive_tour(winterActiv, "/saves/aut_saves/init")
+
+# Figure 9
+load_interactive_tour(winterActiv, "/saves/aut_saves/before")
+
+# Figure 10
+load_interactive_tour(winterActiv, "/saves/aut_saves/after")
+
+# Figure 12 + 13
+load_interactive_tour(ausActiv, "/saves/aus_saves/before")
+
+# Figure 14
+load_interactive_tour(ausActiv, "/saves/aus_saves/after")

### 2024-10-01 manuscript.tex
reworked figures 3, 7, 8, 9 and 10. minor adjustments to text. added scripts to load interfaces and recreate plots

@@
 The frame selection interface (Figure \ref{fig:GUI_overview}E) displays the frames of the currently shown tours and enables users to jump between frames of the tour objects.
-Below this, users can save and load projections and subsets using the respective buttons (Figure \ref{fig:GUI_overview}F).
-New tours can be started using the current settings via the interface at the bottom (Figure \ref{fig:GUI_overview}G), and users can select different metrics for some plot types e.g.
-heatmaps (Figure \ref{fig:GUI_overview}H).
+Below this, there are three additional interfaces (Figure \ref{fig:GUI_overview}F).
+The first can be used to adjust the number of bins in histograms—more bins result in higher resolution but slower display updates.
+The second allows users to animate tours, automatically shifting to the next projection after a specified amount of time.
+The third offers the ability to hide projection axes with a norm smaller than a chosen threshold, helping to reduce clutter.
+Users can also save and load projections and subsets using the respective buttons (Figure \ref{fig:GUI_overview}G).
+New tours can be started using the current settings via the interface at the bottom (Figure \ref{fig:GUI_overview}H), and users can select different metrics for some plot types e.g.
+heatmaps (Figure \ref{fig:GUI_overview}I).
 
@@
     \caption{Overview of the GUI.
-A: Sidebar controls; B: Display area; C: Feature selection checkboxes; D: Subset selection with color indicators; E: Frame selection interface; F: Save and load buttons; G: Interface for starting new tours; H: Metric selection interface.}
+A: Sidebar controls; B: Display area; C: Feature selection checkboxes; D: Subset selection with color indicators; E: Frame selection interface; F: Interfaces for adjusting the number of bins of histograms, animating tours and blending out projection axes with a low norm; G: Save and load buttons; H: Interface for starting new tours; I: Metric selection interface.}
     \label{fig:GUI_overview}
@@
 The first is the overall fraction, where counts are normalised by the overall number of observations $n$.
-It is generally not particularly useful.
+It gives a quick overview on the overall magnitude of the features.
 The second, $f^c$ is normalized relative to the size of each cluster, and is useful for comparing the features which contribute most to each cluster.
-The last, $f_f$ is normalized relative to each feature, so it is useful to examine how features are distributed across clusters \textcolor{red}{Not sure this last explanation is good - at least I don't understand what I've written}
+The last, $f^{f}$ is normalized relative to each feature, so it is useful to examine how features are distributed across clusters.
 
@@
 
-\textcolor{red}{It might be better to standardise these intra fractions.
-For A, one should divide by cluster count, and then compute activity fraction.
-For B, one should divide by activity count and then compute cluster fraction.} In Figure \ref{fig:winteractiv_heatmap}A, we can observe the general interests of tourists within each cluster.
+In Figure \ref{fig:winteractiv_heatmap}A, we can observe the general interests of tourists within each cluster.
 For instance, tourists in clusters 1, 3, 5, and 6 predominantly engaged in alpine skiing, while those in clusters 2 and 4 did not.
@@
     \caption{Display of a projection of the Austrian vacation activities dataset with the six clusters of the k-means cluster solution highlighted in different colors.
+Projection axes with a norm < 0.1 have been blend out to reduce cluttering.
 The colors indicate which clusters the highlighted observations belong to with cluster 1 being blue (A), cluster 2 being orange (B), cluster 3 being green (C), cluster 4 being red (D), cluster 5 being violet (E) and cluster 6 being brown (F).
-We can see that the projection separates some clusters well, however also that there is considerable overlap of clusters 4, 5 and 6}
+Some datapoints appear to be highlighted always, which occurs due to overlap of many datapoints in one spot.
+We can see that the projection separates some clusters well, however also that there is considerable overlap of clusters 4, 5 and 6.}
     \label{fig:winter_activ_cluster_highlights}
@@
     \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset.
+Projection axes with a norm < 0.1 have been blend out to reduce cluttering.
 Top left: 2D tour with the linear discriminant analysis projection pursuit index.
@@
     \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset, with manually adjusted projections.
+Projection axes with a norm < 0.1 have been blend out to reduce cluttering.
 Top left: 2D tour with the linear discriminant analysis projection pursuit index.
@@
     \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset, after new subsets have been selected.
+Projection axes with a norm < 0.1 have been blend out to reduce cluttering.
 Top left: 2D tour with the linear discriminant analysis projection pursuit index.

### 2024-10-01 scripts/load_snapshots.R

@@ -4,9 +4,20 @@ remotes::install_github("mmedl94/lionfish")
 library(lionfish)
 library(dplyr)
 
+library(tourr)
+library(devtools)
+library(reticulate)
+library(flexclust)
+library(stats)
+library(dplyr)
+setwd("/home/matthias/Desktop/Work/gsoc/lionfish")
+load_all()
+
+setwd("/home/matthias/Desktop/Work/gsoc/paper/pytour_paper")
 # set working directory
 setwd("..")
-
+setwd("/home/matthias/Desktop/Work/gsoc/paper/pytour_paper")
+init_env()
 # load Austrian Vacation Activities dataset
 data("winterActiv")
 winterActiv_features <- read.csv("saves/aut_saves/init/feature_selection.csv",
@@ -23,16 +34,21 @@ ausActiv <- ausActiv[, colnames(ausActiv) %in% ausActiv_features$V1]
 ausActiv <- ausActiv[, ausActiv_features$V1]
 
 # Figure 7 + figure 8
-load_interactive_tour(winterActiv, "/saves/aut_saves/init")
+load_interactive_tour(winterActiv, "/saves/aut_saves/init",
+                      display_size = 5)
 
 # Figure 9
-load_interactive_tour(winterActiv, "/saves/aut_saves/before")
+load_interactive_tour(winterActiv, "/saves/aut_saves/before",
+                      display_size = 5)
 
 # Figure 10
-load_interactive_tour(winterActiv, "/saves/aut_saves/after")
+load_interactive_tour(winterActiv, "/saves/aut_saves/after",
+                      display_size = 5)
 
 # Figure 12 + 13
-load_interactive_tour(ausActiv, "/saves/aus_saves/before")
+load_interactive_tour(ausActiv, "/saves/aus_saves/before",
+                      display_size = 5)
 
 # Figure 14
-load_interactive_tour(ausActiv, "/saves/aus_saves/after")
+load_interactive_tour(ausActiv, "/saves/aus_saves/after",
+                      display_size = 5)

### 2024-10-01 manuscript.tex
tidied up heatmap sections more

--- 2024-10-01 manuscript.tex
@@
 When using the manual tour to remove one feature at a time, the relationship between this variable and the partition between clusters can be examined.
-The ideal starting view is that obtained through a projection pursuit guided tour~\citep{CBCH94} that optimized for the separation between the labeled groups~\citep{lckl2005}.
+The ideal starting view is that obtained through a projection pursuit guided tour~\citep{CBCH94} that was optimized for the separation between the labeled groups~\citep{lckl2005}.
 
@@
 
-where $c_{ij}, i=1, ..., p$ (number of features); $j=1, ..., k$ (number of clusters) are a summary of each feature in each cluster.
+where $c_{ij}, i=1, ..., k$ (number of clusters); $j=1, ..., p$ (number of features) are a summary of each feature in each cluster.
 Because our examples use binary features,  the $c_{ij}$ is the number of 1's of feature $j$ in cluster $i$.
@@
 For \(c_{ij}\) the count (number of true values in the binary feature) in the \(i\)-th cluster and \(j\)-th feature, we define
-There are several ways that these values can be normalized to examine different aspects: $f_{ij}^{o} = \frac{c_{ij}}{n}$, $f_{ij}^{c} = \frac{c_{ij}}{n_{i}}$ and $f_{ij}^{f} = \frac{c_{ij}}{n_{j}}$.
-The first is the overall fraction, where counts are normalised by the overall number of observations $n$.
-It is generally not particularly useful.
-The second, $f^c$ is normalized relative to the size of each cluster, and is useful for comparing the features which contribute most to each cluster.
-The last, $f_f$ is normalized relative to each feature, so it is useful to examine how features are distributed across clusters \textcolor{red}{Not sure this last explanation is good - at least I don't understand what I've written}
+There are several ways that these values can be normalized to examine different aspects: $f_{ij}^{o} = \frac{c_{ij}}{n}$, $f_{ij}^{c} = \frac{c_{ij}}{n_{i}}$ and $f_{ij}^{f} = \frac{c_{ij}}{n_{j}}$, where $n_i, n_j$ are the row and column totals.
+The first, $f^{o}$ is the overall fraction, where counts are normalised by the overall number of observations $n$.
+Is generally less useful.
+
+The second, $f^c$ is normalized relative to the size of each cluster and can be considered the distribution of features in each cluster.
+It is useful for comparing the composition of each cluster relative to the features.
+For example, for $p=4$, if $f^c_{1j} = (0.4, 0.4, 0.1, 0.1)$ it suggests that high values (1's) of features 1 and 2 distinguish cluster 1.
+In the context of the later examples, this would mean cluster 1 contains tourists that list engaging in activities 1 and 2 but not 3 and 4.
+This normalisation produces what is called the {\bf intra-cluster fraction}.
+
+The last, $f_f$ is normalized relative to each feature, which can be considered to be the distribution of clusters on each feature.
+It is useful to examine how features are related to a cluster.
+Considering the last metric, for example, if $f^f_{i1} = (0.5, 0.4, 0, 0.1, 0)$ ($k=5$) would indicate that high values on feature 1 primarily are in clusters 1, 2.
+In the context of the later examples, this would mean activity 1 is commonly listed only in clusters 1 and 2.
+This normalisation produces what is called the {\bf intra-feature fraction}.
+
+Generally, if the features are ordered categories or numerical scores, decomposing the cluster-feature summary is still possible by using an analysis of variance breakdown into row and column effects.
 
@@
 Color represents (A) the intra-cluster fraction, and (B) the intra-feature fraction, with lighter indicating higher values.
-From A, we can see that some activities dominate most clusters, e.g.
-relaxing and going for walks.
-Some activities are unique to clusters, e.g.
-skiing is a primary activity for clusters 1, 3, 5, 6.
-From B, we can see which activities dominate each cluster, e.g.
-cluster 4 tourists like going to a spa and using health activities.}
+From A, we can see that cluster 3 tourists like alpine skiing, going out in the evening and going to discos and bars.
+They also like relaxing, shopping and sightseeing but these are popular among all tourists.
+From B, we can see the distribution of activities across clusters, e.g.
+using health facilities is popular in cluster 4 while going to a pool or sauna is popular in clusters 1 and 5.}
     \label{fig:winteractiv_heatmap}
@@
 
-\textcolor{red}{It might be better to standardise these intra fractions.
-For A, one should divide by cluster count, and then compute activity fraction.
-For B, one should divide by activity count and then compute cluster fraction.} In Figure \ref{fig:winteractiv_heatmap}A, we can observe the general interests of tourists within each cluster.
-For instance, tourists in clusters 1, 3, 5, and 6 predominantly engaged in alpine skiing, while those in clusters 2 and 4 did not.
+In Figure \ref{fig:winteractiv_heatmap}A, where colors show normalised column counts, we can observe the general interests of tourists within each cluster.
+Some activities are high on almost all clusters (e.g.
+relaxing, shopping), and some are low on all clusters (e.g.
+ski touring and horseback riding), and ignoring these can be helpful when assessing the distribution of activities between clusters.
+When comparing clusters 5 and 6 we can see that alpine skiing is high in both, but cluster 5 tourists also like going to the pool or sauna, but cluster 6 tourists prefer going for walks.
+%For instance, tourists in clusters 1, 3, 5, and 6 predominantly engaged in alpine skiing, while those in clusters 2 and 4 did not.
 Additionally, we see that activities such as ski touring and horseback riding were generally unpopular across all clusters.
 
-In Figure \ref{fig:winteractiv_heatmap}B, we can determine whether tourists in a particular cluster had a strong preference for certain activities.
+In Figure \ref{fig:winteractiv_heatmap}B, we can determine whether for a particular feature by tourists in which clusters.
 For example, nearly all tourists who visited museums are grouped in cluster 2, and those who used health facilities are primarily attributed to cluster 4.
-We can also identify activities that were similarly popular across all clusters, such as relaxing.
+Some activities, such as relaxing, are popular across all clusters.
 
-By using the heatmap with these metrics, we can perform feature selection by removing unpopular activities and those that were consistently similar across all clusters.
+These heatmaps can help with selecting features to focus on using the tour.
+Unpopular and universally popular activities can be removed.
 After performing the feature selection by unchecking the corresponding checkboxes in the GUI using this strategy, the following 12 activities remained: alpine skiing, going to a spa, using health facilities, hiking, going for walks, excursions, going out in the evening, going to discos/bars, shopping, sightseeing, museums, and pool/sauna.

### 2024-10-02 manuscript.tex

@@
 The first, $f^{o}$ is the overall fraction, where counts are normalised by the overall number of observations $n$.
-Is generally less useful.
+It gives a quick overview on the overall magnitude of the features.
 

### 2024-10-02 manuscript.tex
small changes and comments up to section 3

--- 2024-10-01 manuscript.tex
@@
 \item Visualizing the partitions between clusters in combinations of features using tour methods, with both a grand tour and manual tour.
-\item Interactively brushing groups of points to refine the cluster solution, like done with spin-and-brush tools.
+\item Interactively brushing groups of points to refine the cluster solution, like done with spin-and-brush tools~\citep{ggobi}.
 \item Linking multiple displays to focus on particular clusters, and simplify the problem to better understand the clustering solution. 
@@
 
-The interface was implemented in \texttt{lionfish} and offers a variety of linked interactive plot types, providing users with the flexibility to visualize their data from multiple perspectives.
+The interface was implemented in the R package \texttt{lionfish} and offers a variety of linked interactive plot types, providing users with the flexibility to visualize their data from multiple perspectives.
 The ability to navigate through various projections of the displayed tours directly within a graphical user interface (GUI) enables users to explore different aspects of the dataset.
@@
 These states can be recovered by using pressing the \texttt{Load projections and subsets} button.
-Additionally, users can initiate new tours of various directly from the sidebar.
+Additionally, users can initiate new tours directly from the sidebar.
+The options for new tour paths are: a local tour around the currently shown projection, a guided tour that searches for projections with maximum separation between groups \textcolor{red}{XXX is this true?
+other options?}
+
+\textcolor{red}{Mention additional interactive features in the displays - manual tour, anything else that might be relevant?}
 
@@
 Users can right-click and drag the arrowheads to modify the projections, after which the class orthonormalizes the projection axes and updates both the projection and the transformed data accordingly.
+
+\textcolor{red}{Maybe here you can say something about adding new display types, what would be needed for implementing them}
 
@@
 The first, $f^{o}$ is the overall fraction, where counts are normalised by the overall number of observations $n$.
-Is generally less useful.
+This can only give an overview and is generally less useful for feature selection.
 
@@
 This normalisation produces what is called the {\bf intra-cluster fraction}.
+\textcolor{red}{It corresponds to the cluster means.
+XXX why not larger values in the example, the maximum would be 1 so 0.4 still seems very moderate.}
 
-The last, $f_f$ is normalized relative to each feature, which can be considered to be the distribution of clusters on each feature.
+The last, $f^f$ is normalized relative to each feature, which can be considered to be the distribution of clusters on each feature.
 It is useful to examine how features are related to a cluster.
@@
 This normalisation produces what is called the {\bf intra-feature fraction}.
+\textcolor{red}{Note that this must be interpreted with care in case of inbalanced cluster sizes.}
 
@@
 A similar approach can be used to refine a partitioning solution.
-This is in particular useful to integrate prior knowledge or business interests in a given cluster solution.
+In a \textit{visual analytics} approach we use interactive visualizations to integrate human judgement with statistical and machine learning models~\citep{keim2010mastering} to optimize knowledge extraction from data.
+Here this is in particular useful to integrate prior knowledge or business interests in a given cluster solution.
 In the interface we can keep the provided clustering, but separate out new subsets via manual selection, for example after we found a group of particular interest via a manual tour.

### 2024-10-02 manuscript.tex
changed capitalization of references

--- 2024-10-02 manuscript.tex
@@
 \Plainauthor{Matthias Medl, Dianne Cook, Ursula Laa} %% comma-separated
-\Plaintitle{lionfish} %% without formatting
-\Shorttitle{lionfish} %% a short title (if necessary)
+\Plaintitle{Interactively Visualizing Multivariate Market Segmentation Using the R Package Lionfish} %% without formatting
+\Shorttitle{Market Segmentation using Lionfish} %% a short title (if necessary)
 
@@
 
-The majority of \texttt{lionfish} was written in Python, while the R side of the package handles setting up the Python environment within the interactive interface is being run, launching the interactive tour, and generating new tours when initiated through the interface.
+The majority of \texttt{lionfish} was written in Python \citep{python}, while the R \citep{R} side of the package handles setting up the Python environment within the interactive interface is being run, launching the interactive tour, and generating new tours when initiated through the interface.
 To incorporate the functionality of the \texttt{tourr} package without translating large portions of its code from R to Python, the \texttt{reticulate} package was used.

### 2024-10-02 manuscript.tex
response to Uschis comments. minor adjustments to 2.1 and the heatmap mini example

@@
 The \texttt{Reset original selection} button allows users to revert the subset selections to their initial state.
+All plots displayed in the GUI are linked, meaning that changes in one plot will affect the other plots.
+For instance, if observations are moved from one subset into another subset in one plot then the reassigment will also occur in the other plots.
 
@@
 Additionally, users can initiate new tours directly from the sidebar.
-The options for new tour paths are: a local tour around the currently shown projection, a guided tour that searches for projections with maximum separation between groups \textcolor{red}{XXX is this true?
-other options?}
+The options for new tour paths are: a local tour around the currently shown projection and guided tours that search for projections based on the holes- or the linear discriminant analysis (LDA) index.
+The holes index is sensitive to projections with few points in the center of the projections and the LDA index aims to maximize the distance between the centers of the selectet subsets.
+\textcolor{red}{XXX is this true?
+other options?} Additionally, all tour displays allow users to perform a manual tour; by right-clicking and dragging the arrowheads of projection axes, the projection is recalculated accordingly.
+This allows for manual exploration of the data.
 
@@
 \textcolor{red}{Maybe here you can say something about adding new display types, what would be needed for implementing them}
+\textcolor{blue}{I am unsure if this can easily be generalized.
+Some plots are easy to implement (scatterplot), but others like the mosaic plot come with a lot of complications, since that required a deep understanding on how matplotlib works.
+Implementing new interactive features would be especially challenging.}
 
@@
 It is useful for comparing the composition of each cluster relative to the features.
-For example, for $p=4$, if $f^c_{1j} = (0.4, 0.4, 0.1, 0.1)$ it suggests that high values (1's) of features 1 and 2 distinguish cluster 1.
-In the context of the later examples, this would mean cluster 1 contains tourists that list engaging in activities 1 and 2 but not 3 and 4.
+For example, for $p=4$, if $f^c_{1j} = (0.9, 0.2, 0.1, 0.1)$ it suggests that high values (1's) of feature 1 distinguishes cluster 1, and that values of the other  features were low.
+In the context of the later examples, this would mean cluster 1 contains tourists that especially engaged in activity 1, but not in activities 2, 3 and 4.
 This normalisation produces what is called the {\bf intra-cluster fraction}.
@@
 XXX why not larger values in the example, the maximum would be 1 so 0.4 still seems very moderate.}
+\textcolor{blue}{I adjusted the example a bit.}
 
@@
 It is useful to examine how features are related to a cluster.
-Considering the last metric, for example, if $f^f_{i1} = (0.5, 0.4, 0, 0.1, 0)$ ($k=5$) would indicate that high values on feature 1 primarily are in clusters 1, 2.
-In the context of the later examples, this would mean activity 1 is commonly listed only in clusters 1 and 2.
+Considering the last metric, for example, if $f^f_{i1} = (0.2, 0.7, 0, 0.1, 0)$ ($k=5$) would indicate that high values on feature 1 primarily are in cluster 2.
+In the context of the later examples, this would mean activity 1 is most commonly listed in cluster 2.
 This normalisation produces what is called the {\bf intra-feature fraction}.
 \textcolor{red}{Note that this must be interpreted with care in case of inbalanced cluster sizes.}
+
+In this example, while the intra-cluster fraction for cluster 1 and feature 1 is notably high at 0.9, the intra-feature fraction is comparatively low at 0.2.
+Applied to the subsequent analysis, this indicates that 90 \% of tourists in cluster 1 participated in activity 1, yet only 20 \% of all tourists who engaged in activity 1 were part of cluster 1.
+The majority of individuals participating in activity 1 were members of cluster 2.
+This discrepancy may be attributed to cluster 1 representing a relatively small group of tourists, characterized by a pronounced preference for activity 1 and a lack of interest in other activities.
 

### 2024-10-02 manuscript.tex
added abstract. minor addition to conclusion. marked proramming languages as such. italicize the k's of k means

@@
 \Abstract{
-Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum
-Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum
-Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum
-Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum
-Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum
-Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum
-Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum
-Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum Lorem ipsum
+Market segmentation presents a challenge in data science, particularly when dealing with high-dimensional data where regions of interest often overlap.
+In such cases, traditional feature selection and clustering algorithms can fall short in delivering both optimal and actionable solutions.
+The complexity of customer behavior, characterized by multiple and often overlapping traits, further complicates the identification of distinct market segments that can guide precise and targeted marketing strategies.
+Addressing these challenges requires a more nuanced and interactive approach to data exploration and analysis.
+
+In this study, we analyzed two datasets derived from surveys conducted in Austria and Australia, where tourists reported on their vacation activities.
+These datasets serve as examples of the challenges posed by market segmentation.
+To address these, we developed and employed the \texttt{lionfish} \proglang{R} package, which facilitates interactive and dynamic exploration of clustering solutions.
+Unlike static approaches, \texttt{lionfish} allows users to refine clusters iteratively, focusing on aspects of the data most relevant to their specific analytical objectives.
+The package integrates tour algorithms through the \texttt{tourr} package, enabling advanced visualization of high-dimensional data, while \proglang{Python}-powered interactivity allows for manual adjustments and redefinition of cluster boundaries based on visual feedback.
+
+By bridging the gap between automated algorithms and human intuition, \texttt{lionfish} enables analysts to surpass the limitations of traditional clustering methods.
+It offers a flexible, user-driven approach that reveals deeper insights into complex market behaviors, facilitating more effective segmentation and ultimately improving strategic decision-making.
 }
-\Keywords{interactive graphics, tourr, exploratory data analysis, \proglang{R}, \proglang{python}}
-\Plainkeywords{keywords, comma-separated, non-capitalized,, R} %% without formatting
+\Keywords{interactive graphics, tourr, exploratory data analysis, \proglang{R}, \proglang{Python}}
+\Plainkeywords{interactive graphics, tourr, exploratory data analysis, R, python} %% without formatting
 %% at least one keyword must be supplied
@@
 
-While tour animations are best obtained within R using the tourr package~\citep{tourr}, it does not enable the interactivity required for example for a manual tour~\citep{laa2023new}.
+While tour animations are best obtained within \proglang{R} using the tourr package~\citep{tourr}, it does not enable the interactivity required for example for a manual tour~\citep{laa2023new}.
 Interactive graphics are available when using Javascript, as implemented in the detourr package \citep{RJ-2023-052}.
@@
 For example, we may wish to explore the local neighbourhood of a projection selected by the user with a local tour animation provided by tourr, or we may want to optimize a guided tour path using groups identified via brushing.
-Our solution is using Python for high-performance interactivity, through the packages \texttt{TKinter} \citep{lundh1999introduction}, \texttt{CustomTKinter} \citep{schimansky24}, and \texttt{matplotlib} \citep{Hunter:2007}, with integration to the tourr package \citep{tourr} via \texttt{reticulate} \citep{reticulate}, a framework that facilitates seamless interoperability between Python and R.
+Our solution is using \proglang{Python} for high-performance interactivity, through the packages \texttt{TKinter} \citep{lundh1999introduction}, \texttt{CustomTKinter} \citep{schimansky24}, and \texttt{matplotlib} \citep{Hunter:2007}, with integration to the tourr package \citep{tourr} via \texttt{reticulate} \citep{reticulate}, a framework that facilitates seamless interoperability between \proglang{Python} and \proglang{R}.
 
@@
 
-The interface was implemented in the R package \texttt{lionfish} and offers a variety of linked interactive plot types, providing users with the flexibility to visualize their data from multiple perspectives.
+The interface was implemented in the \proglang{R} package \texttt{lionfish} and offers a variety of linked interactive plot types, providing users with the flexibility to visualize their data from multiple perspectives.
 The ability to navigate through various projections of the displayed tours directly within a graphical user interface (GUI) enables users to explore different aspects of the dataset.
@@
 The holes index is sensitive to projections with few points in the center of the projections and the LDA index aims to maximize the distance between the centers of the selectet subsets.
-\textcolor{red}{XXX is this true?
+\citep{ggobi} \textcolor{red}{XXX is this true?
 other options?} Additionally, all tour displays allow users to perform a manual tour; by right-clicking and dragging the arrowheads of projection axes, the projection is recalculated accordingly.
@@
 
-The majority of \texttt{lionfish} was written in Python \citep{python}, while the R \citep{R} side of the package handles setting up the Python environment within the interactive interface is being run, launching the interactive tour, and generating new tours when initiated through the interface.
-To incorporate the functionality of the \texttt{tourr} package without translating large portions of its code from R to Python, the \texttt{reticulate} package was used.
+The majority of \texttt{lionfish} was written in \proglang{Python} \citep{python}, while the \proglang{R} \citep{R} side of the package handles setting up the \proglang{Python} environment within the interactive interface is being run, launching the interactive tour, and generating new tours when initiated through the interface.
+To incorporate the functionality of the \texttt{tourr} package without translating large portions of its code from \proglang{R} to \proglang{Python}, the \texttt{reticulate} package was used.
 This approach allows \texttt{lionfish} to automatically benefit from updates to \texttt{tourr}.
@@
 However, to reduce inefficiencies caused by cross-language communication and to simplify debugging, \texttt{tourr} functions were only accessed when necessary.
-Core functionalities, such as performing data transformations using projections, were implemented directly in Python.
+Core functionalities, such as performing data transformations using projections, were implemented directly in \proglang{Python}.
 These transformations are based on well-established mathematical principles and are straightforward to replicate, ensuring they remain stable and efficient.
@@
 
-The central component of the Python code is the \texttt{customTKinter} class \texttt{InteractiveTourInterface}.
+The central component of the \proglang{Python} code is the \texttt{customTKinter} class \texttt{InteractiveTourInterface}.
 This class centrally stores attributes related to all plots, such as the dataset, sub-selections, feature selections, and other shared information.
-Plot-specific data is organized in dictionaries (the Python equivalent of named lists in R), including the display type, construction instructions, tour projections (only in case of tour displays), color schemes for the displayed data, and, where applicable, the selector and manual projection manipulation classes.
+Plot-specific data is organized in dictionaries (the \proglang{Python} equivalent of named lists in \proglang{R}), including the display type, construction instructions, tour projections (only in case of tour displays), color schemes for the displayed data, and, where applicable, the selector and manual projection manipulation classes.
 
@@
 In this example, while the intra-cluster fraction for cluster 1 and feature 1 is notably high at 0.9, the intra-feature fraction is comparatively low at 0.2.
-Applied to the subsequent analysis, this indicates that 90 \% of tourists in cluster 1 participated in activity 1, yet only 20 \% of all tourists who engaged in activity 1 were part of cluster 1.
+Applied to the subsequent analysis, this indicates that 90\% of tourists in cluster 1 participated in activity 1, yet only 20\% of all tourists who engaged in activity 1 were part of cluster 1.
 The majority of individuals participating in activity 1 were members of cluster 2.
@@
 
-To gain further insight into the dataset a k-means clustering as described in \cite{leisch2018market} has been performed.
-Therefore, the function \texttt{stepcclust} of the R package \texttt{flexclust}\citep{flexclust} with $k=6$ and $nrep=20$ was used.
+To gain further insight into the dataset a $k$-means clustering as described in \cite{leisch2018market} has been performed.
+Therefore, the function \texttt{stepcclust} of the \proglang{R} package \texttt{flexclust}\citep{flexclust} with $k=6$ and $nrep=20$ was used.
 
@@
 
-We can now repeat the k-means clustering with \texttt{stepcclust} on the reduced dataset.
+We can now repeat the $k$-means clustering with \texttt{stepcclust} on the reduced dataset.
 Silhouette plots of both cluster solutions can be seen in Figure \ref{fig:silhouette_comparison}.
@@
     \includegraphics[width=1\textwidth]{silhouette_comparison.png}
-    \caption{Comparison of silhouette plots of two k-means cluster solutions of the Austrian vacation activities dataset with k=6.
-(A) shows the silhouette plot of the k-means solution of the full dataset and (B) the silhouette plot of the k-means solution of the dataset after manual feature selection.
+    \caption{Comparison of silhouette plots of two $k$-means cluster solutions of the Austrian vacation activities dataset with $k$=6.
+(A) shows the silhouette plot of the $k$-means solution of the full dataset and (B) the silhouette plot of the $k$-means solution of the dataset after manual feature selection.
 We can see that the cluster solution with the reduced dataset achieved better silhouette scores and that clusters 1 and 3 contain observations with negative silhouette scores.}
@@
     \includegraphics[width=1\textwidth]{winter_activ_cluster_highlights.png}
-    \caption{Display of a projection of the Austrian vacation activities dataset with the six clusters of the k-means cluster solution highlighted in different colors.
+    \caption{Display of a projection of the Austrian vacation activities dataset with the six clusters of the $k$-means cluster solution highlighted in different colors.
 Projection axes with a norm < 0.1 have been blend out to reduce cluttering.
@@
     \includegraphics[width=1\textwidth]{winter_cl7_init.png}
-    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset.
+    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the $k$-means solution of the Austrian vacation activities dataset.
 Projection axes with a norm < 0.1 have been blend out to reduce cluttering.
@@
     \includegraphics[width=1\textwidth]{winter_cl7_pre.png}
-    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset, with manually adjusted projections.
+    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the $k$-means solution of the Austrian vacation activities dataset, with manually adjusted projections.
 Projection axes with a norm < 0.1 have been blend out to reduce cluttering.
@@
     \includegraphics[width=1\textwidth]{winter_cl7_post.png}
-    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset, after new subsets have been selected.
+    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the $k$-means solution of the Austrian vacation activities dataset, after new subsets have been selected.
 Projection axes with a norm < 0.1 have been blend out to reduce cluttering.
@@
 Multiple features from this cluster were retained to preserve as much information as possible.
-After feature selection, the observations were reclustered using k-means with \( k = 15 \).
+After feature selection, the observations were reclustered using $k$-means with \( k = 15 \).
 
@@
     \includegraphics[width=1\textwidth]{aus_heatmap.png}
-    \caption{Heatmap of the intra cluster fraction (indicated by color) of the k-means solution with \( k = 6 \) and the reduced feature subset.
+    \caption{Heatmap of the intra cluster fraction (indicated by color) of the $k$-means solution with \( k = 6 \) and the reduced feature subset.
 We can observe that tourists in subsets 1, 2 and 6 all prefer to travel without their friends.}
@@
     \includegraphics[width=1\textwidth]{aus_preselection.png}
-    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Australian vacation activities dataset, with the projection axes of the feature ``Friends'' pointing into one direction and all the other ones into another one.
+    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the $k$-means solution of the Australian vacation activities dataset, with the projection axes of the feature ``Friends'' pointing into one direction and all the other ones into another one.
 Top left: 2D tour.
@@
 Future developments in the software might include expanding the range of interactive features and providing additional visualization methods.
-Because it is built on R and python, it would also be possible to integrate more closely with the machine learning algorithms, such as generating new cluster solutions directly from the GUI.
+Because it is built on \proglang{R} and \proglang{Python}, it would also be possible to integrate more closely with the machine learning algorithms, such as generating new cluster solutions directly from the GUI.
 For more complex shapes, containing concavities, or non-linear boundaries, implementation of sliced tours~\citep{Laa2020}, would be recommended.
+Furthermore, it provides an additional demonstration of how \proglang{R} and \proglang{Python} can be integrated to harness the strengths of both programming languages.
 
 In summary, \texttt{lionfish} represents a significant advancement in the toolkit of data analysts, offering a novel way to balance automated analysis with human intuition and domain expertise, thereby facilitating a deeper and more comprehensive understanding of complex datasets.
+
+\section{Acknowledgements}~\label{acknowledgements}
+
+The development of the \texttt{lionfish} package has been supported by the Google Summer of Code 2024 program.
 

### 2024-10-02 manuscript.tex
changed the capitalisation of the plot indices of the intro plots to be uniform with the rest of the figures. More making of programming languages and packages

@@
 \begin{figure*}[ht]
-\centerline{\includegraphics[width=1\textwidth]{images/intro1.png}}
-\caption{Examples of how $k$-means partitions 2D data with different association structure. If the correlation is high (c), the partitioning happens along the primary direction of the association.}
+\centerline{\includegraphics[width=1\textwidth]{images/intro1.pdf}}
+\caption{Examples of how $k$-means partitions 2D data with different association structure. If the correlation is high (C), the partitioning happens along the primary direction of the association.}
 \label{kmeans-partition}
@@
 \begin{figure*}[ht]
-\centerline{\includegraphics[width=1\textwidth]{images/intro2.png}}
+\centerline{\includegraphics[width=1\textwidth]{images/intro2.pdf}}
 \caption{The four $k$-means partitions plotted as histograms of the individual features.
@@
 While tour animations are best obtained within \proglang{R} using the tourr package~\citep{tourr}, it does not enable the interactivity required for example for a manual tour~\citep{laa2023new}.
-Interactive graphics are available when using Javascript, as implemented in the detourr package \citep{RJ-2023-052}.
+Interactive graphics are available when using \proglang{Javascript}, as implemented in the \texttt{detourr} package \citep{RJ-2023-052}.
 This allows to replay a recorded tour path with interactive graphics, and can also be linked with additional displays, but lacks capabilities for manual tours.
 
-To integrate user interactions with the capabilities of the tourr package an active communication with the interface is required.
+To integrate user interactions with the capabilities of the \texttt{tourr package} an active communication with the interface is required.
 For example, we may wish to explore the local neighbourhood of a projection selected by the user with a local tour animation provided by tourr, or we may want to optimize a guided tour path using groups identified via brushing.
@@
 
-We can further explore the similarities between the clusters by initializing an \texttt{interactive\_tour()} with a 2D tour based on the linear discriminant analysis (LDA) projection pursuit index.
+We can further explore the similarities between the clusters by initializing an \texttt{interactive\_tour()} with a 2D tour based on the LDA projection pursuit index.
 By navigating through the tour, we can observe various projections, and when a projection that separates the clusters is found, we highlight each cluster sequentially.

### 2024-10-02 manuscript.tex
Small changes in the Applications section

--- 2024-10-02 manuscript.tex
@@
 
-The Austrian Vacation Activities \cite{dolnicar2003winter} and the Australian Vacation Activities \cite{cliff2009formative} datasets are used to illustrate the methods.
+The Austrian Vacation Activities \citep{dolnicar2003winter} and the Australian Vacation Activities \cite{cliff2009formative} datasets are used to illustrate the methods.
 
@@
 From B, we can see the distribution of activities across clusters, e.g.
-using health facilities is popular in cluster 4 while going to a pool or sauna is popular in clusters 1 and 5.}
+most tourists who use health facilities are found in cluster 4 while tourist going to a pool or sauna are primarily found in clusters 1 and 5.}
     \label{fig:winteractiv_heatmap}
@@
 
-In Figure \ref{fig:winteractiv_heatmap}B, we can determine whether for a particular feature by tourists in which clusters.
+In Figure \ref{fig:winteractiv_heatmap}B, we can determine whether tourists selecting a particular feature are equally distributed or if they primarily fall within one or a few clusters.
 For example, nearly all tourists who visited museums are grouped in cluster 2, and those who used health facilities are primarily attributed to cluster 4.
@@
 By navigating through the tour, we can observe various projections, and when a projection that separates the clusters is found, we highlight each cluster sequentially.
-The different highlighted clusters can be seen in Figure \ref{fig:silhouette_comparison}.
+The different highlighted clusters can be seen in Figure \ref{fig:winter_activ_cluster_highlights}.
 
@@
     \caption{Display of a projection of the Austrian vacation activities dataset with the six clusters of the k-means cluster solution highlighted in different colors.
-Projection axes with a norm < 0.1 have been blend out to reduce cluttering.
+Projection axes with a norm < 0.1 are not shown to reduce cluttering.
 The colors indicate which clusters the highlighted observations belong to with cluster 1 being blue (A), cluster 2 being orange (B), cluster 3 being green (C), cluster 4 being red (D), cluster 5 being violet (E) and cluster 6 being brown (F).
@@
 The most influential features shown in Figure \ref{fig:winter_activ_cluster_highlights} are pool/sauna, alpine skiing, museums, and going to the spa.
-The projection roughly separates clusters 1 (blue), 2 (orange), and 3 (green) from each other and the other three clusters (red, violet and brown), which appear to be quite similar.
+The projection roughly separates clusters 1 (blue), 2 (orange), and 3 (green) from each other and the other three clusters (red, violet and brown), which appear to be quite similar in the selected projection.
 
@@
     \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset.
-Projection axes with a norm < 0.1 have been blend out to reduce cluttering.
+Projection axes with a norm < 0.1 are not shown to reduce cluttering.
 Top left: 2D tour with the linear discriminant analysis projection pursuit index.
@@
 Multiple features from this cluster were retained to preserve as much information as possible.
-After feature selection, the observations were reclustered using k-means with \( k = 15 \).
+After feature selection, the observations were clustered using k-means with \( k = 6 \).
 
@@
     \includegraphics[width=1\textwidth]{aus_feature_clustering.png}
-    \caption{Dendogram of the features of the Australian Vacation Activities dataset using Ward's method with the Jaccard index.
+    \caption{Dendrogram of the features of the Australian Vacation Activities dataset using Ward's method with the Jaccard index.
 Features which were clustered together are marked by the colored boxes.

### 2024-10-02 manuscript.tex

--- 2024-10-02 manuscript.tex
@@
 
-We can further explore the similarities between the clusters by initializing an \texttt{interactive\_tour()} with a 2D tour based on the LDA projection pursuit index.
+We can further explore the similarities between the clusters by initializing an \texttt{interactive\_tour()} with a 2D tour based on the linear discriminant analysis (LDA) projection pursuit index.
 By navigating through the tour, we can observe various projections, and when a projection that separates the clusters is found, we highlight each cluster sequentially.
-The different highlighted clusters can be seen in Figure \ref{fig:silhouette_comparison}.
+The different highlighted clusters can be seen in Figure \ref{fig:winter_activ_cluster_highlights}.
 
@@
     \includegraphics[width=1\textwidth]{winter_activ_cluster_highlights.png}
-    \caption{Display of a projection of the Austrian vacation activities dataset with the six clusters of the $k$-means cluster solution highlighted in different colors.
-Projection axes with a norm < 0.1 have been blend out to reduce cluttering.
+    \caption{Display of a projection of the Austrian vacation activities dataset with the six clusters of the k-means cluster solution highlighted in different colors.
+Projection axes with a norm < 0.1 are not shown to reduce cluttering.
 The colors indicate which clusters the highlighted observations belong to with cluster 1 being blue (A), cluster 2 being orange (B), cluster 3 being green (C), cluster 4 being red (D), cluster 5 being violet (E) and cluster 6 being brown (F).
@@
     \includegraphics[width=1\textwidth]{winter_cl7_init.png}
-    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the $k$-means solution of the Austrian vacation activities dataset.
-Projection axes with a norm < 0.1 have been blend out to reduce cluttering.
+    \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset.
+Projection axes with a norm < 0.1 are not shown to reduce cluttering.
 Top left: 2D tour with the linear discriminant analysis projection pursuit index.
@@
 Multiple features from this cluster were retained to preserve as much information as possible.
-After feature selection, the observations were reclustered using $k$-means with \( k = 15 \).
+After feature selection, the observations were clustered using k-means with \( k = 6 \).
 

### 2024-10-02 manuscript.tex
Shorter abstract

@@
 \Abstract{
-Market segmentation presents a challenge in data science, particularly when dealing with high-dimensional data where regions of interest often overlap.
+%Market segmentation presents a challenge in data science, particularly when dealing with high-dimensional data where regions of interest often overlap.
 In such cases, traditional feature selection and clustering algorithms can fall short in delivering both optimal and actionable solutions.
@@
 
-In this study, we analyzed two datasets derived from surveys conducted in Austria and Australia, where tourists reported on their vacation activities.
+Market segmentation typically partitions data that does not have any clear cluster structure, resulting in groups of consumers who are similar, but without clear separation.
+The interpretation of such a clustering solution is typically obscured by the multivariate nature of the consumer data.
+The \texttt{lionfish} \proglang{R} package facilitates an interactive and dynamic exploration of clustering solutions.
+Unlike static approaches, \texttt{lionfish} allows users to refine clusters iteratively, focusing on aspects of the data most relevant to their specific analytical objectives.
+The package integrates tour algorithms through the \texttt{tourr} package, enabling multivariate visualization of high-dimensional data, while \proglang{Python}-powered interactivity allows for manual adjustments and redefinition of cluster boundaries based on visual feedback.
+We analyzed two datasets derived from surveys conducted in Austria and Australia, where tourists reported on their vacation activities and show how we can understand and enhance a clustering solution using \texttt{lionfish} in those examples.
+The flexible, user-driven approach provides deeper insights into the complex market behaviours, facilitating more effective segmentation and improving strategic decision-making.
+
+%In this study, we analyzed two datasets derived from surveys conducted in Austria and Australia, where tourists reported on their vacation activities.
 These datasets serve as examples of the challenges posed by market segmentation.
@@
 
-By bridging the gap between automated algorithms and human intuition, \texttt{lionfish} enables analysts to surpass the limitations of traditional clustering methods.
+%By bridging the gap between automated algorithms and human intuition, \texttt{lionfish} enables analysts to surpass the limitations of traditional clustering methods.
 It offers a flexible, user-driven approach that reveals deeper insights into complex market behaviors, facilitating more effective segmentation and ultimately improving strategic decision-making.
 }
-\Keywords{interactive graphics, tourr, exploratory data analysis, \proglang{R}, \proglang{Python}}
+\Keywords{interactive graphics, tourr, exploratory data analysis, \proglang{R}, \proglang{Python}, clustering, market segmentation}
 \Plainkeywords{interactive graphics, tourr, exploratory data analysis, R, python} %% without formatting

### 2024-10-02 manuscript.tex
a few minor adjustments. Expanded the Australian tourism applicaiton section to consider couples and families.

--- 2024-10-02 manuscript.tex
@@
 
-While tour animations are best obtained within \proglang{R} using the tourr package~\citep{tourr}, it does not enable the interactivity required for example for a manual tour~\citep{laa2023new}.
+While tour animations are best obtained within \proglang{R} using the \texttt{tourr} package~\citep{tourr}, it does not enable the interactivity required for example for a manual tour~\citep{laa2023new}.
 Interactive graphics are available when using \proglang{Javascript}, as implemented in the \texttt{detourr} package \citep{RJ-2023-052}.
@@
 
-To integrate user interactions with the capabilities of the \texttt{tourr package} an active communication with the interface is required.
-For example, we may wish to explore the local neighbourhood of a projection selected by the user with a local tour animation provided by tourr, or we may want to optimize a guided tour path using groups identified via brushing.
-Our solution is using \proglang{Python} for high-performance interactivity, through the packages \texttt{TKinter} \citep{lundh1999introduction}, \texttt{CustomTKinter} \citep{schimansky24}, and \texttt{matplotlib} \citep{Hunter:2007}, with integration to the tourr package \citep{tourr} via \texttt{reticulate} \citep{reticulate}, a framework that facilitates seamless interoperability between \proglang{Python} and \proglang{R}.
+To integrate user interactions with the capabilities of the \texttt{tourr} package an active communication with the interface is required.
+For example, we may wish to explore the local neighbourhood of a projection selected by the user with a local tour animation provided by \texttt{tourr}, or we may want to optimize a guided tour path using groups identified via brushing.
+Our solution is using \proglang{Python} for high-performance interactivity, through the packages \texttt{TKinter} \citep{lundh1999introduction}, \texttt{CustomTKinter} \citep{schimansky24}, and \texttt{matplotlib} \citep{Hunter:2007}, with integration to the \texttt{tourr} package \citep{tourr} via \texttt{reticulate} \citep{reticulate}, a framework that facilitates seamless interoperability between \proglang{Python} and \proglang{R}.
 
@@
 At a minimum, the user needs to provide both the dataset and the instructions for constructing the desired plots.
-The dataset must be supplied as a \texttt{data.table}, while the plotting instructions should be passed as a list containing the named elements \texttt{type} and \texttt{obj}.
+The dataset must be supplied as a numeric matrix, while the plotting instructions should be passed as a list containing the named elements \texttt{type} and \texttt{obj}.
 The \texttt{type} element specifies the type of display to generate, such as \texttt{scatter} for a scatterplot or \texttt{2d\textunderscore tour} for a 2-dimensional tour.
@@
 The options for new tour paths are: a local tour around the currently shown projection and guided tours that search for projections based on the holes- or the linear discriminant analysis (LDA) index.
-The holes index is sensitive to projections with few points in the center of the projections and the LDA index aims to maximize the distance between the centers of the selectet subsets.
-\citep{ggobi} \textcolor{red}{XXX is this true?
-other options?} Additionally, all tour displays allow users to perform a manual tour; by right-clicking and dragging the arrowheads of projection axes, the projection is recalculated accordingly.
+The holes index is sensitive to projections with few points in the center of the projections and the LDA index aims to maximize the distance between the centers of the selectet subsets \citep{ggobi}.
+Additionally, all tour displays allow users to perform a manual tour; by right-clicking and dragging the arrowheads of projection axes, the projection is recalculated accordingly.
 This allows for manual exploration of the data.
 
-\textcolor{red}{Mention additional interactive features in the displays - manual tour, anything else that might be relevant?}
+More information on all the features of the GUI can be found on \href{https://github.com/mmedl94/lionfish}{https://github.com/mmedl94/lionfish} and \href{https://mmedl94.github.io/lionfish/}{https://mmedl94.github.io/lionfish/}.
 
@@
 Users can right-click and drag the arrowheads to modify the projections, after which the class orthonormalizes the projection axes and updates both the projection and the transformed data accordingly.
-
-\textcolor{red}{Maybe here you can say something about adding new display types, what would be needed for implementing them}
-\textcolor{blue}{I am unsure if this can easily be generalized.
-Some plots are easy to implement (scatterplot), but others like the mosaic plot come with a lot of complications, since that required a deep understanding on how matplotlib works.
-Implementing new interactive features would be especially challenging.}
 
@@
 This normalisation produces what is called the {\bf intra-cluster fraction}.
-\textcolor{red}{It corresponds to the cluster means.
-XXX why not larger values in the example, the maximum would be 1 so 0.4 still seems very moderate.}
-\textcolor{blue}{I adjusted the example a bit.}
+In case of binary features it equates to the cluster means.
 
@@
 This normalisation produces what is called the {\bf intra-feature fraction}.
-\textcolor{red}{Note that this must be interpreted with care in case of inbalanced cluster sizes.}
+It also has to be noted that this metric is heavily influenced by the cluster sizes.
+Consider a feature that is uniformly distributed across all clusters, then the largest cluster would receive the highest intra-feature fraction.
 
@@
 Projection axes with a norm < 0.1 have been blend out to reduce cluttering.
-Top left: 2D tour with the linear discriminant analysis projection pursuit index.
+Top left: 2D tour with a manually adjusted projection.
 Top right: heatmap with the intra-cluster fraction.
-Bottom left: 1D tour with the linear discriminant analysis projection pursuit index.
+Bottom left: 1D tour with a manually adjusted projection.
 Bottom right: mosaic plot.
@@
 
-\subsubsection{Segmentation of solo travellers}
+\subsubsection{Segmentation of travellers travelling without friends}
 
-In the heatmap displaying the intra-cluster fraction shown in Figure \ref{fig:aus_heatmapg}, we observe that subsets 1, 2, and 6 tend to prefer travelling alone.
-As a tourist agency, we might be interested in targeting solo travellers more effectively.
+In the heatmap displaying the intra-cluster fraction shown in Figure \ref{fig:aus_heatmapg}, we observe that subsets 1, 2, and 6 tend to prefer travelling without friends.
+We can assume that this group consits of tourists travelling alone as couples or with their family.
+Additional features to further subsegment these groups might be interesting for future surveys.
+As a tourist agency, we might be interested in targeting these travellers more effectively.
 However, subsets 2 and 6 also include individuals who enjoy spending time with their friends.
@@
 Since we are particularly interested in tourists who did not spend time with friends, we can extend the projection axis ``Friends'' outward to separate the data based on that feature.
-Datapoints in the opposite direction of the ``Friends'' axes are the solo travellers we are interested in, which can be seen in Figure \ref{fig:aus_preselection} on the left side of the top left plot.
+Datapoints in the opposite direction of the ``Friends'' axes are the tourists we are interested in, which can be seen in Figure \ref{fig:aus_preselection} on the left side of the top left plot.
 Subsequently, we can pull all other projection axes in one direction to separate data points based on their overall activity level.
@@
 
-By comparing subset 3 (green), which contains active tourists who spent time with their friends, with subset 7 (pink), the active tourists who traveled alone, we notice that subset 7 showed less interest in visiting the casino, theatre, and chartering a boat.
+By comparing subset 3 (green), which contains active tourists who spent time with their friends, with subset 7 (pink), the active tourists who traveled without their friends, we notice that subset 7 showed less interest in visiting the casino, theatre, and chartering a boat.
 
-We also observe distinctive differences in the interests of the subsets that traveled alone.
+We also observe distinctive differences in the interests of the subsets that traveled without friends.
 Subset 7 was interested in relaxing, shopping, sightseeing, wildlife, going to the beach, visiting pubs, and exploring farms.
-Given that subset 7 showed a notable interest in going to pubs, we can assume they want to meet new people.
-This insight can be utilized in a marketing campaign that bundles the activities subset 7 is interested in, creating a package tailored for solo travellers.
-This way, they can meet other solo travellers while engaging in activities they enjoy.
+Given that Subset 7 showed a notable interest in going to pubs, we can assume that the solo travelers and couples in this subset are looking to meet new people.
+This insight could be leveraged in a marketing campaign by bundling activities that appeal to Subset 7, creating packages tailored to these tourists.
+Such packages would provide opportunities for them to connect with other travelers who share similar interests while enjoying their preferred activities.
+Additionally, some solo travelers and couples might prefer experiences or environments without children, allowing for the creation of adult-oriented packages.
+The remaining subgroup within Subset 7 consists of families.
+To better cater to this group, packages can be refined to consider that not all museums or festivals are equally suitable for children, enabling more family-friendly customizations.
 
@@
 For them, the focus was on sightseeing, relaxing, shopping, and going to the beach, with much less interest in other activities.
-It can be assumed that they value their time alone.
-These tourists could potentially be targeted more effectively by offering sightseeing options with minimal interaction, such as using a phone app to provide information about interesting locations, rather than relying on a tour guide.
+It can be assumed that solo travellers in this group value their time alone.
+This subgroup could potentially be targeted more effectively by offering sightseeing options with minimal interaction, such as using a phone app to provide information about interesting locations, rather than relying on a tour guide.
+For couples in this subset, the focus may be on enjoying quality time together in a relaxed, low-key environment.
+Marketing strategies could emphasize romantic sightseeing experiences, such as sunset tours or private beach spots.
+Additionally, offering couples' packages that include spa treatments, leisurely dining experiences, and personalized itineraries could resonate well with this group, allowing them to unwind and connect at their own pace.
+Families in subset 8 might prioritize activities that balance relaxation and light exploration, particularly in child-friendly settings.
+Sightseeing tours tailored to families, with engaging and educational content for children, could be a strong fit.
+Beach outings that offer safe, family-oriented areas or interactive experiences like sandcastle-building competitions could also be appealing.
+By curating packages that cater to both relaxation and gentle family activities, families can enjoy their vacation with minimal stress.
 
 Finally, since subset 9 (gold) is notably inactive, one might infer that they prefer spending much of their time in their accommodation.
-Consequently, these individuals might be most interested in accommodations that offer well-equipped, comfortable living spaces with amenities that cater to relaxation and leisure.
+Consequently, these tourists might be most interested in accommodations that offer well-equipped, comfortable living spaces with amenities that cater to relaxation and leisure.
 
@@
 
-Similarly, in the Australian dataset, manual exploration allowed us to refine the understanding of solo travellers by dividing them into three distinct subsegments: highly active, moderately active, and largely inactive tourists.
+Similarly, in the Australian dataset, manual exploration allowed us to refine the understanding of tourists travelling without friends by dividing them into three distinct subsegments: highly active, moderately active, and largely inactive tourists.
 This nuanced segmentation, derived from a blend of automated clustering and manual adjustments, offers a deeper understanding of the varied needs and behaviors within this group, enabling more effective marketing strategies.
@@
 For more complex shapes, containing concavities, or non-linear boundaries, implementation of sliced tours~\citep{Laa2020}, would be recommended.
-Furthermore, it provides an additional demonstration of how \proglang{R} and \proglang{Python} can be integrated to harness the strengths of both programming languages.
+The integration of new plot types could also be streamlined to improve the versitility of the package.
+
+Furthermore, \texttt{lionfish} provides an additional demonstration of how \proglang{R} and \proglang{Python} can be integrated to harness the strengths of both programming languages.
 

### 2024-10-03 manuscript.tex
lots of minor tweaks

@@
 For a scatterplot, the user needs to provide a vector of strings specifying the names of the features to be displayed.
-The user can optionally specify the feature names, the arrangement of the plots, predefined subsets of the data (e.g., cluster solutions), custom names for these subsets, the number of available subsets, and the size of each plot.
+With the optional arguments of \texttt{interactive\textunderscore tour()} the user can specify the feature names, the arrangement of the plots within the GUI, predefined subsets of the data (e.g., cluster solutions), custom names for these subsets, the number of available subsets, and the size of each plot.
 
@@
 Users can also save and load projections and subsets using the respective buttons (Figure \ref{fig:GUI_overview}G).
-New tours can be started using the current settings via the interface at the bottom (Figure \ref{fig:GUI_overview}H), and users can select different metrics for some plot types e.g.
+New tours can also be started directly from within the GUI (Figure \ref{fig:GUI_overview}H), and users can select different metrics for some plot types e.g.
 heatmaps (Figure \ref{fig:GUI_overview}I).
@@
 
-One can move through the projections of a tour by pressing the arrow keys or by specifying the index of the projection the be displayed and pressing the \texttt{Update frames} button.
+One can move through the projections of a tour by pressing the arrow keys or by specifying the frame of the projection to be displayed and pressing the \texttt{Update frames} button.
 Users can animate the tours by toggling the \texttt{Animate} checkbox and specifying a time interval, so that the displays automatically move to the next projections after the specified time interval.
@@
 
-More information on all the features of the GUI can be found on \href{https://github.com/mmedl94/lionfish}{https://github.com/mmedl94/lionfish} and \href{https://mmedl94.github.io/lionfish/}{https://mmedl94.github.io/lionfish/}.
+More information on all the features of the GUI can be found on \href{https://github.com/mmedl94/lionfish}{https://github.com/mmedl94/lionfish} and on \href{https://mmedl94.github.io/lionfish/}{https://mmedl94.github.io/lionfish/}.
 
@@
 
-The majority of \texttt{lionfish} was written in \proglang{Python} \citep{python}, while the \proglang{R} \citep{R} side of the package handles setting up the \proglang{Python} environment within the interactive interface is being run, launching the interactive tour, and generating new tours when initiated through the interface.
+The majority of \texttt{lionfish} was written in \proglang{Python} \citep{python}.
+The \proglang{R} \citep{R} side of the package handles setting up the \proglang{Python} environment within the interactive interface is being run, launching the interactive tour, and generating new tours when initiated through the interface.
 To incorporate the functionality of the \texttt{tourr} package without translating large portions of its code from \proglang{R} to \proglang{Python}, the \texttt{reticulate} package was used.
-This approach allows \texttt{lionfish} to automatically benefit from updates to \texttt{tourr}.
+This approach allows \texttt{lionfish} to automatically benefit from updates to \texttt{tourr} and avoided the necessity to reimplement code in \proglang{Python} that was already written in \proglang{R}.
 
@@
 
-Before launching the \texttt{lionfish} interface the user should have performed a partitioning of their choice, and provide the initial clustering solution used by the interface.
+Before launching the \texttt{lionfish} interface, the user should have performed a partitioning of their choice, and provide the initial clustering solution used by the interface.
 We launch the interface to explore and potentially refine this solution.
@@
 It also has to be noted that this metric is heavily influenced by the cluster sizes.
-Consider a feature that is uniformly distributed across all clusters, then the largest cluster would receive the highest intra-feature fraction.
+Consider a feature that is uniformly distributed across all clusters, then the largest cluster would receive the highest intra-feature fraction for that feature.
 
@@
 By highlighting each cluster individually, we can evaluate their distinctiveness in different projections.
-The most influential features shown in Figure \ref{fig:winter_activ_cluster_highlights} are pool/sauna, alpine skiing, museums, and going to the spa.
+The most influential features shown in Figure \ref{fig:winter_activ_cluster_highlights} are pool/sauna, alpine skiing, museums, going to the spa, going for walks and sight seeing.
 The projection roughly separates clusters 1 (blue), 2 (orange), and 3 (green) from each other and the other three clusters (red, violet and brown), which appear to be quite similar in the selected projection.
@@
 
-\subsubsection{Segmentation of travellers travelling without friends}
+\subsubsection{Segmentation of tourists travelling without friends}
 
 In the heatmap displaying the intra-cluster fraction shown in Figure \ref{fig:aus_heatmapg}, we observe that subsets 1, 2, and 6 tend to prefer travelling without friends.
-We can assume that this group consits of tourists travelling alone as couples or with their family.
+We can assume that this group consists of tourists travelling alone as couples or with their family.
 Additional features to further subsegment these groups might be interesting for future surveys.
@@
 The remaining subgroup within Subset 7 consists of families.
-To better cater to this group, packages can be refined to consider that not all museums or festivals are equally suitable for children, enabling more family-friendly customizations.
+To better cater to this group, packages can be refined for instance to consider that not all museums or festivals are equally suitable for children, enabling more family-friendly customizations.
 
@@
 
-Future developments in the software might include expanding the range of interactive features and providing additional visualization methods.
-Because it is built on \proglang{R} and \proglang{Python}, it would also be possible to integrate more closely with the machine learning algorithms, such as generating new cluster solutions directly from the GUI.
-For more complex shapes, containing concavities, or non-linear boundaries, implementation of sliced tours~\citep{Laa2020}, would be recommended.
-The integration of new plot types could also be streamlined to improve the versitility of the package.
-
-Furthermore, \texttt{lionfish} provides an additional demonstration of how \proglang{R} and \proglang{Python} can be integrated to harness the strengths of both programming languages.
+Future developments in \texttt{lionfish} might include expanding the range of visualization methods and offering additional interactive features, such as enabling the generation of new cluster solutions directly from the GUI.
+Leveraging the integration of both \proglang{R} and \proglang{Python}, future enhancements could include a seamless combination of algorithms from each language, which broadens the expansion potential of the software.
+For more complex data shapes, such as those with concavities or non-linear boundaries, the implementation of sliced tours~\citep{Laa2020} is recommended to improve exploratory analysis.
+Streamlining the addition of new plot types could also further enhance the versatility of the package, making it more adaptable for different data visualization needs.
+This ongoing integration of \proglang{R} and \proglang{Python} exemplifies how the strengths of both languages can be harnessed for the development of new software packages.
 

### 2024-10-03 manuscript.tex
shifted reference to website into own section

@@
 This allows for manual exploration of the data.
-
-More information on all the features of the GUI can be found on \href{https://github.com/mmedl94/lionfish}{https://github.com/mmedl94/lionfish} and on \href{https://mmedl94.github.io/lionfish/}{https://mmedl94.github.io/lionfish/}.
 
@@
 
+\section{Resources and supplementary materials}~\label{resources}
+
+The software can be found at  \href{https://mmedl94.github.io/lionfish/}{https://mmedl94.github.io/lionfish/} which contains more documentation on:
+
+\begin{itemize} \itemsep 0in
+\item installation
+\item currently supported interactivity
+\item the variety of plots available
+\end{itemize}
+
+The supplementary materials for this paper include:
+
+\begin{itemize} \itemsep 0in
+\item scripts to start the GUI and reproduce the graphics in the paper
+\item saved state for different parts of the applications, so that the results can be reproduced
+\end{itemize}
+
 \section{Acknowledgements}~\label{acknowledgements}

### 2024-10-04 scripts/load_snapshots.R
added link to supplementary repo

@@ -1,22 +1,8 @@
-# install lionfish
-install.packages("remotes")
-remotes::install_github("mmedl94/lionfish")
 library(lionfish)
 library(dplyr)
 
-library(tourr)
-library(devtools)
-library(reticulate)
-library(flexclust)
-library(stats)
-library(dplyr)
-setwd("/home/matthias/Desktop/Work/gsoc/lionfish")
-load_all()
-
-setwd("/home/matthias/Desktop/Work/gsoc/paper/pytour_paper")
-# set working directory
+# set working directory to path/to/lionfish_article/
 setwd("..")
-setwd("/home/matthias/Desktop/Work/gsoc/paper/pytour_paper")
 init_env()
 # load Austrian Vacation Activities dataset
 data("winterActiv")
@@ -32,6 +18,9 @@ ausActiv_features <- read.csv("saves/aus_saves/init/feature_selection.csv",
                                  header = FALSE)
 ausActiv <- ausActiv[, colnames(ausActiv) %in% ausActiv_features$V1]
 ausActiv <- ausActiv[, ausActiv_features$V1]
+
+# if the display is too large or too small please adjust the display size
+# argument
 
 # Figure 7 + figure 8
 load_interactive_tour(winterActiv, "/saves/aut_saves/init",

### 2024-10-04 manuscript.tex
minor changes
added link to supplementary repo
added images as .pdf

@@
     \centering
-    \includegraphics[width=1\textwidth]{GUI_overview.png}
+    \includegraphics[width=1\textwidth]{GUI_overview.pdf}
     \caption{Overview of the GUI.
@@
 
-%\[
-%\text{Overall fraction}\ f_{ij}^{o} = \frac{c_{ij}}{n}
-%\]
-%with normalization to the total number of observations $n$ for a general overview,
-
-%\[
-%\text{Intra cluster fraction}\  f_{ij}^{clust} = \frac{c_{ij}}{n_{i}}
-%\]
-%showing the counts relative to the cluster's size \(n_i\), to see the fraction of observations in cluster $i$ for which feature $j$ was true, and
-
-%\[
-%\text{Intra feature fraction}\ f_{ij}^{feat} = \frac{c_{ij}}{c_{j}}
-%\]
-%to see how the true values for feature $j$ are distributed across the different clusters relative, with normalization to the overall count of that feature \(c_j = \sum_i c_{ij}\).
-
 \subsection{Subset selection}
@@
     \centering
-    \includegraphics[width=1\textwidth]{winteractiv_heatmap.png}
+    \includegraphics[width=1\textwidth]{winteractiv_heatmap.pdf}
     \caption{Traditional overview of clusters.
@@
     \centering
-    \includegraphics[width=1\textwidth]{silhouette_comparison.png}
+    \includegraphics[width=1\textwidth]{silhouette_comparison.pdf}
     \caption{Comparison of silhouette plots of two $k$-means cluster solutions of the Austrian vacation activities dataset with $k$=6.
@@
     \centering
-    \includegraphics[width=1\textwidth]{winter_activ_cluster_highlights.png}
+    \includegraphics[width=1\textwidth]{winter_activ_cluster_highlights.pdf}
     \caption{Display of a projection of the Austrian vacation activities dataset with the six clusters of the k-means cluster solution highlighted in different colors.
@@
     \centering
-    \includegraphics[width=1\textwidth]{winter_cl7_init.png}
+    \includegraphics[width=1\textwidth]{winter_cl7_init.pdf}
     \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the k-means solution of the Austrian vacation activities dataset.
@@
     \centering
-    \includegraphics[width=1\textwidth]{winter_cl7_pre.png}
+    \includegraphics[width=1\textwidth]{winter_cl7_pre.pdf}
     \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the $k$-means solution of the Austrian vacation activities dataset, with manually adjusted projections.
@@
     \centering
-    \includegraphics[width=1\textwidth]{winter_cl7_post.png}
+    \includegraphics[width=1\textwidth]{winter_cl7_post.pdf}
     \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the $k$-means solution of the Austrian vacation activities dataset, after new subsets have been selected.
@@
     \centering
-    \includegraphics[width=1\textwidth]{aus_feature_clustering.png}
+    \includegraphics[width=1\textwidth]{aus_feature_clustering.pdf}
     \caption{Dendrogram of the features of the Australian Vacation Activities dataset using Ward's method with the Jaccard index.
@@
     \centering
-    \includegraphics[width=1\textwidth]{aus_heatmap.png}
+    \includegraphics[width=1\textwidth]{aus_heatmap.pdf}
     \caption{Heatmap of the intra cluster fraction (indicated by color) of the $k$-means solution with \( k = 6 \) and the reduced feature subset.
@@
     \centering
-    \includegraphics[width=1\textwidth]{aus_preselection.png}
+    \includegraphics[width=1\textwidth]{aus_preselection.pdf}
     \caption{Interactive tour GUI loaded with multiple plots showing different aspects of the $k$-means solution of the Australian vacation activities dataset, with the projection axes of the feature ``Friends'' pointing into one direction and all the other ones into another one.
@@
     \centering
-    \includegraphics[width=1\textwidth]{aus_selection.png}
+    \includegraphics[width=1\textwidth]{aus_selection.pdf}
     \caption{Interactive tour GUI as seen in \ref{fig:aus_preselection}, but after sub-selection of subsets 7 (pink), 8 (grey) and 9 (gold).
@@
 
-The supplementary materials for this paper include:
+The supplementary materials for this paper can be found at \href{https://github.com/mmedl94/lionfish_article/}{https://github.com/mmedl94/lionfish\_article/} and include:
 
